---
title: "02_Analysis"
author: "Sophie Buysse"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Description

This code runs all linear models for the 2021 and 2022 experiments. 

# Set basic info for all code

This chunk sets things that will be constant for the whole document.

```{r, warning = FALSE, message = FALSE}
#setup code here

# read in packages
library(dplyr)
library(ggplot2)
library(lme4)
library(lmerTest)
library(tidyr)
library(rcompanion)
library(plotly)
#library(predictmeans)
library(emmeans)
library(sjPlot)
# prep for doing statistical analyses
options(contrasts = c("contr.sum", "contr.poly"))

# let's make some functions
std_mean <- function(x){
  sd(x, na.rm = TRUE)/sqrt(length(x))
}

# Confidence interval multiplier for standard error
    # Calculate t-statistic for confidence interval: 
    # e.g., if conf.interval is .95, use .975 (above/below), and use df=N-1

conf_int <- function(x, conf.interval = 0.95){
  ciMult <- qt((1+conf.interval)/2, length(x)-1)
  ci <- std_mean(x) * ciMult
  return(ci)
}

```
# 2021

## Read in data

```{r sum1, eval = TRUE}
Dat_2021 <- read.csv("data/CleanData_2021.csv")
# set factors
Dat_2021 <- Dat_2021 %>% dplyr::mutate(
  Population = as.factor(Population),
  Line = as.factor(Line),
  Treatment = as.factor(Treatment),
  Soil.Mix.Batch.Number.x = as.factor(Soil.Mix.Batch.Number.x),
  BranchStructure = as.factor(BranchStructure),
  DoneFlwr = as.factor(DoneFlwr),
  FruitCounter = as.factor(FruitCounter),
  SeedCounter = as.factor(SeedCounter)
  
)
```


## Summarize some data
```{r}
# for the things I can easily get a mean for (num or int). NOTE: These are raw means and confidence intervals, I might not actually want them later on because I will want model predicted means and standard errors.
ByPop_21 <- Dat_2021 %>%
  dplyr::group_by(Population) %>%
  summarize(across(.col = c(Emergence:LeafNum_08262021, RosetteLeafNum_10042021:Repro_to_Ros, LatBranches:IJ_FruitCount, FruitCollected, NumFlwrLeft:IJ_SeedCount, FruitCount_BL, AG_biomass, AvgSeedWt, AvgSeedNum, fitness), .fns = list(mean = ~mean(.x, na.rm = TRUE, .group = "keep"), ci_95 = ~conf_int(.x), n = ~sum(!is.na(.x)))))

# transpose for easier viewing
ByPop_21 <- data.frame(t(ByPop_21))

# now do by treatment
ByTrt_21 <- Dat_2021 %>%
  dplyr::group_by(Treatment) %>%
  summarize(across(.col = c(Emergence:LeafNum_08262021, RosetteLeafNum_10042021:Repro_to_Ros, LatBranches:IJ_FruitCount, FruitCollected, NumFlwrLeft:IJ_SeedCount, FruitCount_BL, AG_biomass, AvgSeedWt, AvgSeedNum, fitness), .fns = list(mean = ~mean(.x, na.rm = TRUE, .group = "keep"), ci_95 = ~conf_int(.x), n = ~sum(!is.na(.x)))))
ByTrt_21 <- data.frame(t(ByTrt_21))

# Now I want population  means within each treatment
ByTrt_Pop_21 <- Dat_2021 %>%
  dplyr::group_by(Population) %>%
  dplyr::group_by(Treatment, .add = TRUE) %>%
  summarize(across(.col = c(Emergence:LeafNum_08262021, RosetteLeafNum_10042021:Repro_to_Ros, LatBranches:IJ_FruitCount, FruitCollected, NumFlwrLeft:IJ_SeedCount, FruitCount_BL, AG_biomass, AvgSeedWt, AvgSeedNum, fitness), .fns = list(mean = ~mean(.x, na.rm = TRUE, .group = "keep"), ci_95 = ~conf_int(.x), n = ~sum(!is.na(.x)))))

# get just the sample sizes
SampleSizes_2021 <- dplyr::select(ByTrt_Pop_21, "Treatment", "Population", ends_with("_n"))
save(SampleSizes_2021, file = "data/SampleSizes_2021.robj")

# for some ease of viewing
ByTrt_Pop_21 <- data.frame(t(ByTrt_Pop_21))
colnames(ByTrt_Pop_21) <- c("Belm_C", "Belm_F", "Roda_C", "Roda_F")

# save this to keep means
write.csv(ByTrt_Pop_21,"data/PopMeansByTreatment_2021.csv", row.names = TRUE )


```
## Two Treatment Anovas

This experiment included a Current/Future treatment where plants were in the current treatment before vernalization and the future treatment after vernalization to see if early heat and drought was important. We are not analyzing this third treatment here and it was removed during the data cleaning step (commented out code here).

The initial model is was follows:
trait ~ Treatment * Population + (1|Population:Line) + (1|Soil.Mix.Batch.Number.x)

When removing the soil mix random effect, most p values and effect sizes were not meaningfully changed with the following 4 exceptions written as to what happened when soil  mix was removed from the model:\
- dry weight of single leaf: population effect increased, p value decreased \
- relative water content: interaction effect increased, p value decreased (makes sense because dry weight changed a little)\
- rosette dry weight: p value cut in half effect size doubles \
- fruit number: population p value increases, interaction p value decreases \

In general these model changes were slight and soil mix explained very little variance - the exceptions are relative water content (soil mix explained half as much variance as genotype), number of fruits, and number of primary stalks (soil mix explained more variance than genotype but over half the variance is still residual). Thus, soil mix was removed as a random variable from the analysis.

To make plotting easier, I want to have one function that runs the model and the model accuracy tests, then outputs the model object. I then want to create a data_table with predict means and run an anova with the model output object. The data tabel and the anova need to be saved to go into a summary table later or to be used for plotting.

```{r stats}
# subset to only the Current and Future treatments
#Dat_2021_TwoTrt <- Dat_2021[which(Dat_2021$Treatment == "Current"| Dat_2021$Treatment == "Future") , ]
Dat_2021_TwoTrt <- Dat_2021
rm(Dat_2021)
# things I want to use as random variables; Line, soil mix batch number
# for single leaf traits also include leaf_Collected
# for traits at harvest I can use DoneFlwr

# function to run models on 2021 data: 
# the model itself is not saved to a global object. the anova table is saved.

# run the mixed model
do_lmer <- function(trait, data = Dat_2021_TwoTrt){
  lm <- lmer(trait ~ Treatment * Population + (1|Population:Line) , data = data, contrasts = list(Treatment=contr.sum, Population = contr.sum))
  print(plot(lm))
  hist(residuals(lm), breaks = 15)
  plot(fitted(lm), residuals(lm, type = "pearson", scaled = TRUE),
       col = c("red", "blue")[as.numeric(Dat_2021_TwoTrt$Population)[complete.cases(trait)]],
       pch = c(16, 15, 17)[as.numeric(Dat_2021_TwoTrt$Treatment)[complete.cases(trait)]])
  #legend("topleft", legend = c("Italy", "Sweden"), col = c("red", "blue"), pch = 16)
  #legend("topright", legend = c("Current", "Future"), col = "black", pch = c(16,17))
  # plotting sanity check
  print(paste0("the number of complete cases is ", sum(complete.cases(trait))))
  qqnorm(resid(lm))
  qqline(resid(lm))
  print(summary(lm))
  print(confint(lm))
  #sample.size <- data %>% group_by(Treatment) %>% group_by(Treatment, .add = TRUE) %>% summarize(across(.cols = trait, .fns= ~sum(!is.na(.x))))
  return(lm)
}
# do the anova
do_anov <- function(model){
  anov <- anova(model)
  print(anov)
  df <- as.data.frame(anov)
  df$exp <- "2021"
  return(df)
}

# get the means table, no backtransforming - can add pairwise to do comparisons within treatment or within population, but not both at the same time. commented out is the old predict means functions
#get_table <- function(model){
#  tab <- predictmeans(model=model, modelterm = "Treatment:Population", plot = FALSE)
#  return(tab$mean_table)
#}
get_table <- function(model){
  tab <- emmeans(model, c("Treatment", "Population"), type = "response")
  return(tab)
}
# function to backtransform - not needed with emmeans
#undo_log10 <- function(x){
#  10^x
#}

#get the means table with backtransforming from log. predict means version is commented out.
#get_table_bt <- function(model){
#  tab <- predictmeans(model=model, modelterm = "Treatment:Population", trans=undo_log10, plot = FALSE)
#  return(tab$mean_table)
#}
# note this ONLY shows the backtransformed values, so the figures with non-transformed data will need to change (and all figures will actually if I commit to this)
get_table_bt <- function(model){
  tab <- get_table(update(ref_grid(model), tran = "log10"))
  return(tab)
}

```

The modelling function runs a mixed model where each trait is predicted by population, treatment, and their interaction while accounting for random effects of line nested within population and soil mix batch number.

If the interaction term in this model is significant, 4 additional models are run to identify differences between population in each treatment and differences between treatments for each population

### Phenology

Days to Emergence

```{r}
# Emergence
emergence_lm_21 <- do_lmer(Dat_2021_TwoTrt$Emergence)
emergence_anov_21 <- do_anov(emergence_lm_21)
emergence_anov_21$trait <- "Emergence"
emergence_emmeans_21 <- get_table(emergence_lm_21)
emergence_means_21 <- as.data.frame(emergence_emmeans_21)
emergence_pairs_21 <- as.data.frame(pairs(emergence_emmeans_21))
emergence_pairs_21$trait <- "Emergence"
```

There is an interaction, so do post-hoc analyses to identify where there are differences between means within treatments and between treatments within populations.


Days between emergence and bolting
no post hoc models because not in paper

```{r}
# days emergence to bolting
bolting_lm_21 <- do_lmer(Dat_2021_TwoTrt$DayToBolt)
bolting_anov_21 <- do_anov(bolting_lm_21)
bolting_anov_21$trait <- "Bolting"
bolting_means_21 <- as.data.frame(get_table(bolting_lm_21))
```

Days between bolting and flowering
no post hoc models because not in paper

```{r}
# days bolting to flowering
flowering_lm_21 <- do_lmer(Dat_2021_TwoTrt$DayToFlwr)
flowering_anov_21 <- do_anov(flowering_lm_21)
flowering_anov_21$trait <- "Bolting To Flowering"
flowering_means_21 <- as.data.frame(get_table(flowering_lm_21))
```

Days between Emergence and flowering

```{r}
# Emergence to flowering
eTof_lm_21 <- do_lmer(Dat_2021_TwoTrt$EmergeToFlwr)
eTof_anov_21 <- do_anov(eTof_lm_21)
eTof_anov_21$trait <- "Emergence To Flowering"
eTof_emmeans_21 <- get_table(eTof_lm_21)
eTof_means_21 <- as.data.frame(eTof_emmeans_21)
eTof_pairs_21 <- as.data.frame(pairs(eTof_emmeans_21))
eTof_pairs_21$trait <- "Emergence To Flowering"
```

### Single Leaf Collected At Flowering

fresh weight
```{r}
# Fresh weight
fresh_lm_21 <- do_lmer(Dat_2021_TwoTrt$l10_FreshWt)
fresh_anov_21 <- do_anov(fresh_lm_21)
fresh_anov_21$trait <- "l10_FreshWt"
fresh_emmeans_21 <- get_table_bt(fresh_lm_21)
fresh_means_21 <- as.data.frame(fresh_emmeans_21)
fresh_pairs_21 <- as.data.frame(pairs(fresh_emmeans_21))
fresh_pairs_21$trait <- "l10_FreshWt"
```

saturated/hydrated weight

```{r}
# Saturated weight
sat_lm_21 <- do_lmer(Dat_2021_TwoTrt$l10_SatWt)
sat_anov_21 <- do_anov(sat_lm_21)
sat_anov_21$trait <- "l10_SatWt"
sat_emmeans_21 <- get_table_bt(sat_lm_21)
sat_means_21 <- as.data.frame(sat_emmeans_21)
sat_pairs_21 <- as.data.frame(pairs(sat_emmeans_21))
sat_pairs_21$trait <- "l10_SatWt"
```


dried weight
```{r}
# Dried weight
dry_lm_21 <- do_lmer(Dat_2021_TwoTrt$l10_DriedWt)
dry_anov_21 <- do_anov(dry_lm_21)
dry_anov_21$trait <- "l10_DriedWt"
dry_emmeans_21 <- get_table_bt(dry_lm_21)
dry_means_21 <- as.data.frame(dry_emmeans_21)
dry_pairs_21 <- as.data.frame(pairs(dry_emmeans_21))
dry_pairs_21$trait <- "l10_DriedWt"
# no significant interaction term

# testing, see below for more
#anov_lmer(Dat_2021_TwoTrt$DriedWt_g)
#anov_lmer(Dat_2021_TwoTrt$SQR_DriedWt)

```

leaf area

```{r}
# Leaf Area
area_lm_21 <- do_lmer(Dat_2021_TwoTrt$l10_LeafArea)
area_anov_21 <- do_anov(area_lm_21)
area_anov_21$trait <- "l10_area"
area_emmeans_21 <- get_table_bt(area_lm_21)
area_means_21 <- as.data.frame(area_emmeans_21)
area_pairs_21 <- as.data.frame(pairs(area_emmeans_21))
area_pairs_21$trait <- "l10_area"
```


leaf perimeter

```{r}
per_lm_21 <- do_lmer(Dat_2021_TwoTrt$l10_LeafPer)
per_anov_21 <- do_anov(per_lm_21)
per_anov_21$trait <- "l10_Perimeter"
per_emmeans_21 <- get_table_bt(per_lm_21)
per_means_21 <- as.data.frame(per_emmeans_21)
per_pairs_21 <- as.data.frame(pairs(per_emmeans_21))
per_pairs_21$trait <- "l10_Perimeter"
```


specific leaf area

```{r}
# SLA
SLA_lm_21 <- do_lmer(Dat_2021_TwoTrt$l10_SLA)
SLA_anov_21 <- do_anov(SLA_lm_21)
SLA_anov_21$trait <- "l10_SLA"
SLA_emmeans_21 <- get_table_bt(SLA_lm_21)
SLA_means_21 <- as.data.frame(SLA_emmeans_21)
SLA_pairs_21 <- as.data.frame(pairs(SLA_emmeans_21))
SLA_pairs_21$trait <- "l10_SLA"
```



leaf dry matter content
```{r}
# LDMC
LDMC_lm_21 <- do_lmer(Dat_2021_TwoTrt$l10_LDMC)
LDMC_anov_21 <- do_anov(LDMC_lm_21)
LDMC_anov_21$trait <- "l10_LDMC"
LDMC_emmeans_21 <- get_table_bt(LDMC_lm_21)
LDMC_means_21 <- as.data.frame(LDMC_emmeans_21)
LDMC_pairs_21 <- as.data.frame(pairs(LDMC_emmeans_21))
LDMC_pairs_21$trait <- "l10_LDMC"
```


relative water content
```{r}
# RWC
RWC_lm_21 <- do_lmer(Dat_2021_TwoTrt$RWC)
RWC_anov_21 <- do_anov(RWC_lm_21)
RWC_anov_21$trait <- "RWC"
RWC_emmeans_21 <- get_table(RWC_lm_21)
RWC_means_21 <- as.data.frame(RWC_emmeans_21)
RWC_pairs_21 <- as.data.frame(pairs(RWC_emmeans_21))
#no interaction (well, weak)
RWC_pairs_21$trait <- "RWC"
```

### Leaf Number
Note that Basia is potentially interested in looking at growth curves or leaf number changes over time wth a shade or photoperiod response but would also probably be interested in a temperature response but I should do a bit more literature searching on that if we go that route. but that is all to say that all these leaf number traits could be the basis of why she does that, and I could go back and see what Tori did too.
 
leaf num at 4 weeks
no post hoc analyses because not in paper
```{r}
# Leaf Num 7/22/21 - a little under 5 weeks post planting on 6/19/2022 - 4 weeks post moving to the chambers on 6/24
LN_PreVern_lm_21 <- do_lmer(Dat_2021_TwoTrt$LeafNum_.07222021)
LN_PreVern_anov_21 <- do_anov(LN_PreVern_lm_21)
LN_PreVern_anov_21$trait <- "LeafNum_4wks"
LN_PreVern_means_21 <- as.data.frame(get_table(LN_PreVern_lm_21))
# no interaction
# diagonal lines in residual plot are artifact of discrete trait. none in left corner is indicative of zeros.
# double check that residual lines all have same value
lm <- lmer(LeafNum_.07222021 ~ Treatment * Population + (1|Population:Line) , data = Dat_2021_TwoTrt, contrasts = list(Treatment=contr.sum, Population = contr.sum))

plot(fitted(lm), residuals(lm, type = "pearson", scaled = TRUE),
       col = c("red", "blue")[as.numeric(Dat_2021_TwoTrt$Population)],
       pch = c(16, 15, 17)[as.numeric(Dat_2021_TwoTrt$Treatment)])
legend("topleft", legend = c("Italy", "Sweden"), col = c("red", "blue"), pch = 16)
legend("topright", legend = c("Current", "Future"), col = "black", pch = c(16,17))
text(fitted(lm), residuals(lm, type = "pearson", scaled = TRUE)-0.15, labels = Dat_2021_TwoTrt$LeafNum_.07222021, cex = 0.7)

```

Leaf num 5 weeks into vernalization
no post hoc analyses because not in paper
```{r}
# LeafNumber from 08262021 - 5 weeks into vernalization... why did I choose this date? I think I was thinning on this day?
LN_Vern_lm_21 <- do_lmer(Dat_2021_TwoTrt$LeafNum_08262021)
LN_Vern_anov_21 <- do_anov(LN_Vern_lm_21)
LN_Vern_anov_21$trait <- "LeafNum_9wks"
LN_Vern_means_21 <- as.data.frame(get_table(LN_Vern_lm_21))
# no interaction
```

Leaf num about when bolting started
no post hoc analyses becuase not in paper - not at a constant developmental stage
```{r}
# Leaf Num October 4 - picked this day because was about when bolting started but some had already bolted by this day
LN_PostVern_lm_21 <- do_lmer(Dat_2021_TwoTrt$RosetteLeafNum_10042021)
LN_PostVern_anov_21 <- do_anov(LN_PostVern_lm_21)
LN_PostVern_anov_21$trait <- "LeafNum_14wks"
LN_PostVern_means_21 <- as.data.frame(get_table(LN_PostVern_lm_21))
# no interaction
```


Leaf number at bolting and flower were only done on some plants, not all (bolting only for SW plants and flowering only for 16 plants), so there are no models or figures for these traits.


Leaf num at harvest
```{r}
# LN at harvest - from all plants - not sure totally accurate count because hard to count at this point in life cycle
LN_harv_lm_21 <- do_lmer(Dat_2021_TwoTrt$RosetteLeafNum)
LN_harv_anov_21 <- do_anov(LN_harv_lm_21)
LN_harv_anov_21$trait <- "LeafNum_harvest"
LN_harv_emmeans_21 <- get_table(LN_harv_lm_21)
LN_harv_means_21 <- as.data.frame(LN_harv_emmeans_21)
LN_harv_pairs_21 <- as.data.frame(pairs(LN_harv_emmeans_21))
LN_harv_pairs_21$trait <- "LeafNum_harvest"
# no interaction
```

### Biomass

Any biomass traits and biomass allocation

Rosette biomass
```{r}
# dry rosette biomass
Ros_lm_21 <- do_lmer(Dat_2021_TwoTrt$l10_DryRosG)
Ros_anov_21 <- do_anov(Ros_lm_21)
Ros_anov_21$trait <- "l10_DryRosG"
Ros_emmeans_21 <- get_table_bt(Ros_lm_21)
Ros_means_21 <- as.data.frame(Ros_emmeans_21)
Ros_pairs_21 <- as.data.frame(pairs(Ros_emmeans_21))
Ros_pairs_21$trait <- "l10_DryRosG"
#no interaction

## check how it compares to the sqr transformed
#Ros_lm_21SR <- do_lmer(Dat_2021_TwoTrt$SQR_DryRosG)
#Ros_anov_21SR <- do_anov(Ros_lm_21SR)
#Ros_anov_21SR$trait <- "SQR_DryRosG"
#Ros_means_21SR <- get_table_bt(Ros_lm_21SR)
## commented out b/c no difference in signficance levels.

```

Reproductive biomass
```{r}
# dry reproductive biomass
Repro_lm_21 <- do_lmer(Dat_2021_TwoTrt$l10_DryReproG)
Repro_anov_21 <- do_anov(Repro_lm_21)
Repro_anov_21$trait <- "l10_DryReproG"
Repro_emmeans_21 <- get_table_bt(Repro_lm_21)
Repro_means_21 <- as.data.frame(Repro_emmeans_21)
Repro_pairs_21 <- as.data.frame(pairs(Repro_emmeans_21))
Repro_pairs_21$trait <- "l10_DryReproG"
```


Above ground biomass (rosette + reproductive)
no post hoc analyses because not in paper
```{r}
# dry above ground biomass (ros + repro)
AG_lm_21 <- do_lmer(Dat_2021_TwoTrt$l10_AG_biomass)
AG_anov_21 <- do_anov(AG_lm_21)
AG_anov_21$trait <- "l10_AG_biomass"
AG_means_21 <- as.data.frame(get_table_bt(AG_lm_21))
```

Skipping dry root biomass because only have for a few plants. Also skipping root to shoot for the same reason.

Above ground biomass allocation (repro/ros)
```{r}
# reproductive biomass divided by rosette biomass - above ground biomass allocation
RR_lm_21 <- do_lmer(Dat_2021_TwoTrt$l10_R_to_R)
RR_anov_21 <- do_anov(RR_lm_21)
RR_anov_21$trait <- "l10_Repro_to_Ros"
RR_emmeans_21 <- get_table_bt(RR_lm_21)
RR_means_21 <- as.data.frame(RR_emmeans_21)
RR_pairs_21 <- as.data.frame(pairs(RR_emmeans_21))
RR_pairs_21$trait <- "l10_Repro_to_Ros"
```


### Plant Structure
branches and height. Branch structure is about damage so skip. but maybe should be a random effect in other models?

Num lateral branches
```{r}
# number of lateral branches
LatBranch_lm_21 <- do_lmer(Dat_2021_TwoTrt$LatBranches)
LatBranch_anov_21 <- do_anov(LatBranch_lm_21)
LatBranch_anov_21$trait <- "Lateral Branches"
LatBranch_emmeans_21 <- get_table(LatBranch_lm_21)
LatBranch_means_21 <- as.data.frame(LatBranch_emmeans_21)
LatBranch_pairs_21 <- as.data.frame(pairs(LatBranch_emmeans_21))
LatBranch_pairs_21$trait <- "Lateral Branches"
```

Num primary stalks
```{r}
# number of primary stalks
PrimStalks_lm_21 <- do_lmer(Dat_2021_TwoTrt$PrimaryStalks)
PrimStalks_anov_21 <- do_anov(PrimStalks_lm_21)
PrimStalks_anov_21$trait <- "Primary Stalks"
PrimStalks_emmeans_21 <- get_table(PrimStalks_lm_21)
PrimStalks_means_21 <- as.data.frame(PrimStalks_emmeans_21)
PrimStalks_pairs_21 <- as.data.frame(pairs(PrimStalks_emmeans_21))
PrimStalks_pairs_21$trait <- "Primary Stalks"
```

Height main stalk
```{r}
# height of main stalk (cm)
height_lm_21 <- do_lmer(Dat_2021_TwoTrt$Height_cm)
height_anov_21 <- do_anov(height_lm_21)
height_anov_21$trait <- "Height (cm)"
height_emmeans_21 <- get_table(height_lm_21)
height_means_21 <- as.data.frame(height_emmeans_21)
height_pairs_21 <- as.data.frame(pairs(height_emmeans_21))
height_pairs_21$trait <- "Height (cm)"
```


### Fitness

Fruit Count
at first, only using IJ fruit count because of issue with skipping counter for other fruit count.

Now (9/20/2024) only using Basia's fruit counts

```{r, eval = FALSE, echo = FALSE}
# number fruits on plant
# maybe would want to add done flower or not done flower as a random effect in this model?
fruit_lm_21 <- do_lmer(Dat_2021_TwoTrt$l10_IJ_Fruits)
fruit_anov_21 <- do_anov(fruit_lm_21)
fruit_anov_21$trait <- "l10_FruitNum"
fruit_means_21 <- get_table_bt(fruit_lm_21)

## test to see if adding fruit counter changes the model results
#tmp_fruit <- lmer(l10_IJ_Fruits ~ Treatment * Population + (1|Population:Line) + (1|FruitCounter), data = Dat_2021_TwoTrt, #contrasts = list(Treatment=contr.sum, Population = contr.sum))
#
#  hist(residuals(tmp_fruit), breaks = 15)
#  plot(fitted(tmp_fruit), residuals(tmp_fruit, type = "pearson", scaled = TRUE),
#       col = c("red", "blue")[as.numeric(Dat_2021_TwoTrt$Population)[complete.cases(Dat_2021_TwoTrt$l10_IJ_Fruits)]],
#       pch = c(16, 15, 17)[as.numeric(Dat_2021_TwoTrt$Treatment)[complete.cases(Dat_2021_TwoTrt$l10_IJ_Fruits)]])
#    # plotting sanity check
#  print(paste0("the number of complete cases is ", sum(complete.cases(Dat_2021_TwoTrt$l10_IJ_Fruits))))
#  #title(trait)
#  print(summary(tmp_fruit))
#  anova(tmp_fruit)
#  
## no meaningful changes, so FruitCounter will not be a random effect and this code is commented out.

```

```{r}
# number fruits on plant
# maybe would want to add done flower or not done flower as a random effect in this model?
fruit_lm_21 <- do_lmer(Dat_2021_TwoTrt$FruitCount_BL)
fruit_anov_21 <- do_anov(fruit_lm_21)
fruit_anov_21$trait <- "FruitNum"
fruit_emmeans_21 <- get_table(fruit_lm_21)
fruit_means_21 <- as.data.frame(fruit_emmeans_21)
fruit_pairs_21 <- as.data.frame(pairs(fruit_emmeans_21))
fruit_pairs_21$trait <- "FruitNum"

```

```{r}
# try a hurdle model
library(pscl)
mod.hurdle <- hurdle(FruitCount_BL ~ Treatment * Population  , data = Dat_2021_TwoTrt, dist = "poisson", zero.dist = "binomial") #+ (1|Population:Line) contrasts = list(Treatment=contr.sum, Population = contr.sum)
# same as this:
# mod.hurdle <- hurdle(visits ~ ., data = nmes, dist = "poisson", 
#                          zero.dist = "binomial")
summary(mod.hurdle)
hist(residuals(mod.hurdle), breaks = 15) # decent
plot(fitted(mod.hurdle), residuals(mod.hurdle, type = "pearson", scaled = TRUE)) # this looks very similar


mod.zeroinf <- zeroinfl(FruitCount_BL ~ Treatment * Population  , data = Dat_2021_TwoTrt, dist = "poisson")
summary(mod.zeroinf)

# can I match these at all if I remove zeros from my initial model?
mod.nozero <- lmer(FruitCount_BL ~ Treatment * Population + (1|Population:Line) , data = Dat_2021_TwoTrt[Dat_2021_TwoTrt$FruitCount_BL >0, ],contrasts = list(Treatment=contr.sum, Population = contr.sum))
summary(mod.nozero)
hist(residuals(mod.nozero), breaks = 15) # this is more normal
plot(fitted(mod.nozero), residuals(mod.nozero, type = "pearson", scaled = TRUE)) # this looks very similar

# try to get random component back in there
library(glmmTMB)
mod.zeroinf2 <- glmmTMB(FruitCount_BL ~ Treatment * Population + (1|Population:Line), data = Dat_2021_TwoTrt, ziformula = ~Treatment * Population, family = poisson)
summary(mod.zeroinf2)
mod.poisson <- glmmTMB(FruitCount_BL ~ Treatment * Population + (1|Population:Line), data = Dat_2021_TwoTrt, ziformula = ~0, family = poisson)
summary(mod.poisson)
hist(residuals(mod.poisson), breaks = 15) # okay
plot(fitted(mod.poisson), residuals(mod.poisson, type = "pearson")) #, scaled = TRUE; I think this looks better but not sure what taking otu scaled as an argument did
plot(fitted(fruit_lm_21), residuals(fruit_lm_21, type = "pearson"))
# okay, so if I fit a poisson model is really where the changes happen, not for adding inthe zero inflated component


# get emmeans from the poisson
#mod.poisson_anov <- do_anov(mod.poisson) # can't do anova on glmmTMB
#mod.poisson_anov$trait <- "FruitNum_Pois"
mod.poisson_emmeans <- get_table(mod.poisson)
mod.poisson_means <- as.data.frame(mod.poisson_emmeans)
mod.poisson_pairs <- as.data.frame(pairs(mod.poisson_emmeans))
mod.poisson_pairs$trait <- "FruitNumP_Pois"

# tester fig
ggplot(data = mod.poisson_means, aes(x = Treatment, y = rate)) +
  geom_point(aes(col=Population, shape = Population), size = 2, position = position_dodge(width = 0.3), show.legend = TRUE) +
  geom_errorbar(aes(ymin=asymp.LCL, ymax = asymp.UCL, col = Population), width = 0.1, size = 0.8, position = position_dodge(width = 0.3), show.legend = TRUE)+
  geom_line(aes(group = Population, col = Population), size = 1, position = position_dodge(width = 0.3))+ 
  labs(y="Fruit Count Poisson Model (rate)", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_x_discrete(expand = c(0.6,0))



# zero inflated poisson
mod.zeroinf2_emmeans <- get_table(mod.zeroinf2)
mod.zeroinf2_means <- as.data.frame(mod.zeroinf2_emmeans)
mod.zeroinf2_pairs <- as.data.frame(pairs(mod.zeroinf2_emmeans))
mod.zeroinf2_pairs$trait <- "FruitNumP_PoisZI"

# tester fig
ggplot(data = mod.zeroinf2_means, aes(x = Treatment, y = rate)) +
  geom_point(aes(col=Population, shape = Population), size = 2, position = position_dodge(width = 0.3), show.legend = TRUE) +
  geom_errorbar(aes(ymin=asymp.LCL, ymax = asymp.UCL, col = Population), width = 0.1, size = 0.8, position = position_dodge(width = 0.3), show.legend = TRUE)+
  geom_line(aes(group = Population, col = Population), size = 1, position = position_dodge(width = 0.3))+ 
  labs(y="Fruit Count Poisson ZI Model (rate)", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_x_discrete(expand = c(0.6,0))
```


Average weight of a seed (collected seed weight / count of collected seeds)
```{r}
# weight of collected seeds - take weight divided by number of seeds counted
AvSeedWt_lm_21 <- do_lmer(Dat_2021_TwoTrt$AvgSeedWt)
AvSeedWt_anov_21 <- do_anov(AvSeedWt_lm_21)
AvSeedWt_anov_21$trait <- "Average Seed Wt"
AvSeedWt_emmeans_21 <- get_table(AvSeedWt_lm_21)
AvSeedWt_means_21 <- as.data.frame(AvSeedWt_emmeans_21)
AvSeedWt_pairs_21 <- as.data.frame(pairs(AvSeedWt_emmeans_21))
AvSeedWt_pairs_21$trait <- "Average Seed Wt"

```


average Seed number per fruit -
```{r}
# number seeds from collected fruits

#seeds_2021 <- anov_lmer(Dat_2021_TwoTrt$AvgSeedNum)
seeds_lm_21 <- do_lmer(Dat_2021_TwoTrt$l10_AvgSeedNum)
seeds_anov_21 <- do_anov(seeds_lm_21)
seeds_anov_21$trait <- "l10_Average Seeds per Fruit"
seeds_emmeans_21 <- get_table_bt(seeds_lm_21)
seeds_means_21 <- as.data.frame(seeds_emmeans_21)
seeds_pairs_21 <- as.data.frame(pairs(seeds_emmeans_21))
seeds_pairs_21$trait <- "l10_Average Seeds per Fruit"
# why are residuals so odd? because population explains 0 variance.

## test if SeedCounter matters
#seeds_lm <- lmer(l10_AvgSeedNum ~ Treatment * Population + (1|Population:Line) + (1|SeedCounter), data = Dat_2021_TwoTrt, #contrasts = list(Treatment=contr.sum, Population = contr.sum))
#plot(fitted(seeds_lm), residuals(seeds_lm, type = "pearson", scaled = TRUE),
#       col = c("red", "blue")[as.numeric(Dat_2021_TwoTrt$Population)[complete.cases(Dat_2021_TwoTrt$l10_AvgSeedNum)]],
#       pch = c(16, 15, 17)[as.numeric(Dat_2021_TwoTrt$Treatment)[complete.cases(Dat_2021_TwoTrt$l10_AvgSeedNum)]])
#summary(seeds_lm)
#anova(seeds_lm)
#
## when including seed counter, it does explain some variance (0.0001) but there is more residual (0.0004) and population #still explains nothing. The significance levels don't change, and the effect sizes don't really change either though not the #standard error and df are different for each predictor
## so, not going to include in the model.
#
## see if the nesting is doing something weird. if LineID was the random variable, does anything change?
#Dat_2021_TwoTrt$Line.ID <- as.factor(paste0(Dat_2021_TwoTrt$Population, Dat_2021_TwoTrt$Line))
#seeds_lm2 <- lmer(l10_AvgSeedNum ~ Treatment * Population + (1|Line.ID), data = Dat_2021_TwoTrt, contrasts = #list(Treatment=contr.sum, Population = contr.sum))
#plot(fitted(seeds_lm2), residuals(seeds_lm2, type = "pearson", scaled = TRUE),
#       col = c("red", "blue")[as.numeric(Dat_2021_TwoTrt$Population)[complete.cases(Dat_2021_TwoTrt$l10_AvgSeedNum)]],
#       pch = c(16, 15, 17)[as.numeric(Dat_2021_TwoTrt$Treatment)[complete.cases(Dat_2021_TwoTrt$l10_AvgSeedNum)]])
#summary(seeds_lm2)
#anova(seeds_lm2)
#
## no difference, as expected if the model was functioning as expected. All testing code is commented out and no changes were made to the model.

```

Bonus question about a trade off - are seed weight and seeds per fruit negatively correlated?
```{r}
# overall
cor.test(Dat_2021_TwoTrt$AvgSeedNum, Dat_2021_TwoTrt$AvgSeedWt)
plot(Dat_2021_TwoTrt$AvgSeedNum, Dat_2021_TwoTrt$AvgSeedWt)

# just in SW
cor.test(Dat_2021_TwoTrt[Dat_2021_TwoTrt$Population == "RODA", ]$AvgSeedNum, Dat_2021_TwoTrt[Dat_2021_TwoTrt$Population == "RODA", ]$AvgSeedWt)
plot(Dat_2021_TwoTrt[Dat_2021_TwoTrt$Population == "RODA", ]$AvgSeedNum, Dat_2021_TwoTrt[Dat_2021_TwoTrt$Population == "RODA", ]$AvgSeedWt)

# just in IT
cor.test(Dat_2021_TwoTrt[Dat_2021_TwoTrt$Population == "BELM", ]$AvgSeedNum, Dat_2021_TwoTrt[Dat_2021_TwoTrt$Population == "BELM", ]$AvgSeedWt)
plot(Dat_2021_TwoTrt[Dat_2021_TwoTrt$Population == "BELM", ]$AvgSeedNum, Dat_2021_TwoTrt[Dat_2021_TwoTrt$Population == "BELM", ]$AvgSeedWt)

# just cur?
cor.test(Dat_2021_TwoTrt[Dat_2021_TwoTrt$Treatment == "Current", ]$AvgSeedNum, Dat_2021_TwoTrt[Dat_2021_TwoTrt$Treatment == "Current", ]$AvgSeedWt)
plot(Dat_2021_TwoTrt[Dat_2021_TwoTrt$Treatment == "Current", ]$AvgSeedNum, Dat_2021_TwoTrt[Dat_2021_TwoTrt$Treatment == "Current", ]$AvgSeedWt)

# just fut?
cor.test(Dat_2021_TwoTrt[Dat_2021_TwoTrt$Treatment == "Future", ]$AvgSeedNum, Dat_2021_TwoTrt[Dat_2021_TwoTrt$Treatment == "Future", ]$AvgSeedWt)
plot(Dat_2021_TwoTrt[Dat_2021_TwoTrt$Treatment == "Future", ]$AvgSeedNum, Dat_2021_TwoTrt[Dat_2021_TwoTrt$Treatment == "Future", ]$AvgSeedWt)
```



Total Fitness

```{r}
# total fitness 
fitness_lm_21 <- do_lmer(Dat_2021_TwoTrt$fitness)
fitness_anov_21 <- do_anov(fitness_lm_21)
fitness_anov_21$trait <- "Total Fitness"
fitness_emmeans_21 <- get_table(fitness_lm_21)
fitness_means_21 <- as.data.frame(fitness_emmeans_21)
fitness_pairs_21 <- as.data.frame(pairs(fitness_emmeans_21))
fitness_pairs_21$trait <- "Total Fitness"
```


###Outputs

1) Results table

Messing around with the stargazer package, but it ultimately works for model outputs but not for anova tables.
```{r, eval = FALSE}
seeds_lm <- lmer(l10_AvgSeedNum ~ Treatment * Population + (1|Population:Line) + (1|SeedCounter), data = Dat_2021_TwoTrt, contrasts = list(Treatment=contr.sum, Population = contr.sum))
anova(seeds_lm)
class(seeds_lm) <- "lmerMod"

stargazer(seeds_lm, align = TRUE, type = "text")

stargazer(RR_2021,RWC_2021, type = "text")
```

also tried with sjPlot but again is with model outputs and doesn't look great b/c just 'trait' and 'pop1' and 'treat1' instead of more helfpul terms. 
```{r, eval = FALSE}
sjPlot::tab_model(AG_lm_21, area_lm_21, AvSeedWt_lm_21, bolting_lm_21, dry_lm_21, emergence_lm_21, eTof_lm_21, fitness_lm_21, flowering_lm_21, fresh_lm_21, fruit_lm_21, height_lm_21, LatBranch_lm_21, LDMC_lm_21, LN_harv_lm_21, LN_PostVern_lm_21, LN_PreVern_lm_21, LN_Vern_lm_21, per_lm_21, PrimStalks_lm_21, Repro_lm_21, Ros_lm_21, RR_lm_21, RWC_lm_21, sat_lm_21, seeds_lm_21, SLA_lm_21) 
# would actually work to make a table with only p values
sjPlot::tab_model(anova(RWC_lm_21), anova(LDMC_lm_21), show.r2 = FALSE)
```

I made a bunch of anova tables. Can I simply rbind them all together? Current plan is to r bind all these together, r bind all the 2022 models together, and then cbind the 2021 and 2022 models together.

```{r}
AnovaResults_2021 <- rbind(emergence_anov_21, bolting_anov_21, flowering_anov_21, eTof_anov_21, fresh_anov_21, sat_anov_21, dry_anov_21, area_anov_21, per_anov_21, SLA_anov_21, LDMC_anov_21, RWC_anov_21, LN_PreVern_anov_21, LN_Vern_anov_21, LN_PostVern_anov_21, LN_harv_anov_21, Ros_anov_21, Repro_anov_21, RR_anov_21, LatBranch_anov_21, PrimStalks_anov_21, height_anov_21, fruit_anov_21, AG_anov_21, AvSeedWt_anov_21, seeds_anov_21, fitness_anov_21)

write.csv(AnovaResults_2021, file = "data/AnovaResults_2021.csv", row.names = TRUE)
```

2) All the tables to read in to the figures code

```{r}
# change column names - order is odd as artifact from updating code
dfs <- c("emergence_means_21", "bolting_means_21", "flowering_means_21", "eTof_means_21", "RWC_means_21", 'LN_PreVern_means_21', 'LN_Vern_means_21', "LN_PostVern_means_21", "LN_harv_means_21", 'LatBranch_means_21', "PrimStalks_means_21", "height_means_21", "AvSeedWt_means_21", "fruit_means_21", "fitness_means_21", "fresh_means_21", 'sat_means_21', "dry_means_21", "area_means_21", "per_means_21", "SLA_means_21", "LDMC_means_21", "Ros_means_21", "Repro_means_21", 'RR_means_21', "AG_means_21", "seeds_means_21")

for (i in dfs) {
  x=get(i)
  colnames(x) <- c("Treatment", "Population", "Mean", "SE", "Df", "LL_95", "UL_95")
  assign(i,x)
}

# create list of data frames
Means_2021 <- list(emergence_means_21, bolting_means_21, flowering_means_21, eTof_means_21, fresh_means_21, sat_means_21, dry_means_21, area_means_21, per_means_21, SLA_means_21, LDMC_means_21, RWC_means_21, LN_PreVern_means_21, LN_Vern_means_21, LN_PostVern_means_21, LN_harv_means_21, Ros_means_21, Repro_means_21, RR_means_21, LatBranch_means_21, PrimStalks_means_21, height_means_21, fruit_means_21, AG_means_21, AvSeedWt_means_21, seeds_means_21, fitness_means_21)

# name that list
names(Means_2021) <- c("emergence_means_21", "bolting_means_21", "flowering_means_21", "eTof_means_21", "fresh_means_21", 'sat_means_21', "dry_means_21", "area_means_21", "per_means_21", "SLA_means_21", "LDMC_means_21", "RWC_means_21", 'LN_PreVern_means_21', 'LN_Vern_means_21', "LN_PostVern_means_21", "LN_harv_means_21", "Ros_means_21", "Repro_means_21", 'RR_means_21', 'LatBranch_means_21', "PrimStalks_means_21", "height_means_21", "fruit_means_21", "AG_means_21", "AvSeedWt_means_21", "seeds_means_21", "fitness_means_21")

# save named list
save(Means_2021, file = "data/ModelMeans_2021.robj")

```

Save a .csv with the posthoc results
```{r}
PostHocResults_2021 <- rbind(emergence_pairs_21, eTof_pairs_21, fresh_pairs_21, sat_pairs_21, dry_pairs_21, area_pairs_21, per_pairs_21, SLA_pairs_21, LDMC_pairs_21, RWC_pairs_21, LN_harv_pairs_21, Ros_pairs_21, Repro_pairs_21, RR_pairs_21, LatBranch_pairs_21, PrimStalks_pairs_21, height_pairs_21, fruit_pairs_21, AvSeedWt_pairs_21, seeds_pairs_21, fitness_pairs_21)

write.csv(PostHocResults_2021, file = "data/PostHocResults_2021.csv", row.names = TRUE)
```


## Adaptive?
not sure any of this code is useful. doesn't include correlated traits just looks at plasticity. None of this is being run in July 2024.
```{r, eval = FALSE}
CleanData_TwoTrt$Line.ID <- as.factor(paste0(CleanData_TwoTrt$Population, CleanData_TwoTrt$Line))

plasticity <- CleanData_TwoTrt %>% group_by(Treatment) %>% group_by(Line.ID, .add = TRUE) %>%summarize(mean_SLA = mean(SLA, na.rm = TRUE), mean_RR = mean(Repro_to_Ros, na.rm = TRUE), mean_height = mean(Height_cm, na.rm = TRUE), mean_fruit = mean(TotFruit, na.rm = TRUE), mean_emerge = mean(Emergence, na.rm = TRUE),mean_bolt = mean(DayToBolt, na.rm = TRUE), mean_LDMC = mean(LDMC, na.rm = TRUE), mean_RWC = mean(RWC, na.rm = TRUE), mean_flwr = mean(DayToFlwr, na.rm = TRUE), mean_dehis = mean(DayToDehis, na.rm = TRUE), mean_LeafNum = mean(RosetteLeafNum, na.rm = TRUE), mean_RosWt = mean(DryRosetteG, na.rm = TRUE), mean_LB = mean(LatBranches, na.rm = TRUE), mean_PS = mean(PrimaryStalks, na.rm = TRUE), )

# need to calculate plasticity for each line.

# first calcualte relative fitness
fitness <- CleanData_TwoTrt %>% group_by(Treatment) %>% group_by(Line.ID, .add = TRUE) %>%
  summarize(mean_fruits = mean(TotFruit, na.rm = TRUE))
tmp2 <- merge(fitness, plasticity, by = c("Treatment", "Line.ID"))
tmp2$Population <- substr(tmp2$Line.ID,1,4)
# test with SLA
Fut_SLA <- tmp2[tmp2$Treatment == "Future", c("Treatment", "Line.ID", "mean_fruits", "mean_SLA", "Population")]
Cur_SLA <- tmp2[tmp2$Treatment == "Current", c("Treatment", "Line.ID", "mean_fruits", "mean_SLA", "Population")]

# now everything:
Fut_tmp <- tmp2[tmp2$Treatment == "Future", ]
Cur_tmp <- tmp2[tmp2$Treatment == "Current", ]
# make fitness relative
Fut_tmp <- cbind(Fut_tmp, "rel_fitness" = Fut_tmp$mean_fruits/mean(Fut_tmp$mean_fruits))
Cur_tmp <- cbind(Cur_tmp, "rel_fitness" = Cur_tmp$mean_fruits/mean(Cur_tmp$mean_fruits))
# and standardize
Fut_tmp[4:17] <- scale(Fut_tmp[4:17], center = TRUE, scale = TRUE)

Cur_tmp[4:17] <- scale(Cur_tmp[4:17], center = TRUE, scale = TRUE)

# test on just SLA
summary(lm(rel_fitness ~ mean_SLA * Population, data = Fut_tmp))
# so not significant, but trait term is barely positive, so positive selection
summary(lm(rel_fitness ~ mean_SLA * Population, data = Cur_tmp))
# hey, trait term is slightly negative! so different directions in different environments.
# trait and pop terms are also significant.

# now I want to include like all the traits
# the interaction is only for the population term... maybe I need to fit each population separately? or type * population a million times?
# I probably want to standardize everything too because the scales are so different.
fit_fut <- lm(rel_fitness ~ mean_SLA*Population + mean_RR*Population + mean_height*Population + mean_emerge*Population + mean_bolt*Population + mean_LDMC*Population + mean_RWC*Population + mean_flwr*Population + mean_dehis*Population + mean_LeafNum*Population + mean_RosWt*Population + mean_LB*Population + mean_PS*Population, data = Fut_tmp)
summary(fit_fut)
# ahahahahaha the R2 is one. so some super overfitting which I already knew, and maybe the interaction effects are actually wrong too.

# I think the model should actually be quite a bit more complicated?
# I can also tell my committee that i haven't done these yet. I did a full regression for SLA but I haven't done partial regressions.

tmp3 <- merge(Fut_tmp, Cur_tmp, by = c("Line.ID"))
# not the cleanest data frame but whatever
tmp3$mean_fit <- rowMeans(tmp3[, c("mean_fruits.x", "mean_fruits.y")], na.rm = TRUE)
# difference is current minus future

# calculate plasticity for SLA. I would need to do this for every trait which seems like a lot of work
tmp3$dif_SLA <- tmp3$mean_SLA.y - tmp3$mean_SLA.x
#tmp3 <- tmp3[,c("Line.ID", "Treatment.x", "Population.x", "mean_fruits.x", "mean_fit", "dif_SLA")]
## mean fitness
#summary(lm(mean_fit ~ dif_SLA * Population.x, data = tmp3))
## negative and significant

# future fitness
summary(lm(rel_fitness.x ~ dif_SLA + mean_SLA.x + Population.x + dif_SLA:Population.x + mean_SLA.x:Population.x, data = tmp3))
# negative on trait and on plasticity but not significant


ggplot(data = tmp3, aes(x = dif_SLA, y = mean_fit)) +
  geom_point(aes(col=Population.x), size = 4) +
  labs(y="Mean fitness", 
      x="SLA plasticity")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("maroon3", "darkgreen"))+
  scale_linetype_manual(name = "Population",
                        labels = c("IT", "SW"),
                        values = c("solid", "dotted"))+
  theme_classic() +
  theme(
    legend.title = element_text(family = "serif", color = "black", size = 20),
    legend.text = element_text(family = "serif", color = "black", size = 20),
    axis.title = element_text(family = "serif", color = "black", size = 20),
    axis.text = element_text(family = "serif", color = "black", size = 20),
    axis.line = element_line(colour = 'black', size = 1),
    axis.ticks = element_line(color = 'black', size = 1),
    axis.ticks.length=unit(.15, "cm"),
    legend.spacing.y = unit(0.03, "cm"),
    panel.border = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank()
    )

ggplot(data = tmp3, aes(x = dif_SLA, y = mean_fruits.x)) +
  geom_point(aes(col=Population.x), size = 4) +
  labs(y="Future fitness", 
      x="SLA plasticity")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("maroon3", "darkgreen"))+
  scale_linetype_manual(name = "Population",
                        labels = c("IT", "SW"),
                        values = c("solid", "dotted"))+
  theme_classic() +
  theme(
    legend.title = element_text(family = "serif", color = "black", size = 20),
    legend.text = element_text(family = "serif", color = "black", size = 20),
    axis.title = element_text(family = "serif", color = "black", size = 20),
    axis.text = element_text(family = "serif", color = "black", size = 20),
    axis.line = element_line(colour = 'black', size = 1),
    axis.ticks = element_line(color = 'black', size = 1),
    axis.ticks.length=unit(.15, "cm"),
    legend.spacing.y = unit(0.03, "cm"),
    panel.border = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank()
    )

ggplot(data = Fut_SLA, aes(x = mean_SLA, y = mean_fruits)) +
  geom_point(aes(col=Population), size = 4) +
  labs(y="Future Fruit Number", 
      x="Future SLA")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("maroon3", "darkgreen"))+
  scale_linetype_manual(name = "Population",
                        labels = c("IT", "SW"),
                        values = c("solid", "dotted"))+
  theme_classic() +
  theme(
    legend.title = element_text(family = "serif", color = "black", size = 20),
    legend.text = element_text(family = "serif", color = "black", size = 20),
    axis.title = element_text(family = "serif", color = "black", size = 20),
    axis.text = element_text(family = "serif", color = "black", size = 20),
    axis.line = element_line(colour = 'black', size = 1),
    axis.ticks = element_line(color = 'black', size = 1),
    axis.ticks.length=unit(.15, "cm"),
    legend.spacing.y = unit(0.03, "cm"),
    panel.border = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank()
    )

ggplot(data = Cur_SLA, aes(x = mean_SLA, y = mean_fruits)) +
  geom_point(aes(col=Population), size = 4) +
  labs(y="Current Fruit Number", 
      x="Current SLA")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("maroon3", "darkgreen"))+
  scale_linetype_manual(name = "Population",
                        labels = c("IT", "SW"),
                        values = c("solid", "dotted"))+
  theme_classic() +
  theme(
    legend.title = element_text(family = "serif", color = "black", size = 20),
    legend.text = element_text(family = "serif", color = "black", size = 20),
    axis.title = element_text(family = "serif", color = "black", size = 20),
    axis.text = element_text(family = "serif", color = "black", size = 20),
    axis.line = element_line(colour = 'black', size = 1),
    axis.ticks = element_line(color = 'black', size = 1),
    axis.ticks.length=unit(.15, "cm"),
    legend.spacing.y = unit(0.03, "cm"),
    panel.border = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank()
    )
```

## Sample size information

The goal here is to know how many lines and how many replicates per line for this experiment.

What I want is a count of total plants, a mean, min, and max of plants per line, and the number of lines per population

I got the number of observations that were not NA from earlier in the summarize section

```{r}
# sample size of 2021 total
Dat_2021_TwoTrt %>% count(Treatment)
# 50 plants per treatments
Dat_2021_TwoTrt %>% count(Treatment, Population)
# in each treatment, 20 are IT, 30 are SW

# from models know that there are 21 unique lines. 8 IT, 13 SW.

n_2021 <- Dat_2021_TwoTrt %>%
  group_by(Treatment) %>%
  group_by(Population, .add = TRUE) %>%
  group_by(Line, .add = TRUE) %>%
  count

# so only line with 1 rep is Belm 1. Then everything has 2 but RIL parents. B12 has 7, R47 has 6.

n_2021 %>% 
  group_by(Treatment) %>%
  group_by(Population, .add = TRUE) %>%
  summarize(min = min(n), mean = mean(n), median = median(n), max = max(n))

```


# 2022

## Read in data
```{r}
Dat_2022 <- read.csv("data/CleanData_2022.csv", header= TRUE)

# make some things factors
Dat_2022$Treatment <- as.factor(Dat_2022$Treatment)
Dat_2022$Population <- as.factor(Dat_2022$Population)
Dat_2022$PotID <- as.factor(Dat_2022$PotID)
Dat_2022$Flat <- as.factor(Dat_2022$Flat)
Dat_2022$Chamber <- as.factor(Dat_2022$Chamber)
Dat_2022$Transplanted <- as.factor(Dat_2022$Transplanted)

```

## Summarize!

```{r sum}
# by population only
ByPop_22 <- Dat_2022 %>%
  dplyr::group_by(Population) %>%
  summarize(across(.col = c(DaysToEmergence:l10_Stomata_density), .fns = list(mean = ~mean(.x, na.rm = TRUE, .group = "keep"), ci_95 = ~conf_int(.x), n = ~sum(!is.na(.x)))))

# transpose for easier viewing
ByPop_22 <- data.frame(t(ByPop_22))

# now do by treatment
ByTrt_22 <- Dat_2022 %>%
  dplyr::group_by(Treatment) %>%
  summarize(across(.col = c(DaysToEmergence:l10_Stomata_density), .fns = list(mean = ~mean(.x, na.rm = TRUE, .group = "keep"), ci_95 = ~conf_int(.x), n = ~sum(!is.na(.x)))))
ByTrt_22 <- data.frame(t(ByTrt_22))

# Now I want population  means within each treatment
ByTrt_Pop_22 <- Dat_2022 %>%
  dplyr::group_by(Population) %>%
  dplyr::group_by(Treatment, .add = TRUE) %>%
  summarize(across(.col = c(DaysToEmergence:l10_Stomata_density), .fns = list(mean = ~mean(.x, na.rm = TRUE, .group = "keep"), ci_95 = ~conf_int(.x), n = ~sum(!is.na(.x)))))

# get just the sample sizes
SampleSizes_2022 <- dplyr::select(ByTrt_Pop_22, "Treatment", "Population", ends_with("_n"))
save(SampleSizes_2022, file = "data/SampleSizes_2022.robj")

# for some ease of viewing
ByTrt_Pop_22 <- data.frame(t(ByTrt_Pop_22))
colnames(ByTrt_Pop_22) <- c("Belm_C", "Belm_F", "Roda_C", "Roda_F")

# save this to keep means
write.csv(ByTrt_Pop_22,"data/PopMeansByTreatment_2022.csv", row.names = TRUE )
```


## Anovas
Now, we can make some models. Initially there were three models for this analysis.

For traits collected before bolting: trait ~ Treatment * Population + (1|Population:Line) + (1|Treatment:Chamber) + (1|Transplanted)

For traits collected when a single leaf was harvesting: trait ~ Treatment * Population + (1|BoltingToHarvest) + (1|Population:Line) + (1|Treatment:Chamber) + (1|Transplanted)

For root harvesting traits: trait ~ Treatment * Population + (1|BoltingToRootWashing) + (1|Population:Line) + (1|Treatment:Chamber) + (1|Transplanted)

After discussing with Jeff, chamber nested within treatment was changed to a fixed effect because models struggle to fit a variance with few groups. The Bolting to harvest and bolting to root washing variables were also removed because they are continuous but being treated like categorical as a random effect. These two variables did explain some variation and removing sometimes increased and sometimes decreased both effect sizes and p values. 

If included, the transplanting effect should be fixed. There is a population bias where only italian lines were transplanted. 17 plants from 6 genotypes were transplanted. They became 10 current plants and 7 future plants. **However, when included, the transplanting effect sometimes explains 0 variance** Because this is such a small number (17 plants of 127 plants), the transplanting effect was removed from the model. All models were run with and without transplanting as  a fixed effect, but only small changes were seen in 3 traits (see data/ModelResults_InitalModels.xlsx for notes.)

Final model (only 1): trait ~ Treatment * Population + Treatment:Chamber + (1|Population:Line)


```{r}
# set contrasts for models to ensure we are doing the correct type of ANOVA (type-III)
contrasts(Dat_2022$Population) <- contr.sum
contrasts(Dat_2022$Treatment) <- contr.sum

# make a modelling function to streamline analyses. This function runs a mixed model, then plots the residuals, the model summary, and runs an ANOVA

# to avoid needing nesting notation, make a column with a unique value for each block (chamber within treatment). This changes the treatment effect and I decided to go back to the nested structure because it is important to think about the two current blocks as being more similar than the two future blocks rather than thinking just about differences between each of the 4 blocks.
#Dat_2022$block <- as.factor(paste0(Dat_2022$Treatment, Dat_2022$Chamber))

do_lmer2 <- function(trait, data = Dat_2022){
  lm<- lmer(trait ~ Treatment * Population + Treatment/Chamber + (1|Population:Line), data = data, contrasts = list(Treatment=contr.sum, Population = contr.sum))
  print(plot(lm))
  hist(residuals(lm), breaks = 15)
  # in the residual plots, red is italy, blue is sweden; filled circle is current, filled triangle is future
  plot(fitted(lm), residuals(lm, type = "pearson", scaled = TRUE),
       col = c("red", "blue")[as.numeric(Dat_2022$Population)[complete.cases(trait)]],
       pch = c(16, 17)[as.numeric(Dat_2022$Treatment)[complete.cases(trait)]])
  #legend("topleft", legend = c("Italy", "Sweden"), col = c("red", "blue"), pch = 16)
  #legend("bottomleft", legend = c("Current", "Future"), col = "black", pch = c(16,17))
  #plotting sanity check
  print(paste0("the number of complete cases is ", sum(complete.cases(trait))))
  # make qq plot
  qqnorm(resid(lm))
  qqline(resid(lm))
  print(summary(lm))
  print(confint(lm))
  return(lm)
}
# do the anova
do_anov2 <- function(model){
  anov <- anova(model)
  print(anov)
  df <- as.data.frame(anov)
  df$exp <- "2022"
  return(df)
}

# get table functions are the same as from 2021

  
```

```{r, eval = FALSE, echo = FALSE}
# no longer used functions
# include days between bolting and harvesting as random effect
anov_lmer3 <- function(trait, data = Dat_2022){
  lm<- lmer(trait ~ Treatment * Population + (1|BoltingToHarvest) + (1|Population:Line) + (1|Treatment:Chamber) + (1|Transplanted), data = data, contrasts = list(Treatment=contr.sum, Population = contr.sum))
  print(plot(lm))
  hist(residuals(lm), breaks = 15)
  # in the residual plots, red is italy, blue is sweden; filled circle is current, filled triangle is future
  plot(fitted(lm), residuals(lm, type = "pearson", scaled = TRUE),
       col = c("red", "blue")[as.numeric(Dat_2022$Population)],
       pch = c(16, 17)[as.numeric(Dat_2022$Treatment)])
  legend("topleft", legend = c("Italy", "Sweden"), col = c("red", "blue"), pch = 16)
  legend("bottomleft", legend = c("Current", "Future"), col = "black", pch = c(16,17))
  # make qq plot
  qqnorm(resid(lm))
  qqline(resid(lm))
  print(summary(lm))
  print(confint(lm))
  anov <- anova(lm)
  print(anov)
  return(anov)
}

# include days between bolting and root washing as random effect
anov_lmer4 <- function(trait, data = Dat_2022){
  lm<- lmer(trait ~ Treatment * Population + (1|BoltingToRootWashing) + (1|Population:Line) + (1|Treatment:Chamber) + (1|Transplanted), data = data, contrasts = list(Treatment=contr.sum, Population = contr.sum))
  print(plot(lm))
  hist(residuals(lm), breaks = 15)
  # in the residual plots, red is italy, blue is sweden; filled circle is current, filled triangle is future
  plot(fitted(lm), residuals(lm, type = "pearson", scaled = TRUE),
       col = c("red", "blue")[as.numeric(Dat_2022$Population)],
       pch = c(16, 17)[as.numeric(Dat_2022$Treatment)])
  legend("topleft", legend = c("Italy", "Sweden"), col = c("red", "blue"), pch = 16)
  legend("bottomleft", legend = c("Current", "Future"), col = "black", pch = c(16,17))
  # make qq plot
  qqnorm(resid(lm))
  qqline(resid(lm))
  print(summary(lm))
  print(confint(lm))
  anov <- anova(lm)
  print(anov)
  return(anov)
}
```

Run the models
### Phenology

Days to emergence - log10 transformed this time!
no post hoc because not in manuscript
```{r}
# days to emergence
emergence_lm_22 <- do_lmer2(Dat_2022$l10_DaysToEmergence)
emergence_anov_22 <- do_anov2(emergence_lm_22)
emergence_anov_22$trait <- "Emergence"
emergence_means_22 <- as.data.frame(get_table_bt(emergence_lm_22))
# no interaction
```

Days between emergence and bolting
```{r}
# days emergence to bolting
bolting_lm_22 <- do_lmer2(Dat_2022$EmergenceToBolting)
bolting_anov_22 <- do_anov2(bolting_lm_22)
bolting_anov_22$trait <- "Bolting"
bolting_emmeans_22 <- get_table(bolting_lm_22)
bolting_means_22 <- as.data.frame(bolting_emmeans_22)
bolting_pairs_22 <- as.data.frame(pairs(bolting_emmeans_22))
bolting_pairs_22$trait <- "Bolting"
```

 Not analyzing bolting to root washing or bolting to harvest. Those could be included as random effects, but most were close together. maybe should do analysis on if that should matter. should look at summary sheet to see. There do seem to be differences so the root washing traits (biomass) should have days to root wash as a random effect and other bolting day traits should have bolting to harvest as a random effect. might need to redo the model or not use the function to do this.
 
Not looking an number emerged because doesn't have a 2021 comparison for emergence success... ignoring but might want to look at later.

### Single Leaf Traits: bolting

Fresh weight
```{r}
# single leaf fresh weight
fresh_lm_22 <- do_lmer2(Dat_2022$l10_SL_FreshWt)
fresh_anov_22 <- do_anov2(fresh_lm_22)
fresh_anov_22$trait <- "l10_FreshWt"
fresh_emmeans_22 <- get_table_bt(fresh_lm_22)
fresh_means_22 <- as.data.frame(fresh_emmeans_22)
fresh_pairs_22 <- as.data.frame(pairs(fresh_emmeans_22))
fresh_pairs_22$trait <- "l10_FreshWt"
```


Saturated/Hydrated weight
```{r}
# single leaf hydrated (also called saturated) weight
hyd_lm_22 <- do_lmer2(Dat_2022$l10_SL_HydWt)
hyd_anov_22 <- do_anov2(hyd_lm_22)
hyd_anov_22$trait <- "l10_SatWt"
hyd_emmeans_22 <- get_table_bt(hyd_lm_22)
hyd_means_22 <- as.data.frame(hyd_emmeans_22)
hyd_pairs_22 <- as.data.frame(pairs(hyd_emmeans_22))
hyd_pairs_22$trait <- "l10_SatWt"

## also has SQR transformation. Pop becomes more significant but no other significance changes
#hyd_lm_22SQR <- do_lmer2(Dat_2022$SQR_SL_HydWt)
#hyd_anov_22SQR <- do_anov2(hyd_lm_22SQR)
#hyd_anov_22SQR$trait <- "l10_SatWt"
```


Dried weight
```{r}
# single leaf dry weight
dry_lm_22 <- do_lmer2(Dat_2022$SL_DryWt)
dry_anov_22 <- do_anov2(dry_lm_22)
dry_anov_22$trait <- "DriedWt"
dry_emmeans_22 <- get_table(dry_lm_22)
dry_means_22 <- as.data.frame(dry_emmeans_22)
dry_pairs_22 <- as.data.frame(pairs(dry_emmeans_22))
dry_pairs_22$trait <- "DriedWt"
# also has SQR transformation
# no interaction

# what about SQR transformation - no meaningful change
#anov_lmer2(Dat_2022$SQR_SL_DryWt)

# and what about l10 transformation - no meaningful change
#anov_lmer2(log10(Dat_2022$SL_DryWt))
```

leaf area
```{r}
# single leaf area
area_lm_22 <- do_lmer2(Dat_2022$SL_Area)
area_anov_22 <- do_anov2(area_lm_22)
area_anov_22$trait <- "Area"
area_emmeans_22 <- get_table(area_lm_22)
area_means_22 <- as.data.frame(area_emmeans_22)
area_pairs_22 <- as.data.frame(pairs(area_emmeans_22))
area_pairs_22$trait <- "Area"
## also has SQR - no significance changes
#area_lm_22SQR <- do_lmer2(Dat_2022$SQR_SL_Area)
#area_anov_22SQR <- do_anov2(area_lm_22SQR)
#area_anov_22SQR$trait <- "SQR_Area"
```


leaf perimeter
```{r}
# single leaf perimeter
per_lm_22 <- do_lmer2(Dat_2022$SL_Perim)
per_anov_22 <- do_anov2(per_lm_22)
per_anov_22$trait <- "Perimeter"
per_emmeans_22 <- get_table(per_lm_22)
per_means_22 <- as.data.frame(per_emmeans_22)
per_pairs_22 <- as.data.frame(pairs(per_emmeans_22))
per_pairs_22$trait <- "Perimeter"
```


specific leaf area
```{r}
# SLA
sla_lm_22 <- do_lmer2(Dat_2022$l10_SLA)
sla_anov_22 <- do_anov2(sla_lm_22)
sla_anov_22$trait <- "l10_SLA"
sla_emmeans_22 <- get_table_bt(sla_lm_22)
sla_means_22 <- as.data.frame(sla_emmeans_22)
sla_pairs_22 <- as.data.frame(pairs(sla_emmeans_22))
sla_pairs_22$trait <- "l10_SLA"
```

leaf dry matter content
```{r}
# leaf dry matter content
ldmc_lm_22 <- do_lmer2(Dat_2022$l10_LDMC)
ldmc_anov_22 <- do_anov2(ldmc_lm_22)
ldmc_anov_22$trait <- "l10_LDMC"
ldmc_emmeans_22 <- get_table_bt(ldmc_lm_22)
ldmc_means_22 <- as.data.frame(ldmc_emmeans_22)
ldmc_pairs_22 <- as.data.frame(pairs(ldmc_emmeans_22))
ldmc_pairs_22$trait <- "l10_LDMC"
```


relative water content
```{r}
# relative water content
rwc_lm_22 <- do_lmer2(Dat_2022$RWC)
rwc_anov_22 <- do_anov2(rwc_lm_22)
rwc_anov_22$trait <- "RWC"
rwc_emmeans_22 <- get_table(rwc_lm_22)
rwc_means_22 <- as.data.frame(rwc_emmeans_22)
rwc_pairs_22 <- as.data.frame(pairs(rwc_emmeans_22))
rwc_pairs_22$trait <- "RWC"
# tori did log for RWC.
```

### Leaf Number

leaf num at 4 weeks
no post hoc because not in manuscript
```{r}
# pre vern leaf number
LN_PreVern_lm_22 <- do_lmer2(Dat_2022$LeafNumber_PreVern)
LN_PreVern_anov_22 <- do_anov2(LN_PreVern_lm_22)
LN_PreVern_anov_22$trait <- "LeafNum_5wks"
LN_PreVern_means_22 <- as.data.frame(get_table(LN_PreVern_lm_22))
# no interaction
```

leaf num first week post vern
no post hoc because not in manuscript
```{r}
# june 6 leaf number(first week post vern?)
LN_Jun6_lm_22 <- do_lmer2(Dat_2022$LeafNumber_Jun6)
LN_Jun6_anov_22 <- do_anov2(LN_Jun6_lm_22)
LN_Jun6_anov_22$trait <- "LeafNum_9wks"
LN_Jun6_means_22 <- as.data.frame(get_table(LN_Jun6_lm_22))
# no interaction
```

leaf num second week post vern
no post hoc because not in manuscript
```{r}
# june 14 leaf number (second week post vern?)
LN_Jun13_lm_22 <- do_lmer2(Dat_2022$LeafNumber_Jun13)
LN_Jun13_anov_22 <- do_anov2(LN_Jun13_lm_22)
LN_Jun13_anov_22$trait <- "LeafNum_10wks"
LN_Jun13_means_22 <- as.data.frame(get_table(LN_Jun13_lm_22))
# no interaction
```

leaf num at bolting
```{r}
# total leaf number counted at bolting when plants were harvested
LN_bolt_lm_22 <- do_lmer2(Dat_2022$l10_LeafNumber_Total)
LN_bolt_anov_22 <- do_anov2(LN_bolt_lm_22)
LN_bolt_anov_22$trait <- "l10_LeafNumb_bolting"
LN_bolt_emmeans_22 <- get_table_bt(LN_bolt_lm_22)
LN_bolt_means_22<- as.data.frame(LN_bolt_emmeans_22)
LN_bolt_pairs_22 <- as.data.frame(pairs(LN_bolt_emmeans_22))
LN_bolt_pairs_22$trait <- "l10_LeafNumb_bolting"
# no interaction
```

Not analyzing rosette leaf number or under leaves because this split was started partway through the experiment.


### Biomass

above ground biomass
```{r}
# above ground biomass (same as rosette biomass)
AG_lm_22 <- do_lmer2(Dat_2022$l10_AG_DryBiomass)
AG_anov_22 <- do_anov2(AG_lm_22)
AG_anov_22$trait <- "l10_Above Ground Biomass"
AG_emmeans_22 <- get_table_bt(AG_lm_22)
AG_means_22 <- as.data.frame(AG_emmeans_22)
AG_pairs_22 <- as.data.frame(pairs(AG_emmeans_22))
AG_pairs_22$trait <- "l10_Above Ground Biomass"
## also has SQR transformation -interaction gets a little weaker
#AG_lm_22SQR <- do_lmer2(Dat_2022$SQR_AG_DryBiomass)
#AG_anov_22SQR <- do_anov2(AG_lm_22SQR)
#AG_anov_22SQR$trait <- "SQR_Above Ground Biomass"
```

below ground biomass
```{r}
# below ground biomass
BG_lm_22 <- do_lmer2(Dat_2022$BG_DryBiomass)
BG_anov_22 <- do_anov2(BG_lm_22)
BG_anov_22$trait <- "Below Ground Biomass"
BG_emmeans_22 <- get_table(BG_lm_22)
BG_means_22 <- as.data.frame(BG_emmeans_22)
BG_pairs_22 <- as.data.frame(pairs(BG_emmeans_22))
BG_pairs_22$trait <- "Below Ground Biomass"
### also has SQR transformation; non significance changes
#BG_lm_22SQR <- do_lmer2(Dat_2022$SQR_BG_DryBiomass)
#BG_anov_22SQR <- do_anov2(BG_lm_22SQR)
#BG_anov_22SQR$trait <- "SQR_Below Ground Biomass"
# no interaction
```

Root to shoot ration (below ground / above ground)
```{r}
# Root to Shoot ratio
RS_lm_22 <- do_lmer2(Dat_2022$l10_Root_to_Shoot)
RS_anov_22 <- do_anov2(RS_lm_22)
RS_anov_22$trait <- "l10_Root to Shoot Ratio"
RS_emmeans_22 <- get_table_bt(RS_lm_22)
RS_means_22 <- as.data.frame(RS_emmeans_22)
RS_pairs_22 <- as.data.frame(pairs(RS_emmeans_22))
RS_pairs_22$trait <- "l10_Root to Shoot Ratio"
```


### Stomatal Density
Not analyzing stomata average becuase density is the average information with the viewing window area included in the calculation.
```{r}
# stomatal density - imaged on harvest day
sto_den_lm_22 <- do_lmer2(Dat_2022$l10_Stomata_density)
sto_den_anov_22 <- do_anov2(sto_den_lm_22)
sto_den_anov_22$trait <- "l10_Stomatal Density"
sto_den_emmeans_22 <- get_table_bt(sto_den_lm_22)
sto_den_means_22 <- as.data.frame(sto_den_emmeans_22)
sto_den_pairs_22 <- as.data.frame(pairs(sto_den_emmeans_22))
sto_den_pairs_22$trait <- "l10_Stomatal Density"
# no interaction

# check results don't change if using stomatal average
do_anov2(do_lmer2(Dat_2022$l10_Stomata_avg))
# results don't change. might want to switch to this if can't figure out actual area of those photos. Published rates for Arabidopsis are 87-204 stomata/mm2

```

### Outputs

1) results table

```{r}
AnovaResults_2022 <- rbind(emergence_anov_22, bolting_anov_22, fresh_anov_22, hyd_anov_22, dry_anov_22, area_anov_22, per_anov_22, sla_anov_22, ldmc_anov_22, rwc_anov_22, LN_PreVern_anov_22, LN_Jun6_anov_22, LN_Jun13_anov_22, LN_bolt_anov_22, AG_anov_22, BG_anov_22, RS_anov_22, sto_den_anov_22 )

write.csv(AnovaResults_2022, file = "data/AnovaResults_2022.csv", row.names = TRUE)
```

2) All the tables to read in to the figures code

```{r}
# change column names
dfs3 <- c("bolting_means_22", "dry_means_22", "area_means_22", "per_means_22", "rwc_means_22", "LN_PreVern_means_22", "LN_Jun6_means_22", "LN_Jun13_means_22", "BG_means_22", "emergence_means_22", "fresh_means_22", "hyd_means_22", "sla_means_22", "ldmc_means_22", "LN_bolt_means_22", "AG_means_22", "RS_means_22", "sto_den_means_22")

for (i in dfs3) {
  x=get(i)
  colnames(x) <- c("Treatment", "Population", "Mean", "SE", "Df", "LL_95", "UL_95")
  assign(i,x)
}


# create list of data frames
Means_2022 <- list(emergence_means_22, bolting_means_22, fresh_means_22, hyd_means_22, dry_means_22, area_means_22, per_means_22, sla_means_22, ldmc_means_22, rwc_means_22, LN_PreVern_means_22, LN_Jun6_means_22, LN_Jun13_means_22, LN_bolt_means_22, AG_means_22, BG_means_22, RS_means_22, sto_den_means_22)

# name that list
names(Means_2022) <- c("emergence_means_22", "bolting_means_22", "fresh_means_22", "hyd_means_22", "dry_means_22", "area_means_22", "per_means_22", "sla_means_22", "ldmc_means_22", "rwc_means_22", "LN_PreVern_means_22", "LN_Jun6_means_22", "LN_Jun13_means_22", "LN_bolt_means_22", "AG_means_22", "BG_means_22", "RS_means_22", "sto_den_means_22")

# save named list
save(Means_2022, file = "data/ModelMeans_2022.robj")

```

3) Post hoc test results
```{r}
PostHoc_2022 <- rbind(bolting_pairs_22, fresh_pairs_22, hyd_pairs_22, dry_pairs_22, area_pairs_22, per_pairs_22, sla_pairs_22, ldmc_pairs_22, rwc_pairs_22, LN_bolt_pairs_22, AG_pairs_22, BG_pairs_22, RS_pairs_22, sto_den_pairs_22 )

write.csv(PostHoc_2022, file = "data/PostHoc_2022.csv", row.names = TRUE)
```

## Sample size information

The goal here is to know how many lines and how many replicates per line for this experiment.

What I want is a count of total plants, a mean, min, and max of plants per line, and the number of lines per population

```{r}
# sample size of 2022 total
Dat_2022 %>% count(Treatment)
# 72 plants in current, 69 plants in future
Dat_2022 %>% count(Treatment, Population)
# in each treatment, 20 are IT, 30 are SW

# from models know that there are 16 unique lines. 8 IT, 8 SW. ( but 1 line fully died in IT cur)

n_2022 <- Dat_2022 %>%
  group_by(Treatment) %>%
  group_by(Population, .add = TRUE) %>%
  group_by(Line, .add = TRUE) %>%
  count

# so only line with 1 rep is Belm 1. Then everything has 2 but RIL parents. B12 has 7, R47 has 6.
n_2022 %>% 
  group_by(Treatment) %>%
  summarize(min = min(n), mean = mean(n), median = median(n), max = max(n))

n_2022 %>% 
  group_by(Treatment) %>%
  group_by(Population, .add = TRUE) %>%
  summarize(min = min(n), mean = mean(n), median = median(n), max = max(n))

```

#Notes
The AnovaResults_YEAR files that were output have been copied over to the data/ModelResults.xlsx on 8/2/2024 to make formatting changes and color code for easier comparison and analysis. This was erpeated on 7/22/2025 to update with the emmeans package outputs.

The spreadsheet is color coded so p values below 0.05 are green, between 0.1 and 0.05 are orange, and others are left white. The model terms are red if there is disagreement in significance between experiments for traits that were measured in both experiments. Post hoc analyses if there was an interaction between population and treatment are only recorded here (as of 7/19/2024) and are only done for 2021 as of 8/2/2024.More were done after and added to the spreadsheet with a note of if there was an interaction or not. Now with emmeans there is a post hoc results spreadsheet as well which is included in the tables for the manuscript

Choosing to only include SLA and move LDMC to the supplemental so I want correlation information for the two traits.
```{r}
# 2021
# overall
cor.test(Dat_2021_TwoTrt$SLA, Dat_2021_TwoTrt$LDMC)
plot(Dat_2021_TwoTrt$SLA, Dat_2021_TwoTrt$LDMC)

# just in SW
cor.test(Dat_2021_TwoTrt[Dat_2021_TwoTrt$Population == "RODA", ]$SLA, Dat_2021_TwoTrt[Dat_2021_TwoTrt$Population == "RODA", ]$LDMC)
plot(Dat_2021_TwoTrt[Dat_2021_TwoTrt$Population == "RODA", ]$SLA, Dat_2021_TwoTrt[Dat_2021_TwoTrt$Population == "RODA", ]$LDMC)

# just in IT
cor.test(Dat_2021_TwoTrt[Dat_2021_TwoTrt$Population == "BELM", ]$SLA, Dat_2021_TwoTrt[Dat_2021_TwoTrt$Population == "BELM", ]$LDMC)
plot(Dat_2021_TwoTrt[Dat_2021_TwoTrt$Population == "BELM", ]$SLA, Dat_2021_TwoTrt[Dat_2021_TwoTrt$Population == "BELM", ]$LDMC)

# just cur?
cor.test(Dat_2021_TwoTrt[Dat_2021_TwoTrt$Treatment == "Current", ]$SLA, Dat_2021_TwoTrt[Dat_2021_TwoTrt$Treatment == "Current", ]$LDMC)
plot(Dat_2021_TwoTrt[Dat_2021_TwoTrt$Treatment == "Current", ]$SLA, Dat_2021_TwoTrt[Dat_2021_TwoTrt$Treatment == "Current", ]$LDMC)

# just fut?
cor.test(Dat_2021_TwoTrt[Dat_2021_TwoTrt$Treatment == "Future", ]$SLA, Dat_2021_TwoTrt[Dat_2021_TwoTrt$Treatment == "Future", ]$LDMC)
plot(Dat_2021_TwoTrt[Dat_2021_TwoTrt$Treatment == "Future", ]$SLA, Dat_2021_TwoTrt[Dat_2021_TwoTrt$Treatment == "Future", ]$LDMC)

# 2022
# overall
cor.test(Dat_2022$SLA, Dat_2022$LDMC)
plot(Dat_2022$SLA, Dat_2022$LDMC)

# just in SW
cor.test(Dat_2022[Dat_2022$Population == "R", ]$SLA, Dat_2022[Dat_2022$Population == "R", ]$LDMC)
plot(Dat_2022[Dat_2022$Population == "R", ]$SLA, Dat_2022[Dat_2022$Population == "R", ]$LDMC)

# just in IT
cor.test(Dat_2022[Dat_2022$Population == "B", ]$SLA, Dat_2022[Dat_2022$Population == "B", ]$LDMC)
plot(Dat_2022[Dat_2022$Population == "B", ]$SLA, Dat_2022[Dat_2022$Population == "B", ]$LDMC)

# just cur?
cor.test(Dat_2022[Dat_2022$Treatment == "Current", ]$SLA, Dat_2022[Dat_2022$Treatment == "Current", ]$LDMC)
plot(Dat_2022[Dat_2022$Treatment == "Current", ]$SLA, Dat_2022[Dat_2022$Treatment == "Current", ]$LDMC)

# just fut?
cor.test(Dat_2022[Dat_2022$Treatment == "Current", ]$SLA, Dat_2022[Dat_2022$Treatment == "Future", ]$LDMC)
plot(Dat_2022[Dat_2022$Treatment == "Future", ]$SLA, Dat_2022[Dat_2022$Treatment == "Current", ]$LDMC)
```


# Results Tables

## Main Text Tables - auto formatted options
```{r}
sjPlot::tab_model(eTof_lm_21, RWC_lm_21, rwc_lm_22, sto_den_lm_22, SLA_lm_21, sla_lm_22, RS_lm_22 ,  fitness_lm_21, 
                   collapse.ci = TRUE, 
                  dv.labels = c("Emergence To Flower 21", "RWC 21", "RWC 22", "Stomatal Density 22","SLA 21", "SLA 22",  "Root-to-Shoot 22", "Fitness 21"), 
                 pred.labels = c("(Intercept)", "Treatment", "Population", "Trt x Pop Interaction", "Current Trt x Chamber", "Future Trt x Chamber")
                 
        )

#show.reflvl = TRUE 
#p.style = "stars" or "scientific"
#digits.p=
# I think I can use css code from this tutorial to remove the intercept row if I want?? https://strengejacke.github.io/sjPlot/articles/table_css.html
```

## Supplemental Tables
```{r}

drought_extra <- sjPlot::tab_model(fresh_lm_21, fresh_lm_22, sat_lm_21, hyd_lm_22, dry_lm_21, dry_lm_22, area_lm_21, area_lm_22, per_lm_21, per_lm_22, LDMC_lm_21, ldmc_lm_22,
                                   collapse.ci = TRUE, 
                                   dv.labels = c("Fresh 21", "Fresh 22", "Sat 21", "Sat 22", "Dry 21", "Dry 22", "area 21", "area 22", "perimeter 21", "perimeter 22", "LDMC 21", "LDMC 22")
                                   )
growth_extra <- sjPlot::tab_model(emergence_lm_21, emergence_lm_22, bolting_lm_21, bolting_lm_22, Ros_lm_21, AG_lm_22, RR_lm_21, Repro_lm_21, AG_lm_21, LN_Vern_lm_21, LN_PreVern_lm_21, LN_PreVern_lm_22, LN_PostVern_lm_21, LN_Jun6_lm_22, LN_Jun13_lm_22, LN_harv_lm_21, LN_bolt_lm_22, BG_lm_22, AvSeedWt_lm_21,
                                  collapse.ci = TRUE, 
                                  dv.labels = c("Emergence 21", "Emergence 22", "Bolting 21", "Bolting 22", "Ros Biomass 21", "Ros Biomass 22", "Ros:Repro 21", "Repro 21", "AG 21", "LN Vern 21", "LN Prevern 21", "Ln Prevern 22", "LN Postvern 21", "LN June 6 22", "LN June 13 22", "LN harvest 21", "Ln bolt 22", "BG 22", "Avg. Seed Wt 21"))
fitness_extra <- sjPlot::tab_model(seeds_lm_21, PrimStalks_lm_21, LatBranch_lm_21, height_lm_21, fruit_lm_21,
                                   collapse.ci = TRUE,
                                   dv.labels = c("SeedsPerFruit", "Primary Stalks", "Lateral Branches", "Height", "Fruit Number"))

#flowering_lm_21 vs eTof_lm_21??
 
                                                    
```

# Problem Solve
## Dry weight difference in signficance
For dry weight, there is a huge difference in significance between 2021 (p value below 0.001) and 2022 (p value above 0.95). Quickly looking at the spreadsheets I can not tell a difference and the values are of similar magnitudes. I then did untransformed models and square root transformed models because those were reported as having the most normal residuals in the previous code. None of the models had differences in p values due to transformation that changed the interpretation of the results. I also excluded the "days between bolting and harvest" random effect for the 2022 data but did not find a major difference in the models but the denominator degrees of freedom for treatment did change a lot. As a final check for if there is a potential mistake, I am going to look at some data distributions.

```{r}
ggplot(dat = Dat_2021_TwoTrt)+
  geom_histogram(aes(x = DriedWt_g, fill = Population), alpha = 0.5, position = "identity")+
  theme_classic()
ggplot(dat = Dat_2021_TwoTrt)+
  geom_histogram(aes(x = DriedWt_g, fill = Treatment), alpha = 0.5, position = "identity")+
  theme_classic()


ggplot(dat = Dat_2022)+
  geom_histogram(aes(x = SL_DryWt, fill = Population), alpha = 0.5, position = "identity")+
  theme_classic()
ggplot(dat = Dat_2022)+
  geom_histogram(aes(x = SL_DryWt, fill = Treatment), alpha = 0.5, position = "identity")+
  theme_classic()
```

Visually looking at the histograms it does seems like 2021 has a treatment difference and 2022 has a population difference.Scales of both are comparable. 

## 2022 B1 to B12

I am trying to figure out if I need to adjust these things or not, so I want to look at how different their trait values are. so first let's subset down to just these from Dat_2022 and then we can summarize by line to see if the means are like identical or not identical. Could also make figures if I think it is needed. 
```{r}
B_comp <- Dat_2022[Dat_2022$Population == "B" & Dat_2022$Line %in% c(1,12), ]

B_comp <- B_comp %>%
  dplyr::group_by(Line) %>%
  dplyr::group_by(Treatment, .add = TRUE) %>%
  summarize(across(.col = c(DaysToEmergence:Stomata_density), .fns = list(mean = ~mean(.x, na.rm = TRUE, .group = "keep"))))

# for some ease of viewing
B_comp <- data.frame(t(B_comp))
```

They honestly don't seem to be close enough to the same for me to think they are the same genotype. So I am going to leave them as labelled because I don't know for sure that I planted stock center seeds instead of going of the labelled tubes. The only indication to me that B1 might be stock center seeds is the higher number of seeds that emerged, but in the later growing seasons B1 also had good emergence (1 and 3 in GS2, 7 and 7 in GS3). Also, since I fixed it immediately once found for the pilot results, I think I would have also fixed it immediately for Tori's results since we were still working together when I found the mistake.

## 2021 Avg Seed per fruit
Concerned that line does not explain any variance. Is it true that all the variation is within lines and not between lines? to check this: 1) make histogram of all the values (not line means) - done in CleanData but repeat here

```{r}
hist(Dat_2021_TwoTrt$AvgSeedNum)
hist(Dat_2021_TwoTrt$l10_AvgSeedNum)
```
2) look within lines for difference in avg seeds/fruit - maybe do subtraction and histogram the differences? but won't work if more than two values

```{r}
# start with copy of code from 03_Figures that makes the by line plot
Dat_2021_TwoTrt$Line.ID <- as.factor(paste0(Dat_2021_TwoTrt$Population, Dat_2021_TwoTrt$Line))
Dat_2021_TwoTrt$Line.ID.T <- as.factor(paste0(Dat_2021_TwoTrt$Line.ID, Dat_2021_TwoTrt$Treatment))
df_seed_line <- Dat_2021_TwoTrt %>%
  dplyr::group_by(Line.ID.T, .add = TRUE) %>%
  summarize(across(.col = c(AvgSeedNum, l10_AvgSeedNum), .fns = list(mean = ~mean(.x, na.rm = TRUE, .group = "keep"), ci_95 = ~conf_int(.x), n = ~sum(!is.na(.)))))
df_seed_line$Population <- substr(df_seed_line$Line.ID.T, start = 1, stop = 4)

hist(df_seed_line$AvgSeedNum_n)
# some lines only have 1 pot per treatment -- 10 instances, 7 in Cur
# some lines only have pots in one treatment 4 - all in Fut are the dead ones so only a rep in cur
# 3 and 4 are Roda parent. 6 and 7 are Belm parent.

# get list of Line.ID with 2 replicates
two_reps <- df_seed_line[df_seed_line$AvgSeedNum_n == 2, "Line.ID.T"]$Line.ID.T
# pull those out of the main dataframe
two_reps_seeds <- Dat_2021_TwoTrt[Dat_2021_TwoTrt$Line.ID.T %in% two_reps, c("Treatment", "Population", "Line", "Replicate", "Line.ID", "Line.ID.T", "AvgSeedNum", "l10_AvgSeedNum")]

# find the difference in values
two_reps_diff <-  two_reps_seeds %>% 
  group_by(Line.ID.T) %>% 
  summarize(across(.col = c(AvgSeedNum, l10_AvgSeedNum), .fns = ~diff(.x)))


# table of just RIL parent seed numbers
ril_pars_seed <- Dat_2021_TwoTrt[Dat_2021_TwoTrt$Line.ID %in% c("BELM12", "RODA47"), c("Treatment", "Population", "Line", "Replicate", "Line.ID", "Line.ID.T", "AvgSeedNum", "l10_AvgSeedNum")]
ril_pars_seed[order(ril_pars_seed$Line.ID.T), ]
```


3) figure out which genotypes are the outliers on the residuals plot.The low outlier on the residuals plot is B12. Note for interpretation that the residuals are scales (so not log scale but divided by the mean)
```{r}
# test out ggplotly
tmp_seed <- lmer(l10_AvgSeedNum ~ Treatment * Population + (1|Population:Line) , data = Dat_2021_TwoTrt, contrasts = list(Treatment=contr.sum, Population = contr.sum))

df_seed <- data.frame(fitted = fitted(tmp_seed),
                      resid = residuals(tmp_seed, type = "pearson", scaled = TRUE),
                      pop = Dat_2021_TwoTrt$Population[complete.cases(Dat_2021_TwoTrt$l10_AvgSeedNum)],
                      trt = Dat_2021_TwoTrt$Treatment[complete.cases(Dat_2021_TwoTrt$l10_AvgSeedNum)],
                      line = Dat_2021_TwoTrt$Line[complete.cases(Dat_2021_TwoTrt$l10_AvgSeedNum)],
                      pot.id = Dat_2021_TwoTrt$Pot.ID[complete.cases(Dat_2021_TwoTrt$l10_AvgSeedNum)])

plot_seed <-  ggplot(df_seed) +
  geom_point(aes(x=fitted, y=resid,
       col = pop, shape = trt, text=pot.id))+
  scale_color_manual(values = c("red", "blue"))+
  theme_classic()
ggplotly(plot_seed)
```

make a plot with genotype on the x axis and seed count per fruit on the y axis. color by pop and treatment.

```{r, echo = F}
ps_plot <- ggplot(Dat_2021_TwoTrt)+
  geom_point(aes(x = Line.ID.T, y = AvgSeedNum, col = Population, shape = Treatment, text = Pot.ID))+
  scale_color_manual(values = c("red", "blue"))+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
ggplotly(ps_plot)

#ggplot(Dat_2021_TwoTrt)+
#  geom_point(aes(x = Line.ID, y = AvgSeedNum, col = Population, shape = Treatment))+
#  scale_color_manual(values = c("red", "blue"))+
#  theme_classic()+
#  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
#
#ggplot(Dat_2021_TwoTrt)+
#  geom_point(aes(x = Pot.ID, y = AvgSeedNum, col = Population, shape = Treatment))+
#  scale_color_manual(values = c("red", "blue"))+
#  theme_classic()+
#  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```


Compare that to a plot of a trait that does have a genotype effect (just the first plot)
```{r, echo = FALSE}
ps_plot2 <- ggplot(Dat_2021_TwoTrt)+
  geom_point(aes(x = Line.ID.T, y = IJ_FruitCount, col = Population, shape = Treatment, text = Pot.ID))+
  scale_color_manual(values = c("red", "blue"))+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
ggplotly(ps_plot2)

ps_plot3 <- ggplot(Dat_2021_TwoTrt)+
  geom_point(aes(x = Line.ID.T, y = DayToBolt, col = Population, shape = Treatment, text = Pot.ID))+
  scale_color_manual(values = c("red", "blue"))+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
ggplotly(ps_plot3)

ps_plot4 <- ggplot(Dat_2021_TwoTrt)+
  geom_point(aes(x = Line.ID.T, y = LeafPerimeter, col = Population, shape = Treatment, text = Pot.ID))+
  scale_color_manual(values = c("red", "blue"))+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
ggplotly(ps_plot4)

```

## Why are some confidence intervals so small?

SLA - focus on 2022 SW

Plot the histograms for everything, then add the means and confidence intervals on top of the distributions
```{r, eval = F}
ggplot(dat = Dat_2021_TwoTrt)+
  geom_histogram(aes(x = SLA, fill = Treatment:Population), alpha = 0.5, position = "identity", binwidth = 25)+
  geom_vline(data = SLA_means_21, aes(xintercept = Bk_Mean, color = Treatment:Population),  linetype = "solid")+
  geom_vline(data = SLA_means_21, aes(xintercept = Bk_LL_95, color = Treatment:Population),  linetype = "dashed")+
  geom_vline(data = SLA_means_21, aes(xintercept = Bk_UL_95, color = Treatment:Population),  linetype = "dashed")+
  labs(title = "2021 SLA Back Transformed")+
  facet_grid(Treatment ~ Population)+
  theme_classic()

ggplot(dat = Dat_2022)+
  geom_histogram(aes(x = SLA, fill = Treatment:Population), alpha = 0.5, position = "identity", binwidth = 25)+
  geom_vline(data = sla_means_22, aes(xintercept = Bk_Mean, color = Treatment:Population),  linetype = "solid")+
  geom_vline(data = sla_means_22, aes(xintercept = Bk_LL_95, color = Treatment:Population),  linetype = "dashed")+
  geom_vline(data = sla_means_22, aes(xintercept = Bk_UL_95, color = Treatment:Population),  linetype = "dashed")+
  labs(title = "2022 SLA Back Transformed")+
  facet_grid(Treatment ~ Population)+
  theme_classic()


# transformed

ggplot(dat = Dat_2021_TwoTrt)+
  geom_histogram(aes(x = l10_SLA, fill = Treatment:Population), alpha = 0.5, position = "identity", binwidth = .025)+
  geom_vline(data = SLA_means_21, aes(xintercept = Mean, color = Treatment:Population),  linetype = "solid")+
  geom_vline(data = SLA_means_21, aes(xintercept = LL_95, color = Treatment:Population),  linetype = "dashed")+
  geom_vline(data = SLA_means_21, aes(xintercept = UL_95, color = Treatment:Population),  linetype = "dashed")+
  labs(title = "2021 SLA")+
  facet_grid(Treatment ~ Population)+
  theme_classic()

ggplot(dat = Dat_2022)+
  geom_histogram(aes(x = l10_SLA, fill = Treatment:Population), alpha = 0.5, position = "identity", binwidth = .025)+
  geom_vline(data = sla_means_22, aes(xintercept = Mean, color = Treatment:Population),  linetype = "solid")+
  geom_vline(data = sla_means_22, aes(xintercept = LL_95, color = Treatment:Population),  linetype = "dashed")+
  geom_vline(data = sla_means_22, aes(xintercept = UL_95, color = Treatment:Population),  linetype = "dashed")+
  labs(title = "2022 SLA")+
  facet_grid(Treatment ~ Population)+
  theme_classic()
```


```{r eval = F}
ggplot(dat = Dat_2021_TwoTrt)+
  geom_histogram(aes(x = LDMC, fill = Treatment:Population), alpha = 0.5, position = "identity", binwidth = 0.01)+
  geom_vline(data = LDMC_means_21, aes(xintercept = Bk_Mean, color = Treatment:Population),  linetype = "solid")+
  geom_vline(data = LDMC_means_21, aes(xintercept = Bk_LL_95, color = Treatment:Population),  linetype = "dashed")+
  geom_vline(data = LDMC_means_21, aes(xintercept = Bk_UL_95, color = Treatment:Population),  linetype = "dashed")+
  labs(title = "2021 LDMC Back Transformed")+
  facet_grid(Treatment ~ Population)+
  theme_classic()


ggplot(dat = Dat_2022)+
  geom_histogram(aes(x = LDMC, fill = Treatment:Population), alpha = 0.5, position = "identity", binwidth = 0.015)+
  geom_vline(data = ldmc_means_22, aes(xintercept = Bk_Mean, color = Treatment:Population),  linetype = "solid")+
  geom_vline(data = ldmc_means_22, aes(xintercept = Bk_LL_95, color = Treatment:Population),  linetype = "dashed")+
  geom_vline(data = ldmc_means_22, aes(xintercept = Bk_UL_95, color = Treatment:Population),  linetype = "dashed")+
  labs(title = "2022 LDMC Back Transformed")+
  facet_grid(Treatment ~ Population)+
  theme_classic()

# transformed scale
ggplot(dat = Dat_2021_TwoTrt)+
  geom_histogram(aes(x = l10_LDMC, fill = Treatment:Population), alpha = 0.5, position = "identity", binwidth = 0.05)+
  geom_vline(data = LDMC_means_21, aes(xintercept = Mean, color = Treatment:Population),  linetype = "solid")+
  geom_vline(data = LDMC_means_21, aes(xintercept = LL_95, color = Treatment:Population),  linetype = "dashed")+
  geom_vline(data = LDMC_means_21, aes(xintercept = UL_95, color = Treatment:Population),  linetype = "dashed")+
  labs(title = "2021 l10 LDMC")+
  facet_grid(Treatment ~ Population)+
  theme_classic()


ggplot(dat = Dat_2022)+
  geom_histogram(aes(x = l10_LDMC, fill = Treatment:Population), alpha = 0.5, position = "identity", binwidth = 0.05)+
  geom_vline(data = ldmc_means_22, aes(xintercept = Mean, color = Treatment:Population),  linetype = "solid")+
  geom_vline(data = ldmc_means_22, aes(xintercept = LL_95, color = Treatment:Population),  linetype = "dashed")+
  geom_vline(data = ldmc_means_22, aes(xintercept = UL_95, color = Treatment:Population),  linetype = "dashed")+
  labs(title = "2022 l10 LDMC")+
  facet_grid(Treatment ~ Population)+
  theme_classic()
```


## Test: Predicted means
Test new plotting function with predictmeans package. Use emergence to flowering trait to test things out. This works and gives an output of values, including the mean_table which I think is what I want to plot with. I don't want their premade plots becuase I want more customization. The results match the tables from emmeans for 2022 but I think the uneven sample sizes are not dealt with correctly because the SEs don't match emmeans for 2021. I then changed to using emmeans package in July 2025.

```{r eval = F}
# testing plotting functions
# run lm
lm <- lmer(EmergeToFlwr ~ Treatment * Population + (1|Population:Line) , data = Dat_2021_TwoTrt, contrasts = list(Treatment=contr.sum, Population = contr.sum))

predictmeans(lm, modelterm="Treatment:Population",  plotord = c(1,2), lineplot = TRUE, plotxlab = "Treatment", plotylab = "Days between Emergence and Flowering", mplot = TRUE, pplot = FALSE, bkplot = FALSE, plot = TRUE, jitterv = 0.05, prtplt = TRUE, newwd= FALSE)

```


# test selection gradient?
Do I want this? discuss with Emily and Jeff
```{r}
# create subset for each pop x trt combination (4 subsets) - only keeping traits related to avoidance/escape that are not too highly correlated and keeps the number of predictors decently low

# sweden - cur has 27 plants, fut has 25
sw_cur <- Dat_2021_TwoTrt[Dat_2021_TwoTrt$Population == "RODA" & Dat_2021_TwoTrt$Treatment == "Current", c("Pot.ID", "Population", "Line", "Replicate", "Treatment", "RWC", "SLA", "EmergeToFlwr", "fitness")]

sw_fut <- Dat_2021_TwoTrt[Dat_2021_TwoTrt$Population == "RODA" & Dat_2021_TwoTrt$Treatment == "Future", c("Pot.ID", "Population", "Line", "Replicate", "Treatment", "RWC", "SLA", "EmergeToFlwr", "fitness")]

# italy - cur has 18 plants, fut has 15 plants
it_cur <- Dat_2021_TwoTrt[Dat_2021_TwoTrt$Population == "BELM" & Dat_2021_TwoTrt$Treatment == "Current", c("Pot.ID", "Population", "Line", "Replicate", "Treatment", "RWC", "SLA", "EmergeToFlwr", "fitness")]

it_fut <- Dat_2021_TwoTrt[Dat_2021_TwoTrt$Population == "BELM" & Dat_2021_TwoTrt$Treatment == "Future", c("Pot.ID", "Population", "Line", "Replicate", "Treatment", "RWC", "SLA", "EmergeToFlwr", "fitness")]

# need to calculate relative fitness and scale the variables
sw_cur <- sw_cur %>%
  mutate(rel.fit = fitness / max(fitness, na.rm = T),
         RWC = c(scale(RWC, center = T, scale = T)),
         SLA = c(scale(SLA, center = T, scale = T)),
         EmergeToFlwr = c(scale(EmergeToFlwr, center = T, scale = T))
  )

sw_fut <- sw_fut %>%
  mutate(rel.fit = fitness / max(fitness, na.rm = T),
         RWC = c(scale(RWC, center = T, scale = T)),
         SLA = c(scale(SLA, center = T, scale = T)),
         EmergeToFlwr = c(scale(EmergeToFlwr, center = T, scale = T))
  )

it_cur <- it_cur %>%
  mutate(rel.fit = fitness / max(fitness, na.rm = T),
         RWC = c(scale(RWC, center = T, scale = T)),
         SLA = c(scale(SLA, center = T, scale = T)),
         EmergeToFlwr = c(scale(EmergeToFlwr, center = T, scale = T))
  )

it_fut <- it_fut %>%
  mutate(rel.fit = fitness / max(fitness, na.rm = T),
         RWC = c(scale(RWC, center = T, scale = T)),
         SLA = c(scale(SLA, center = T, scale = T)),
         EmergeToFlwr = c(scale(EmergeToFlwr, center = T, scale = T))
  )

# run the selection gradients
sg_sw_cur <- lmer(rel.fit ~ RWC + SLA + EmergeToFlwr + (1|Line), data = sw_cur)
summary(sg_sw_cur) # nothing significant

sg_sw_fut <- lmer(rel.fit ~ RWC + SLA + EmergeToFlwr + (1|Line), data = sw_fut)
summary(sg_sw_fut) # actually does show selection for earlier flowering

sg_it_cur <- lmer(rel.fit ~ RWC + SLA + EmergeToFlwr + (1|Line), data = it_cur)
summary(sg_it_cur) # nothing significant

sg_it_fut <- lmer(rel.fit ~ RWC + SLA + EmergeToFlwr + (1|Line), data = it_fut)
summary(sg_it_fut) # nothing significant

```

        
