---
title: "02_Analysis"
author: "Sophie Buysse"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Description

This code runs all linear models for the 2021 and 2022 experiments. To get this to knit with the update for revisions on 3/18/2026, a bunch of traits that aren't in the manuscript and the problem solving code needed to be set to eval = FALSE for a successful knit to happen. Should come back and fix all that later on.

# Set basic info for all code

This chunk sets things that will be constant for the whole document.

```{r, warning = FALSE, message = FALSE}
# read in packages
library(dplyr)
library(ggplot2)
library(lme4)
library(lmerTest)
library(tidyr)
library(rcompanion)
library(plotly)
library(emmeans)
library(sjPlot)
# prep for doing statistical analyses
options(contrasts = c("contr.sum", "contr.poly"))

# let's make some functions
std_mean <- function(x){
  sd(x, na.rm = TRUE)/sqrt(length(x))
}

# Confidence interval multiplier for standard error
    # Calculate t-statistic for confidence interval: 
    # e.g., if conf.interval is .95, use .975 (above/below), and use df=N-1

conf_int <- function(x, conf.interval = 0.95){
  ciMult <- qt((1+conf.interval)/2, length(x)-1)
  ci <- std_mean(x) * ciMult
  return(ci)
}

```
# 2021

## Read in data

```{r sum1, eval = TRUE}
Dat_2021 <- read.csv("data/CleanData_2021.csv")
# set factors
Dat_2021 <- Dat_2021 %>% dplyr::mutate(
  Population = as.factor(Population),
  Line = as.factor(Line),
  Treatment = as.factor(Treatment),
  Soil.Mix.Batch.Number.x = as.factor(Soil.Mix.Batch.Number.x),
  BranchStructure = as.factor(BranchStructure),
  DoneFlwr = as.factor(DoneFlwr),
  FruitCounter = as.factor(FruitCounter),
  SeedCounter = as.factor(SeedCounter)
  
)

Dat_2021$Geno_ID <- as.factor(paste0(Dat_2021$Population, Dat_2021$Line))
```


## Summarize
```{r}
# NOTE: These are raw means and confidence intervals, used to get a quick look at the data.
ByPop_21 <- Dat_2021 %>%
  dplyr::group_by(Population) %>%
  summarize(across(.col = c(Emergence:LeafNum_08262021, RosetteLeafNum_10042021:Repro_to_Ros, LatBranches:IJ_FruitCount, FruitCollected, NumFlwrLeft:IJ_SeedCount, FruitCount_BL, AG_biomass, AvgSeedWt, AvgSeedNum, fitness), .fns = list(mean = ~mean(.x, na.rm = TRUE, .group = "keep"), ci_95 = ~conf_int(.x), n = ~sum(!is.na(.x)))))

# transpose for easier viewing
ByPop_21 <- data.frame(t(ByPop_21))

# by treatment
ByTrt_21 <- Dat_2021 %>%
  dplyr::group_by(Treatment) %>%
  summarize(across(.col = c(Emergence:LeafNum_08262021, RosetteLeafNum_10042021:Repro_to_Ros, LatBranches:IJ_FruitCount, FruitCollected, NumFlwrLeft:IJ_SeedCount, FruitCount_BL, AG_biomass, AvgSeedWt, AvgSeedNum, fitness), .fns = list(mean = ~mean(.x, na.rm = TRUE, .group = "keep"), ci_95 = ~conf_int(.x), n = ~sum(!is.na(.x)))))
ByTrt_21 <- data.frame(t(ByTrt_21))

#population means within each treatment
ByTrt_Pop_21 <- Dat_2021 %>%
  dplyr::group_by(Population) %>%
  dplyr::group_by(Treatment, .add = TRUE) %>%
  summarize(across(.col = c(Emergence:LeafNum_08262021, RosetteLeafNum_10042021:Repro_to_Ros, LatBranches:IJ_FruitCount, FruitCollected, NumFlwrLeft:IJ_SeedCount, FruitCount_BL, AG_biomass, AvgSeedWt, AvgSeedNum, fitness), .fns = list(mean = ~mean(.x, na.rm = TRUE, .group = "keep"), ci_95 = ~conf_int(.x), n = ~sum(!is.na(.x)))))

# get sample sizes
SampleSizes_2021 <- dplyr::select(ByTrt_Pop_21, "Treatment", "Population", ends_with("_n"))
save(SampleSizes_2021, file = "data/SampleSizes_2021.robj")

# for some ease of viewing
ByTrt_Pop_21 <- data.frame(t(ByTrt_Pop_21))
colnames(ByTrt_Pop_21) <- c("Belm_C", "Belm_F", "Roda_C", "Roda_F")

# save this to keep raw means
write.csv(ByTrt_Pop_21,"data/PopMeansByTreatment_2021.csv", row.names = TRUE )

```
subset to traits in the manuscript
```{r}
Dat_2021_TwoTrt <- Dat_2021[ , c("Pot.ID", "Population", "Line", "Geno_ID", "Replicate", "Treatment", "Soil.Mix.Batch.Number.x", "EmergeToFlwr", "l10_FreshWt", "l10_SatWt", "l10_DriedWt", "l10_LeafArea", "l10_SLA", "l10_LDMC", "RWC", "RosetteLeafNum", "l10_DryRosG", "l10_DryReproG", "l10_R_to_R", "FruitCount_BL", "AvgSeedWt", "AvgSeedNum", "l10_AvgSeedNum", "fitness", "SLA", "LDMC")]

# need to make subsets so only genotypes represented in both treatments are present. this is difficult because it differs between traits.

# need four dataframes. one with everything for fitness measurements of FruitCount_BL and fitness
Dat_2021_fitness <- Dat_2021_TwoTrt[ , c("Pot.ID", "Population", "Line", "Geno_ID", "Replicate", "Treatment", "Soil.Mix.Batch.Number.x","FruitCount_BL", "fitness")]

#one that removes B4, R5, and B13 from all traits left. this will be the starting point for the final two dataframes.
Dat_2021_most <- Dat_2021_TwoTrt[!((Dat_2021_TwoTrt$Population == "BELM" & Dat_2021_TwoTrt$Line %in% c(4, 13)) | (Dat_2021_TwoTrt$Population == "RODA" & Dat_2021_TwoTrt$Line == 5)), c("Pot.ID", "Population", "Line", "Geno_ID", "Replicate", "Treatment", "Soil.Mix.Batch.Number.x", "EmergeToFlwr", "l10_FreshWt", "l10_SatWt", "l10_DriedWt", "l10_LeafArea", "l10_SLA", "l10_LDMC", "RWC", "RosetteLeafNum", "l10_DryRosG", "l10_DryReproG", "l10_R_to_R",  "AvgSeedWt", "AvgSeedNum", "l10_AvgSeedNum", "SLA", "LDMC")]

# need to remove Belm 1 from leaf area and SLA because no leaf area
Dat_2021_la <- Dat_2021_most[!((Dat_2021_most$Population == "BELM" & Dat_2021_most$Line == 1)), c("Pot.ID", "Population", "Line", "Geno_ID", "Replicate", "Treatment", "Soil.Mix.Batch.Number.x", "l10_LeafArea", "l10_SLA", "l10_LDMC", "SLA", "LDMC")]

# need to remove Roda 26 from rosette biomass, repro ros, seed weight and seed number because of labeling error
Dat_2021_RosSeed <- Dat_2021_most[!((Dat_2021_most$Population == "RODA" & Dat_2021_most$Line == 26)), c("Pot.ID", "Population", "Line", "Geno_ID", "Replicate", "Treatment", "Soil.Mix.Batch.Number.x", "l10_DryRosG", "l10_R_to_R",  "AvgSeedWt", "l10_AvgSeedNum", "AvgSeedNum")]
```


## Two Treatment Anovas

The initial model is was follows:
trait ~ Treatment * Population + (1|Population:Line) + (1|Soil.Mix.Batch.Number.x)

When removing the soil mix random effect, most p values and effect sizes were not meaningfully changed with the following 4 exceptions written as to what happened when soil  mix was removed from the model:\
- dry weight of single leaf: population effect increased, p value decreased \
- relative water content: interaction effect increased, p value decreased (makes sense because dry weight changed a little)\
- rosette dry weight: p value cut in half and effect size doubles (did not write down which predictor this was for) \
- fruit number: population p value increases, interaction p value decreases \

In general these model changes were slight and soil mix explained very little variance - the exceptions are relative water content (soil mix explained half as much variance as genotype), number of fruits, and number of primary stalks (soil mix explained more variance than genotype but over half the variance is still residual). Thus, soil mix was removed as a random variable from the analysis.

```{r stats}
# random variables: Line, soil mix batch number
# for single leaf traits also include leaf_Collected
# for traits at harvest: DoneFlwr

# function to run models on 2021 data:
do_lmer <- function(trait, data = Dat_2021_most){
  lm <- lmer(trait ~ Treatment * Population + (1|Population:Line) , data = data, contrasts = list(Treatment=contr.sum, Population = contr.sum))
  print(plot(lm))
  hist(residuals(lm), breaks = 15)
  # plotting sanity check
  print(paste0("the number of complete cases is ", sum(complete.cases(trait))))
  qqnorm(resid(lm))
  qqline(resid(lm))
  print(summary(lm))
  print(confint(lm))
  return(lm)
}
# do the anova
do_anov <- function(model){
  anov <- anova(model)
  print(anov)
  df <- as.data.frame(anov)
  df$exp <- "2021"
  return(df)
}

# get the means table, no backtransforming
get_table <- function(model){
  tab <- emmeans(model, c("Treatment", "Population"), type = "response")
  return(tab)
}
#for backtransformed
get_table_bt <- function(model){
  tab <- get_table(update(ref_grid(model), tran = "log10"))
  return(tab)
}

# genotype level emmeans for looking at within population variation
get_geno_table <- function (trait, data = Dat_2021_most){
  tab2 <- emmeans(lm(trait ~ Treatment * Geno_ID, data = data, contrasts = list(Treatment=contr.sum, Geno_ID = contr.sum)), c("Treatment", "Geno_ID"), type = "reponse")
  return(tab2)
}

get_geno_table_bt <- function(trait, data = Dat_2021_most){
  mod <- lm(trait ~ Treatment * Geno_ID, data = data, contrasts = list(Treatment=contr.sum, Geno_ID = contr.sum))
  tab2 <- emmeans(update(ref_grid(mod), tran = "log10"), c("Treatment", "Geno_ID"), type = "response") #
  return(tab2)
}
```

The modelling function runs a mixed model where each trait is predicted by population, treatment, and their interaction while accounting for random effects of line nested within population and soil mix batch number -> and just line in the final model.

### Phenology

Days between Emergence and flowering

```{r}
# Emergence to flowering
eTof_lm_21 <- do_lmer(Dat_2021_most$EmergeToFlwr)
eTof_anov_21 <- do_anov(eTof_lm_21)
eTof_anov_21$trait <- "Emergence To Flowering"
eTof_emmeans_21 <- get_table(eTof_lm_21)
eTof_means_21 <- as.data.frame(eTof_emmeans_21)
eTof_pairs_21 <- as.data.frame(pairs(eTof_emmeans_21))
eTof_pairs_21$trait <- "Emergence To Flowering"

geno_eTof <- as.data.frame(get_geno_table(Dat_2021_most$EmergeToFlwr))
geno_eTof$trait <- "Emergence To Flowering"
```

### Single Leaf Collected At Flowering

fresh weight
```{r}
# Fresh weight
fresh_lm_21 <- do_lmer(Dat_2021_most$l10_FreshWt)
fresh_anov_21 <- do_anov(fresh_lm_21)
fresh_anov_21$trait <- "l10_FreshWt"
fresh_emmeans_21 <- get_table_bt(fresh_lm_21)
fresh_means_21 <- as.data.frame(fresh_emmeans_21)
fresh_pairs_21 <- as.data.frame(pairs(fresh_emmeans_21))
fresh_pairs_21$trait <- "l10_FreshWt"
```

saturated/hydrated weight

```{r}
# Saturated weight
sat_lm_21 <- do_lmer(Dat_2021_most$l10_SatWt)
sat_anov_21 <- do_anov(sat_lm_21)
sat_anov_21$trait <- "l10_SatWt"
sat_emmeans_21 <- get_table_bt(sat_lm_21)
sat_means_21 <- as.data.frame(sat_emmeans_21)
sat_pairs_21 <- as.data.frame(pairs(sat_emmeans_21))
sat_pairs_21$trait <- "l10_SatWt"
```


dried weight
```{r}
# Dried weight
dry_lm_21 <- do_lmer(Dat_2021_most$l10_DriedWt)
dry_anov_21 <- do_anov(dry_lm_21)
dry_anov_21$trait <- "l10_DriedWt"
dry_emmeans_21 <- get_table_bt(dry_lm_21)
dry_means_21 <- as.data.frame(dry_emmeans_21)
dry_pairs_21 <- as.data.frame(pairs(dry_emmeans_21))
dry_pairs_21$trait <- "l10_DriedWt"

# testing, no meaningfully changes - done before updating lmer and anov to differnt functions
#anov_lmer(Dat_2021_TwoTrt$DriedWt_g)
#anov_lmer(Dat_2021_TwoTrt$SQR_DriedWt)

```

leaf area

```{r}
# Leaf Area
area_lm_21 <- do_lmer(Dat_2021_la$l10_LeafArea, data = Dat_2021_la)
area_anov_21 <- do_anov(area_lm_21)
area_anov_21$trait <- "l10_area"
area_emmeans_21 <- get_table_bt(area_lm_21)
area_means_21 <- as.data.frame(area_emmeans_21)
area_pairs_21 <- as.data.frame(pairs(area_emmeans_21))
area_pairs_21$trait <- "l10_area"
```


specific leaf area

```{r}
# SLA
SLA_lm_21 <- do_lmer(Dat_2021_la$l10_SLA, data = Dat_2021_la)
SLA_anov_21 <- do_anov(SLA_lm_21)
SLA_anov_21$trait <- "l10_SLA"
SLA_emmeans_21 <- get_table_bt(SLA_lm_21)
SLA_means_21 <- as.data.frame(SLA_emmeans_21)
SLA_pairs_21 <- as.data.frame(pairs(SLA_emmeans_21))
SLA_pairs_21$trait <- "l10_SLA"

geno_sla_21 <- as.data.frame(get_geno_table_bt(Dat_2021_la$l10_SLA, data = Dat_2021_la))
geno_sla_21$trait <- "l10_SLA"
```



leaf dry matter content
```{r}
# LDMC
LDMC_lm_21 <- do_lmer(Dat_2021_most$l10_LDMC)
LDMC_anov_21 <- do_anov(LDMC_lm_21)
LDMC_anov_21$trait <- "l10_LDMC"
LDMC_emmeans_21 <- get_table_bt(LDMC_lm_21)
LDMC_means_21 <- as.data.frame(LDMC_emmeans_21)
LDMC_pairs_21 <- as.data.frame(pairs(LDMC_emmeans_21))
LDMC_pairs_21$trait <- "l10_LDMC"
```


relative water content
```{r}
# RWC
RWC_lm_21 <- do_lmer(Dat_2021_most$RWC)
RWC_anov_21 <- do_anov(RWC_lm_21)
RWC_anov_21$trait <- "RWC"
RWC_emmeans_21 <- get_table(RWC_lm_21)
RWC_means_21 <- as.data.frame(RWC_emmeans_21)
RWC_pairs_21 <- as.data.frame(pairs(RWC_emmeans_21))
RWC_pairs_21$trait <- "RWC"

geno_rwc_21 <- as.data.frame(get_geno_table(Dat_2021_most$RWC))
geno_rwc_21$trait <- "RWC"
```

### Leaf Number

Leaf number at harvest
```{r}
# LN at harvest - from all plants - hard to count at this point in life cycle
LN_harv_lm_21 <- do_lmer(Dat_2021_most$RosetteLeafNum)
LN_harv_anov_21 <- do_anov(LN_harv_lm_21)
LN_harv_anov_21$trait <- "LeafNum_harvest"
LN_harv_emmeans_21 <- get_table(LN_harv_lm_21)
LN_harv_means_21 <- as.data.frame(LN_harv_emmeans_21)
LN_harv_pairs_21 <- as.data.frame(pairs(LN_harv_emmeans_21))
LN_harv_pairs_21$trait <- "LeafNum_harvest"
```

### Biomass

Rosette biomass
```{r}
# dry rosette biomass
Ros_lm_21 <- do_lmer(Dat_2021_RosSeed$l10_DryRosG, data = Dat_2021_RosSeed)
Ros_anov_21 <- do_anov(Ros_lm_21)
Ros_anov_21$trait <- "l10_DryRosG"
Ros_emmeans_21 <- get_table_bt(Ros_lm_21)
Ros_means_21 <- as.data.frame(Ros_emmeans_21)
Ros_pairs_21 <- as.data.frame(pairs(Ros_emmeans_21))
Ros_pairs_21$trait <- "l10_DryRosG"
#no interaction

## check how it compares to the sqr transformed
#Ros_lm_21SR <- do_lmer(Dat_2021_TwoTrt$SQR_DryRosG)
#Ros_anov_21SR <- do_anov(Ros_lm_21SR)
#Ros_anov_21SR$trait <- "SQR_DryRosG"
#Ros_means_21SR <- get_table_bt(Ros_lm_21SR)
## commented out b/c no difference in signficance levels.

```

Reproductive biomass
```{r}
# dry reproductive biomass
Repro_lm_21 <- do_lmer(Dat_2021_most$l10_DryReproG)
Repro_anov_21 <- do_anov(Repro_lm_21)
Repro_anov_21$trait <- "l10_DryReproG"
Repro_emmeans_21 <- get_table_bt(Repro_lm_21)
Repro_means_21 <- as.data.frame(Repro_emmeans_21)
Repro_pairs_21 <- as.data.frame(pairs(Repro_emmeans_21))
Repro_pairs_21$trait <- "l10_DryReproG"
```

Above ground biomass allocation (repro/ros)
```{r}
# reproductive biomass divided by rosette biomass (i.e., above ground biomass allocation)
RR_lm_21 <- do_lmer(Dat_2021_RosSeed$l10_R_to_R, data = Dat_2021_RosSeed)
RR_anov_21 <- do_anov(RR_lm_21)
RR_anov_21$trait <- "l10_Repro_to_Ros"
RR_emmeans_21 <- get_table_bt(RR_lm_21)
RR_means_21 <- as.data.frame(RR_emmeans_21)
RR_pairs_21 <- as.data.frame(pairs(RR_emmeans_21))
RR_pairs_21$trait <- "l10_Repro_to_Ros"
```

### Fitness

Fruit Count: (9/20/2024) using BL's fruit counts

```{r}
# number fruits on plant
fruit_lm_21 <- do_lmer(Dat_2021_fitness$FruitCount_BL, data = Dat_2021_fitness)
fruit_anov_21 <- do_anov(fruit_lm_21)
fruit_anov_21$trait <- "FruitNum"
fruit_emmeans_21 <- get_table(fruit_lm_21)
fruit_means_21 <- as.data.frame(fruit_emmeans_21)
fruit_pairs_21 <- as.data.frame(pairs(fruit_emmeans_21))
fruit_pairs_21$trait <- "FruitNum"

geno_fruits <- as.data.frame(get_geno_table(Dat_2021_fitness$FruitCount_BL, data = Dat_2021_fitness))
geno_fruits$trait <- "FruitNum"

```

Code to see if a hurdle model changes results since there is a pattern of zero inflation. Not evaluated because not used in the final manuscript.
```{r eval = FALSE}
# try a hurdle model
library(pscl)
mod.hurdle <- hurdle(FruitCount_BL ~ Treatment * Population  , data = Dat_2021_TwoTrt, dist = "poisson", zero.dist = "binomial") 
summary(mod.hurdle)
hist(residuals(mod.hurdle), breaks = 15) # decent
plot(fitted(mod.hurdle), residuals(mod.hurdle, type = "pearson", scaled = TRUE)) # this looks very similar


mod.zeroinf <- zeroinfl(FruitCount_BL ~ Treatment * Population  , data = Dat_2021_TwoTrt, dist = "poisson")
summary(mod.zeroinf)

#remove zeros from my initial model
mod.nozero <- lmer(FruitCount_BL ~ Treatment * Population + (1|Population:Line) , data = Dat_2021_TwoTrt[Dat_2021_TwoTrt$FruitCount_BL >0, ],contrasts = list(Treatment=contr.sum, Population = contr.sum))
summary(mod.nozero)
hist(residuals(mod.nozero), breaks = 15) # this is more normal
plot(fitted(mod.nozero), residuals(mod.nozero, type = "pearson", scaled = TRUE)) # this looks very similar

# try to get random component back in there
library(glmmTMB)
mod.zeroinf2 <- glmmTMB(FruitCount_BL ~ Treatment * Population + (1|Population:Line), data = Dat_2021_TwoTrt, ziformula = ~Treatment * Population, family = poisson)
summary(mod.zeroinf2)
mod.poisson <- glmmTMB(FruitCount_BL ~ Treatment * Population + (1|Population:Line), data = Dat_2021_TwoTrt, ziformula = ~0, family = poisson)
summary(mod.poisson)
hist(residuals(mod.poisson), breaks = 15) # okay
plot(fitted(mod.poisson), residuals(mod.poisson, type = "pearson"))
plot(fitted(fruit_lm_21), residuals(fruit_lm_21, type = "pearson"))
# okay, so if I fit a poisson model is really where the changes happen, not for adding in the zero inflated component


# get emmeans from the poisson
mod.poisson_emmeans <- get_table(mod.poisson)
mod.poisson_means <- as.data.frame(mod.poisson_emmeans)
mod.poisson_pairs <- as.data.frame(pairs(mod.poisson_emmeans))
mod.poisson_pairs$trait <- "FruitNumP_Pois"

# tester fig
ggplot(data = mod.poisson_means, aes(x = Treatment, y = rate)) +
  geom_point(aes(col=Population, shape = Population), size = 2, position = position_dodge(width = 0.3), show.legend = TRUE) +
  geom_errorbar(aes(ymin=asymp.LCL, ymax = asymp.UCL, col = Population), width = 0.1, size = 0.8, position = position_dodge(width = 0.3), show.legend = TRUE)+
  geom_line(aes(group = Population, col = Population), size = 1, position = position_dodge(width = 0.3))+ 
  labs(y="Fruit Count Poisson Model (rate)", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_x_discrete(expand = c(0.6,0))



# zero inflated poisson
mod.zeroinf2_emmeans <- get_table(mod.zeroinf2)
mod.zeroinf2_means <- as.data.frame(mod.zeroinf2_emmeans)
mod.zeroinf2_pairs <- as.data.frame(pairs(mod.zeroinf2_emmeans))
mod.zeroinf2_pairs$trait <- "FruitNumP_PoisZI"

# tester fig
ggplot(data = mod.zeroinf2_means, aes(x = Treatment, y = rate)) +
  geom_point(aes(col=Population, shape = Population), size = 2, position = position_dodge(width = 0.3), show.legend = TRUE) +
  geom_errorbar(aes(ymin=asymp.LCL, ymax = asymp.UCL, col = Population), width = 0.1, size = 0.8, position = position_dodge(width = 0.3), show.legend = TRUE)+
  geom_line(aes(group = Population, col = Population), size = 1, position = position_dodge(width = 0.3))+ 
  labs(y="Fruit Count Poisson ZI Model (rate)", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_x_discrete(expand = c(0.6,0))
```

Average weight of a seed (collected seed weight / count of collected seeds)
```{r}
# weight of collected seeds - take weight divided by number of seeds counted
AvSeedWt_lm_21 <- do_lmer(Dat_2021_RosSeed$AvgSeedWt, data = Dat_2021_RosSeed)
AvSeedWt_anov_21 <- do_anov(AvSeedWt_lm_21)
AvSeedWt_anov_21$trait <- "Average Seed Wt"
AvSeedWt_emmeans_21 <- get_table(AvSeedWt_lm_21)
AvSeedWt_means_21 <- as.data.frame(AvSeedWt_emmeans_21)
AvSeedWt_pairs_21 <- as.data.frame(pairs(AvSeedWt_emmeans_21))
AvSeedWt_pairs_21$trait <- "Average Seed Wt"

```

average Seed number per fruit -
```{r}
# number seeds from collected fruits
seeds_lm_21 <- do_lmer(Dat_2021_RosSeed$l10_AvgSeedNum, data = Dat_2021_RosSeed)
seeds_anov_21 <- do_anov(seeds_lm_21)
seeds_anov_21$trait <- "l10_Average Seeds per Fruit"
seeds_emmeans_21 <- get_table_bt(seeds_lm_21)
seeds_means_21 <- as.data.frame(seeds_emmeans_21)
seeds_pairs_21 <- as.data.frame(pairs(seeds_emmeans_21))
seeds_pairs_21$trait <- "l10_Average Seeds per Fruit"
# why are residuals so odd? because population explains 0 variance.

geno_seeds <- as.data.frame(get_geno_table_bt(Dat_2021_RosSeed$l10_AvgSeedNum, data = Dat_2021_RosSeed))
geno_seeds$trait <- "l10_seeds per fruit"

## test if SeedCounter matters
#seeds_lm <- lmer(l10_AvgSeedNum ~ Treatment * Population + (1|Population:Line) + (1|SeedCounter), data = Dat_2021_TwoTrt, #contrasts = list(Treatment=contr.sum, Population = contr.sum))
#plot(fitted(seeds_lm), residuals(seeds_lm, type = "pearson", scaled = TRUE),
#       col = c("red", "blue")[as.numeric(Dat_2021_TwoTrt$Population)[complete.cases(Dat_2021_TwoTrt$l10_AvgSeedNum)]],
#       pch = c(16, 15, 17)[as.numeric(Dat_2021_TwoTrt$Treatment)[complete.cases(Dat_2021_TwoTrt$l10_AvgSeedNum)]])
#summary(seeds_lm)
#anova(seeds_lm)
#
## when including seed counter, it does explain some variance (0.0001) but there is more residual (0.0004) and population still explains nothing. The significance levels don't change, and the effect sizes don't really change either though not the standard error and df are different for each predictor
## not going to include in the model.
#
## see if the nesting is doing something unintended. if LineID was the random variable, does anything change?
#Dat_2021_TwoTrt$Line.ID <- as.factor(paste0(Dat_2021_TwoTrt$Population, Dat_2021_TwoTrt$Line))
#seeds_lm2 <- lmer(l10_AvgSeedNum ~ Treatment * Population + (1|Line.ID), data = Dat_2021_TwoTrt, contrasts = #list(Treatment=contr.sum, Population = contr.sum))
#plot(fitted(seeds_lm2), residuals(seeds_lm2, type = "pearson", scaled = TRUE),
#       col = c("red", "blue")[as.numeric(Dat_2021_TwoTrt$Population)[complete.cases(Dat_2021_TwoTrt$l10_AvgSeedNum)]],
#       pch = c(16, 15, 17)[as.numeric(Dat_2021_TwoTrt$Treatment)[complete.cases(Dat_2021_TwoTrt$l10_AvgSeedNum)]])
#summary(seeds_lm2)
#anova(seeds_lm2)
#
## no difference, as expected if the model was functioning as expected. All testing code is commented out and no changes were made to the model.

```

Bonus question about a trade off - are seed weight and seeds per fruit negatively correlated?
```{r}
# this includes unequal genotype representation because 
# overall
cor.test(Dat_2021_RosSeed$AvgSeedNum, Dat_2021_RosSeed$AvgSeedWt)
plot(Dat_2021_RosSeed$AvgSeedNum, Dat_2021_RosSeed$AvgSeedWt)

# just in SW
cor.test(Dat_2021_RosSeed[Dat_2021_RosSeed$Population == "RODA", ]$AvgSeedNum, Dat_2021_RosSeed[Dat_2021_RosSeed$Population == "RODA", ]$AvgSeedWt)
plot(Dat_2021_RosSeed[Dat_2021_RosSeed$Population == "RODA", ]$AvgSeedNum, Dat_2021_RosSeed[Dat_2021_RosSeed$Population == "RODA", ]$AvgSeedWt)

# just in IT
cor.test(Dat_2021_RosSeed[Dat_2021_RosSeed$Population == "BELM", ]$AvgSeedNum, Dat_2021_RosSeed[Dat_2021_RosSeed$Population == "BELM", ]$AvgSeedWt)
plot(Dat_2021_RosSeed[Dat_2021_RosSeed$Population == "BELM", ]$AvgSeedNum, Dat_2021_RosSeed[Dat_2021_RosSeed$Population == "BELM", ]$AvgSeedWt)

# just cur?
cor.test(Dat_2021_RosSeed[Dat_2021_RosSeed$Treatment == "Current", ]$AvgSeedNum, Dat_2021_RosSeed[Dat_2021_RosSeed$Treatment == "Current", ]$AvgSeedWt)
plot(Dat_2021_RosSeed[Dat_2021_RosSeed$Treatment == "Current", ]$AvgSeedNum, Dat_2021_RosSeed[Dat_2021_RosSeed$Treatment == "Current", ]$AvgSeedWt)

# just fut?
cor.test(Dat_2021_RosSeed[Dat_2021_RosSeed$Treatment == "Future", ]$AvgSeedNum, Dat_2021_RosSeed[Dat_2021_RosSeed$Treatment == "Future", ]$AvgSeedWt)
plot(Dat_2021_RosSeed[Dat_2021_RosSeed$Treatment == "Future", ]$AvgSeedNum, Dat_2021_RosSeed[Dat_2021_RosSeed$Treatment == "Future", ]$AvgSeedWt)
```

Total Fitness

```{r}
# total fitness 
fitness_lm_21 <- do_lmer(Dat_2021_fitness$fitness, data = Dat_2021_fitness)
fitness_anov_21 <- do_anov(fitness_lm_21)
fitness_anov_21$trait <- "Total Fitness"
fitness_emmeans_21 <- get_table(fitness_lm_21)
fitness_means_21 <- as.data.frame(fitness_emmeans_21)
fitness_pairs_21 <- as.data.frame(pairs(fitness_emmeans_21))
fitness_pairs_21$trait <- "Total Fitness"

geno_fitness <- as.data.frame(get_geno_table(Dat_2021_fitness$fitness, data = Dat_2021_fitness))
geno_fitness$trait <- "TotalFitness"
```

###Outputs

1) Results table

```{r}
AnovaResults_2021 <- rbind(eTof_anov_21, fresh_anov_21, sat_anov_21, dry_anov_21, area_anov_21, SLA_anov_21, LDMC_anov_21, RWC_anov_21, LN_harv_anov_21, Ros_anov_21, Repro_anov_21, RR_anov_21, fruit_anov_21, AvSeedWt_anov_21, seeds_anov_21, fitness_anov_21)

write.csv(AnovaResults_2021, file = "data/AnovaResults_2021.csv", row.names = TRUE)
```

2) All the tables to read in to the figures code

```{r}
# change column names - order is odd as artifact from updating code
dfs <- c("eTof_means_21", "RWC_means_21", "LN_harv_means_21", "AvSeedWt_means_21", "fruit_means_21", "fitness_means_21", "fresh_means_21", 'sat_means_21', "dry_means_21", "area_means_21", "SLA_means_21", "LDMC_means_21", "Ros_means_21", "Repro_means_21", 'RR_means_21', "seeds_means_21")

for (i in dfs) {
  x=get(i)
  colnames(x) <- c("Treatment", "Population", "Mean", "SE", "Df", "LL_95", "UL_95")
  assign(i,x)
}

# create list of data frames
Means_2021 <- list(eTof_means_21, fresh_means_21, sat_means_21, dry_means_21, area_means_21, SLA_means_21, LDMC_means_21, RWC_means_21, LN_harv_means_21, Ros_means_21, Repro_means_21, RR_means_21, fruit_means_21, AvSeedWt_means_21, seeds_means_21, fitness_means_21)

# name that list
names(Means_2021) <- c("eTof_means_21", "fresh_means_21", 'sat_means_21', "dry_means_21", "area_means_21", "SLA_means_21", "LDMC_means_21", "RWC_means_21", "LN_harv_means_21", "Ros_means_21", "Repro_means_21", 'RR_means_21', "fruit_means_21", "AvSeedWt_means_21", "seeds_means_21", "fitness_means_21")

# save named list
save(Means_2021, file = "data/ModelMeans_2021.robj")

```

Save a .csv with the posthoc results
```{r}
PostHocResults_2021 <- rbind(eTof_pairs_21, fresh_pairs_21, sat_pairs_21, dry_pairs_21, area_pairs_21, SLA_pairs_21, LDMC_pairs_21, RWC_pairs_21, LN_harv_pairs_21, Ros_pairs_21, Repro_pairs_21, RR_pairs_21, fruit_pairs_21, AvSeedWt_pairs_21, seeds_pairs_21, fitness_pairs_21)

write.csv(PostHocResults_2021, file = "data/PostHocResults_2021.csv", row.names = TRUE)
```

genotype level information
```{r}
GenotypeMeans_2021 <- rbind(geno_rwc_21, geno_sla_21, geno_eTof, geno_fruits, geno_seeds, geno_fitness)
GenotypeMeans_2021$year <- "2021"
```


## Sample size information

Identify know how many lines and how many replicates per line for this experiment. The number of observations that were not NA comes from earlier in the "summarize" section.

```{r}
# sample size of 2021 total
Dat_2021_TwoTrt %>% count(Treatment)
# 50 plants per treatments
Dat_2021_TwoTrt %>% count(Treatment, Population)
# in each treatment, 20 are IT, 30 are SW

# there are 21 unique lines. 8 IT, 13 SW.
n_2021 <- Dat_2021_TwoTrt %>%
  group_by(Treatment) %>%
  group_by(Population, .add = TRUE) %>%
  group_by(Line, .add = TRUE) %>%
  count

# only line with 1 rep is Belm 1. Then everything has 2 but RIL parents. B12 has 7, R47 has 6.
n_2021 %>% 
  group_by(Treatment) %>%
  group_by(Population, .add = TRUE) %>%
  summarize(min = min(n), mean = mean(n), median = median(n), max = max(n))


# want to know sample sizes for fitness and such
Dat_2021_fitness %>% count(Treatment, Population)
Dat_2021_RosSeed %>% count(Treatment, Population)
Dat_2021_most %>% count(Treatment, Population)
Dat_2021_la %>% count(Treatment, Population)
```


# 2022

## Read in data
```{r}
Dat_2022 <- read.csv("data/CleanData_2022.csv", header= TRUE)

# make factors
Dat_2022$Treatment <- as.factor(Dat_2022$Treatment)
Dat_2022$Population <- as.factor(Dat_2022$Population)
Dat_2022$PotID <- as.factor(Dat_2022$PotID)
Dat_2022$Flat <- as.factor(Dat_2022$Flat)
Dat_2022$Chamber <- as.factor(Dat_2022$Chamber)
Dat_2022$Transplanted <- as.factor(Dat_2022$Transplanted)
Dat_2022$Geno_ID <- as.factor(paste0(Dat_2022$Population, Dat_2022$Line))
```

## Summarize!

```{r sum}
# by population only
ByPop_22 <- Dat_2022 %>%
  dplyr::group_by(Population) %>%
  summarize(across(.col = c(DaysToEmergence:l10_Stomata_density), .fns = list(mean = ~mean(.x, na.rm = TRUE, .group = "keep"), ci_95 = ~conf_int(.x), n = ~sum(!is.na(.x)))))

# transpose for easier viewing
ByPop_22 <- data.frame(t(ByPop_22))

# by treatment
ByTrt_22 <- Dat_2022 %>%
  dplyr::group_by(Treatment) %>%
  summarize(across(.col = c(DaysToEmergence:l10_Stomata_density), .fns = list(mean = ~mean(.x, na.rm = TRUE, .group = "keep"), ci_95 = ~conf_int(.x), n = ~sum(!is.na(.x)))))
ByTrt_22 <- data.frame(t(ByTrt_22))

# population means within each treatment
ByTrt_Pop_22 <- Dat_2022 %>%
  dplyr::group_by(Population) %>%
  dplyr::group_by(Treatment, .add = TRUE) %>%
  summarize(across(.col = c(DaysToEmergence:l10_Stomata_density), .fns = list(mean = ~mean(.x, na.rm = TRUE, .group = "keep"), ci_95 = ~conf_int(.x), n = ~sum(!is.na(.x)))))

# get the sample sizes
SampleSizes_2022 <- dplyr::select(ByTrt_Pop_22, "Treatment", "Population", ends_with("_n"))
save(SampleSizes_2022, file = "data/SampleSizes_2022.robj")

# for some ease of viewing
ByTrt_Pop_22 <- data.frame(t(ByTrt_Pop_22))
colnames(ByTrt_Pop_22) <- c("Belm_C", "Belm_F", "Roda_C", "Roda_F")

# save this to keep means
write.csv(ByTrt_Pop_22,"data/PopMeansByTreatment_2022.csv", row.names = TRUE )
```

subset to the trait that are in the manuscript
```{r}
Dat_2022c <- Dat_2022[,c("PotID", "Treatment", "Population", "Line", "Geno_ID", "Replicate", "Chamber", "Flat", "Transplanted", "l10_SL_FreshWt", "l10_SL_HydWt", "SL_DryWt", "SL_Area", "l10_SLA", "l10_LDMC", "RWC", "l10_LeafNumber_Total", "l10_AG_DryBiomass", "BG_DryBiomass", "l10_Root_to_Shoot", "TotalBiomass", "l10_Stomata_density", "SLA", "LDMC") ]

# remove B4 Fut because there is no matching B4 current
Dat_2022c <- Dat_2022c[!(Dat_2022c$PotID %in% c("B4-2-F")), ]

# to check for outliers, remove Roda-11 which is a high outlier in the biomass traits 
Dat_2022c_outliers <- Dat_2022c[!(Dat_2022c$Geno_ID %in% c("R11")), ]
```


## Anovas
Now, we can make some models. Initially there were three models for this analysis.

For traits collected before bolting: trait ~ Treatment * Population + (1|Population:Line) + (1|Treatment:Chamber) + (1|Transplanted)

For traits collected when a single leaf was harvesting: trait ~ Treatment * Population + (1|BoltingToHarvest) + (1|Population:Line) + (1|Treatment:Chamber) + (1|Transplanted)

For root harvesting traits: trait ~ Treatment * Population + (1|BoltingToRootWashing) + (1|Population:Line) + (1|Treatment:Chamber) + (1|Transplanted)

Chamber nested within treatment was changed to a fixed effect because models struggle to fit a variance with few groups. The Bolting to harvest and bolting to root washing variables were also removed because they are continuous but being treated like categorical as a random effect. These two variables did explain some variation and removing sometimes increased and sometimes decreased both effect sizes and p values. 

There is a population bias where only Italian lines were transplanted. 17 plants from 6 genotypes were transplanted. They became 10 current plants and 7 future plants. **However, when included, the transplanting effect sometimes explains 0 variance** Because this is such a small number (17 plants of 127 plants), the transplanting effect was removed from the model. All models were run with and without transplanting as a fixed effect, but only small changes were seen in 3 traits (see data/ModelResults_InitalModels.xlsx for notes.)

Final model: trait ~ Treatment * Population + Treatment:Chamber + (1|Population:Line)

```{r}
# set contrasts for models to ensure we are doing the correct type of ANOVA (type-III)
contrasts(Dat_2022c$Population) <- contr.sum
contrasts(Dat_2022c$Treatment) <- contr.sum

# make a modelling function to streamline analyses. This function follows the same pattern as from 2021 analysis

do_lmer2 <- function(trait, data = Dat_2022c){
  lm<- lmer(trait ~ Treatment * Population + Treatment/Chamber + (1|Population:Line), data = data, contrasts = list(Treatment=contr.sum, Population = contr.sum))
  print(plot(lm))
  hist(residuals(lm), breaks = 15)
  #plotting sanity check
  print(paste0("the number of complete cases is ", sum(complete.cases(trait))))
  # make qq plot
  qqnorm(resid(lm))
  qqline(resid(lm))
  print(summary(lm))
  print(confint(lm))
  return(lm)
}
# do the anova
do_anov2 <- function(model){
  anov <- anova(model)
  print(anov)
  df <- as.data.frame(anov)
  df$exp <- "2022"
  return(df)
}

# get table functions are the same as from 2021

# geno get table functions need to be updated
get_geno_table22 <- function (trait, data = Dat_2022c){
  tab2 <- emmeans(lm(trait ~ Treatment * Geno_ID, data = data, contrasts = list(Treatment=contr.sum, Geno_ID = contr.sum)), c("Treatment", "Geno_ID"), type = "reponse")
  return(tab2)
}

get_geno_table_bt22 <- function(trait, data = Dat_2022c){
  mod <- lm(trait ~ Treatment * Geno_ID, data = data, contrasts = list(Treatment=contr.sum, Geno_ID = contr.sum))
  tab2 <- emmeans(update(ref_grid(mod), tran = "log10"), c("Treatment", "Geno_ID"), type = "response") #
  return(tab2)
}
  
```

Run the models

### Single Leaf Traits: bolting

Fresh weight
```{r}
# single leaf fresh weight
fresh_lm_22 <- do_lmer2(Dat_2022c$l10_SL_FreshWt)
fresh_anov_22 <- do_anov2(fresh_lm_22)
fresh_anov_22$trait <- "l10_FreshWt"
fresh_emmeans_22 <- get_table_bt(fresh_lm_22)
fresh_means_22 <- as.data.frame(fresh_emmeans_22)
fresh_pairs_22 <- as.data.frame(pairs(fresh_emmeans_22))
fresh_pairs_22$trait <- "l10_FreshWt"
```

Saturated/Hydrated weight
```{r}
# single leaf hydrated (also called saturated) weight
hyd_lm_22 <- do_lmer2(Dat_2022c$l10_SL_HydWt)
hyd_anov_22 <- do_anov2(hyd_lm_22)
hyd_anov_22$trait <- "l10_SatWt"
hyd_emmeans_22 <- get_table_bt(hyd_lm_22)
hyd_means_22 <- as.data.frame(hyd_emmeans_22)
hyd_pairs_22 <- as.data.frame(pairs(hyd_emmeans_22))
hyd_pairs_22$trait <- "l10_SatWt"

## also has SQR transformation. Pop becomes more significant but no other significance changes
#hyd_lm_22SQR <- do_lmer2(Dat_2022c$SQR_SL_HydWt)
#hyd_anov_22SQR <- do_anov2(hyd_lm_22SQR)
#hyd_anov_22SQR$trait <- "l10_SatWt"
```

Dried weight
```{r}
# single leaf dry weight
dry_lm_22 <- do_lmer2(Dat_2022c$SL_DryWt)
dry_anov_22 <- do_anov2(dry_lm_22)
dry_anov_22$trait <- "DriedWt"
dry_emmeans_22 <- get_table(dry_lm_22)
dry_means_22 <- as.data.frame(dry_emmeans_22)
dry_pairs_22 <- as.data.frame(pairs(dry_emmeans_22))
dry_pairs_22$trait <- "DriedWt"

# what about SQR transformation - no meaningful change
#anov_lmer2(Dat_2022c$SQR_SL_DryWt)

# and what about l10 transformation - no meaningful change
#anov_lmer2(log10(Dat_2022c$SL_DryWt))
```

leaf area
```{r}
# single leaf area
area_lm_22 <- do_lmer2(Dat_2022c$SL_Area)
area_anov_22 <- do_anov2(area_lm_22)
area_anov_22$trait <- "Area"
area_emmeans_22 <- get_table(area_lm_22)
area_means_22 <- as.data.frame(area_emmeans_22)
area_pairs_22 <- as.data.frame(pairs(area_emmeans_22))
area_pairs_22$trait <- "Area"
## also has SQR - no significance changes
#area_lm_22SQR <- do_lmer2(Dat_2022c$SQR_SL_Area)
#area_anov_22SQR <- do_anov2(area_lm_22SQR)
#area_anov_22SQR$trait <- "SQR_Area"
```


leaf perimeter
```{r, eval = FALSE}
# single leaf perimeter
per_lm_22 <- do_lmer2(Dat_2022c$SL_Perim)
per_anov_22 <- do_anov2(per_lm_22)
per_anov_22$trait <- "Perimeter"
per_emmeans_22 <- get_table(per_lm_22)
per_means_22 <- as.data.frame(per_emmeans_22)
per_pairs_22 <- as.data.frame(pairs(per_emmeans_22))
per_pairs_22$trait <- "Perimeter"
```


specific leaf area
```{r}
# SLA
sla_lm_22 <- do_lmer2(Dat_2022c$l10_SLA)
sla_anov_22 <- do_anov2(sla_lm_22)
sla_anov_22$trait <- "l10_SLA"
sla_emmeans_22 <- get_table_bt(sla_lm_22)
sla_means_22 <- as.data.frame(sla_emmeans_22)
sla_pairs_22 <- as.data.frame(pairs(sla_emmeans_22))
sla_pairs_22$trait <- "l10_SLA"

geno_sla_22 <- as.data.frame(get_geno_table_bt22(Dat_2022c$l10_SLA))
geno_sla_22$trait <- "l10_SLA"
```

leaf dry matter content
```{r}
# leaf dry matter content
ldmc_lm_22 <- do_lmer2(Dat_2022c$l10_LDMC)
ldmc_anov_22 <- do_anov2(ldmc_lm_22)
ldmc_anov_22$trait <- "l10_LDMC"
ldmc_emmeans_22 <- get_table_bt(ldmc_lm_22)
ldmc_means_22 <- as.data.frame(ldmc_emmeans_22)
ldmc_pairs_22 <- as.data.frame(pairs(ldmc_emmeans_22))
ldmc_pairs_22$trait <- "l10_LDMC"
```


relative water content
```{r}
# relative water content
rwc_lm_22 <- do_lmer2(Dat_2022c$RWC)
rwc_anov_22 <- do_anov2(rwc_lm_22)
rwc_anov_22$trait <- "RWC"
rwc_emmeans_22 <- get_table(rwc_lm_22)
rwc_means_22 <- as.data.frame(rwc_emmeans_22)
rwc_pairs_22 <- as.data.frame(pairs(rwc_emmeans_22))
rwc_pairs_22$trait <- "RWC"


geno_rwc_22 <- as.data.frame(get_geno_table22(Dat_2022c$RWC))
geno_rwc_22$trait <- "RWC"
```

### Leaf Number

leaf num at bolting
```{r}
# total leaf number counted at bolting when plants were harvested
LN_bolt_lm_22 <- do_lmer2(Dat_2022c$l10_LeafNumber_Total)
LN_bolt_anov_22 <- do_anov2(LN_bolt_lm_22)
LN_bolt_anov_22$trait <- "l10_LeafNumb_bolting"
LN_bolt_emmeans_22 <- get_table_bt(LN_bolt_lm_22)
LN_bolt_means_22<- as.data.frame(LN_bolt_emmeans_22)
LN_bolt_pairs_22 <- as.data.frame(pairs(LN_bolt_emmeans_22))
LN_bolt_pairs_22$trait <- "l10_LeafNumb_bolting"
```

### Biomass

above ground biomass
```{r}
# above ground biomass (same as rosette biomass)
AG_lm_22 <- do_lmer2(Dat_2022c$l10_AG_DryBiomass)
AG_anov_22 <- do_anov2(AG_lm_22)
AG_anov_22$trait <- "l10_Above Ground Biomass"
AG_emmeans_22 <- get_table_bt(AG_lm_22)
AG_means_22 <- as.data.frame(AG_emmeans_22)
AG_pairs_22 <- as.data.frame(pairs(AG_emmeans_22))
AG_pairs_22$trait <- "l10_Above Ground Biomass"
## also has SQR transformation -interaction gets a little weaker
#AG_lm_22SQR <- do_lmer2(Dat_2022c$SQR_AG_DryBiomass)
#AG_anov_22SQR <- do_anov2(AG_lm_22SQR)
#AG_anov_22SQR$trait <- "SQR_Above Ground Biomass"

geno_ag_22 <- as.data.frame(get_geno_table_bt22(Dat_2022c$l10_AG_DryBiomass))
geno_ag_22$trait <- "l10_AG"

# check outliers.
AG_outliers <- do_lmer2(Dat_2022c_outliers$l10_AG_DryBiomass, data = Dat_2022c_outliers)
do_anov2(AG_outliers)
```

below ground biomass
```{r}
# below ground biomass
BG_lm_22 <- do_lmer2(Dat_2022c$BG_DryBiomass)
BG_anov_22 <- do_anov2(BG_lm_22)
BG_anov_22$trait <- "Below Ground Biomass"
BG_emmeans_22 <- get_table(BG_lm_22)
BG_means_22 <- as.data.frame(BG_emmeans_22)
BG_pairs_22 <- as.data.frame(pairs(BG_emmeans_22))
BG_pairs_22$trait <- "Below Ground Biomass"
### also has SQR transformation; non significance changes
#BG_lm_22SQR <- do_lmer2(Dat_2022c$SQR_BG_DryBiomass)
#BG_anov_22SQR <- do_anov2(BG_lm_22SQR)
#BG_anov_22SQR$trait <- "SQR_Below Ground Biomass"

geno_bg_22 <- as.data.frame(get_geno_table22(Dat_2022c$BG_DryBiomass))
geno_bg_22$trait <- "BG_DryBiomass"

# check outliers
BG_outliers <- do_lmer2(Dat_2022c_outliers$BG_DryBiomass, data = Dat_2022c_outliers)
do_anov2(BG_outliers)
```

Root to shoot ratio (below ground / above ground)
```{r}
# Root to Shoot ratio
RS_lm_22 <- do_lmer2(Dat_2022c$l10_Root_to_Shoot)
RS_anov_22 <- do_anov2(RS_lm_22)
RS_anov_22$trait <- "l10_Root to Shoot Ratio"
RS_emmeans_22 <- get_table_bt(RS_lm_22)
RS_means_22 <- as.data.frame(RS_emmeans_22)
RS_pairs_22 <- as.data.frame(pairs(RS_emmeans_22))
RS_pairs_22$trait <- "l10_Root to Shoot Ratio"

geno_rs_22 <- as.data.frame(get_geno_table_bt22(Dat_2022c$l10_Root_to_Shoot))
geno_rs_22$trait <- "l10_Root_to_Shoot"

# check outliers
RS_outliers <- do_lmer2(Dat_2022c_outliers$l10_Root_to_Shoot, data = Dat_2022c_outliers)
do_anov2(RS_outliers)
get_table_bt(RS_outliers)
```

Total biomass
```{r}
# total biomass
TotBio_lm_22 <- do_lmer2(Dat_2022c$TotalBiomass) # residuals pretty normal, don't worry about sqrt transformation
TotBio_anov_22 <- do_anov2(TotBio_lm_22) # these output stats are more different than I would expect...
TotBio_anov_22$trait <- "Total Biomass"
TotBio_emmeans_22 <- get_table(TotBio_lm_22)
TotBio_means_22 <- as.data.frame(TotBio_emmeans_22)
TotBio_pairs_22 <- as.data.frame(pairs(TotBio_emmeans_22))
TotBio_pairs_22$trait <- "Total Biomass"

geno_TotBio_22 <- as.data.frame(get_geno_table22(Dat_2022c$TotalBiomass))
geno_TotBio_22$trait <- "TotalBiomass"

# check outliers
totbio_outliers <- do_lmer2(Dat_2022c_outliers$TotalBiomass, data = Dat_2022c_outliers)
do_anov2(totbio_outliers)
```


### Stomatal Density
```{r}
# stomatal density - imaged on harvest day
sto_den_lm_22 <- do_lmer2(Dat_2022c$l10_Stomata_density)
sto_den_anov_22 <- do_anov2(sto_den_lm_22)
sto_den_anov_22$trait <- "l10_Stomatal Density"
sto_den_emmeans_22 <- get_table_bt(sto_den_lm_22)
sto_den_means_22 <- as.data.frame(sto_den_emmeans_22)
sto_den_pairs_22 <- as.data.frame(pairs(sto_den_emmeans_22))
sto_den_pairs_22$trait <- "l10_Stomatal Density"

# check results don't change if using stomatal average - they didn't change, commented out with update
#do_anov2(do_lmer2(Dat_2022c$l10_Stomata_avg))

geno_stoden_22 <- as.data.frame(get_geno_table_bt22(Dat_2022c$l10_Stomata_density))
geno_stoden_22$trait <- "l10_Stomata_density"

```

### Outputs

1) results table

```{r}
AnovaResults_2022 <- rbind(fresh_anov_22, hyd_anov_22, dry_anov_22, area_anov_22, sla_anov_22, ldmc_anov_22, rwc_anov_22, LN_bolt_anov_22, AG_anov_22, BG_anov_22, RS_anov_22, TotBio_anov_22, sto_den_anov_22 )

write.csv(AnovaResults_2022, file = "data/AnovaResults_2022.csv", row.names = TRUE)
```

2) All the tables to read in to the figures code

```{r}
# change column names
dfs3 <- c("dry_means_22", "area_means_22","rwc_means_22", "BG_means_22", "fresh_means_22", "hyd_means_22", "sla_means_22", "ldmc_means_22", "LN_bolt_means_22", "AG_means_22", "RS_means_22", "TotBio_means_22", "sto_den_means_22")

for (i in dfs3) {
  x=get(i)
  colnames(x) <- c("Treatment", "Population", "Mean", "SE", "Df", "LL_95", "UL_95")
  assign(i,x)
}

# create list of data frames
Means_2022 <- list(fresh_means_22, hyd_means_22, dry_means_22, area_means_22, sla_means_22, ldmc_means_22, rwc_means_22, LN_bolt_means_22, AG_means_22, BG_means_22, RS_means_22, TotBio_means_22, sto_den_means_22)

# name that list
names(Means_2022) <- c("fresh_means_22", "hyd_means_22", "dry_means_22", "area_means_22", "sla_means_22", "ldmc_means_22", "rwc_means_22", "LN_bolt_means_22", "AG_means_22", "BG_means_22", "RS_means_22", "TotBio_means_22", "sto_den_means_22")

# save named list
save(Means_2022, file = "data/ModelMeans_2022.robj")

```

3) Post hoc test results
```{r}
PostHoc_2022 <- rbind(fresh_pairs_22, hyd_pairs_22, dry_pairs_22, area_pairs_22, sla_pairs_22, ldmc_pairs_22, rwc_pairs_22, LN_bolt_pairs_22, AG_pairs_22, BG_pairs_22, RS_pairs_22, TotBio_pairs_22, sto_den_pairs_22 )

write.csv(PostHoc_2022, file = "data/PostHoc_2022.csv", row.names = TRUE)
```

genotype level information
```{r}
GenotypeMeans_2022 <- rbind(geno_rwc_22, geno_sla_22, geno_stoden_22, geno_rs_22, geno_ag_22, geno_bg_22, geno_TotBio_22)
GenotypeMeans_2022$year <- "2022"

GenotypeMeans <- rbind(GenotypeMeans_2021, GenotypeMeans_2022)
write.csv(GenotypeMeans, file = "data/GenotypeMeans.csv", row.names = FALSE)
```

## Sample size information

Identify how many lines and how many replicates per line for this experiment.

```{r}
# sample size of 2022 total
Dat_2022c %>% count(Treatment)
# 64 plants in both
Dat_2022c %>% count(Treatment, Population)
# 32 in each group, with removing the only B4 fut that survived, down to 31 in italy future

# from models know that there are 16 unique lines. 8 IT, 8 SW. ( but 1 line fully died in IT cur)

n_2022 <- Dat_2022c %>%
  group_by(Treatment) %>%
  group_by(Population, .add = TRUE) %>%
  group_by(Line, .add = TRUE) %>%
  count

n_2022 %>% 
  group_by(Treatment) %>%
  summarize(min = min(n), mean = mean(n), median = median(n), max = max(n))

n_2022 %>% 
  group_by(Treatment) %>%
  group_by(Population, .add = TRUE) %>%
  summarize(min = min(n), mean = mean(n), median = median(n), max = max(n))

```

# Notes
The AnovaResults_YEAR files that were output have been copied over to the data/ModelResults.xlsx  to make formatting changes and color code for easier comparison and analysis.

# Problem Solve
## SLA and LDMC correlation
Choosing to only include SLA and move LDMC to the supplemental so I want correlation information for the two traits.
```{r, eval = TRUE}
# 2021
# overall -usng the most dataframe. this does include one genotype for sla that is only in one treatment but the other option is to not include one genotype for ldmc. check with la dataframe to make sure no major differences. 
cor.test(Dat_2021_most$SLA, Dat_2021_most$LDMC)
plot(Dat_2021_most$SLA, Dat_2021_most$LDMC)

cor.test(Dat_2021_la$SLA, Dat_2021_la$LDMC)
# slight difference but not meaningful interpretation wise -> round to same thing to the hundreths place

# just in SW
cor.test(Dat_2021_most[Dat_2021_most$Population == "RODA", ]$SLA, Dat_2021_most[Dat_2021_most$Population == "RODA", ]$LDMC)
plot(Dat_2021_most[Dat_2021_most$Population == "RODA", ]$SLA, Dat_2021_most[Dat_2021_most$Population == "RODA", ]$LDMC)

# just in IT
cor.test(Dat_2021_la[Dat_2021_la$Population == "BELM", ]$SLA, Dat_2021_la[Dat_2021_la$Population == "BELM", ]$LDMC)
plot(Dat_2021_la[Dat_2021_la$Population == "BELM", ]$SLA, Dat_2021_la[Dat_2021_la$Population == "BELM", ]$LDMC)

# just cur?
cor.test(Dat_2021_la[Dat_2021_la$Treatment == "Current", ]$SLA, Dat_2021_la[Dat_2021_la$Treatment == "Current", ]$LDMC)
plot(Dat_2021_la[Dat_2021_la$Treatment == "Current", ]$SLA, Dat_2021_la[Dat_2021_la$Treatment == "Current", ]$LDMC)

# just fut?
cor.test(Dat_2021_la[Dat_2021_la$Treatment == "Future", ]$SLA, Dat_2021_la[Dat_2021_la$Treatment == "Future", ]$LDMC)
plot(Dat_2021_la[Dat_2021_la$Treatment == "Future", ]$SLA, Dat_2021_la[Dat_2021_la$Treatment == "Future", ]$LDMC)

# 2022
# overall
cor.test(Dat_2022c$SLA, Dat_2022c$LDMC)
plot(Dat_2022c$SLA, Dat_2022c$LDMC)

# just in SW
cor.test(Dat_2022c[Dat_2022c$Population == "R", ]$SLA, Dat_2022c[Dat_2022c$Population == "R", ]$LDMC)
plot(Dat_2022c[Dat_2022c$Population == "R", ]$SLA, Dat_2022c[Dat_2022c$Population == "R", ]$LDMC)

# just in IT
cor.test(Dat_2022c[Dat_2022c$Population == "B", ]$SLA, Dat_2022c[Dat_2022c$Population == "B", ]$LDMC)
plot(Dat_2022c[Dat_2022c$Population == "B", ]$SLA, Dat_2022c[Dat_2022c$Population == "B", ]$LDMC)

# just cur?
cor.test(Dat_2022c[Dat_2022c$Treatment == "Current", ]$SLA, Dat_2022c[Dat_2022c$Treatment == "Current", ]$LDMC)
plot(Dat_2022c[Dat_2022c$Treatment == "Current", ]$SLA, Dat_2022c[Dat_2022c$Treatment == "Current", ]$LDMC)

# just fut?
cor.test(Dat_2022c[Dat_2022c$Treatment == "Future", ]$SLA, Dat_2022c[Dat_2022c$Treatment == "Future", ]$LDMC)
plot(Dat_2022c[Dat_2022c$Treatment == "Future", ]$SLA, Dat_2022c[Dat_2022c$Treatment == "Future", ]$LDMC)
```
## Dry weight difference in signficance
For dry weight, there is a huge difference in significance between 2021 (p value below 0.001) and 2022 (p value above 0.95). Quickly looking at the spreadsheets I can not tell a difference and the values are of similar magnitudes. I then did untransformed models and square root transformed models because those were reported as having the most normal residuals in the previous code. None of the models had differences in p values due to transformation that changed the interpretation of the results. I also excluded the "days between bolting and harvest" random effect for the 2022 data but did not find a major difference in the models but the denominator degrees of freedom for treatment did change a lot. As a final check for if there is a potential mistake, I am going to look at some data distributions.

```{r, eval = F}
ggplot(dat = Dat_2021_most)+
  geom_histogram(aes(x = DriedWt_g, fill = Population), alpha = 0.5, position = "identity")+
  theme_classic()
ggplot(dat = Dat_2021_RosSeed)+
  geom_histogram(aes(x = DriedWt_g, fill = Treatment), alpha = 0.5, position = "identity")+
  theme_classic()


ggplot(dat = Dat_2022c)+
  geom_histogram(aes(x = SL_DryWt, fill = Population), alpha = 0.5, position = "identity")+
  theme_classic()
ggplot(dat = Dat_2022c)+
  geom_histogram(aes(x = SL_DryWt, fill = Treatment), alpha = 0.5, position = "identity")+
  theme_classic()
```

Visually looking at the histograms it does seems like 2021 has a treatment difference and 2022 has a population difference. Scales of both are comparable. A log 10 transformation was chosen to be used and these hsitograms are commented out.

## 2021 Avg Seed per fruit
Concerned that line does not explain any variance. Is it true that all the variation is within lines and not between lines? to check this: 1) make histogram of all the values (not line means) - done in CleanData but repeat here

```{r}
hist(Dat_2021_TwoTrt$AvgSeedNum)
hist(Dat_2021_TwoTrt$l10_AvgSeedNum)
```
2) look within lines for difference in avg seeds/fruit - maybe do subtraction and histogram the differences? but won't work if more than two values

```{r}
# start with copy of code from 03_Figures that makes the by line plot
Dat_2021_TwoTrt$Line.ID <- as.factor(paste0(Dat_2021_TwoTrt$Population, Dat_2021_TwoTrt$Line))
Dat_2021_TwoTrt$Line.ID.T <- as.factor(paste0(Dat_2021_TwoTrt$Line.ID, Dat_2021_TwoTrt$Treatment))
df_seed_line <- Dat_2021_TwoTrt %>%
  dplyr::group_by(Line.ID.T, .add = TRUE) %>%
  summarize(across(.col = c(AvgSeedNum, l10_AvgSeedNum), .fns = list(mean = ~mean(.x, na.rm = TRUE, .group = "keep"), ci_95 = ~conf_int(.x), n = ~sum(!is.na(.)))))
df_seed_line$Population <- substr(df_seed_line$Line.ID.T, start = 1, stop = 4)

hist(df_seed_line$AvgSeedNum_n)
# some lines only have 1 pot per treatment -- 10 instances, 7 in Cur
# some lines only have pots in one treatment 4 - all in Fut are the dead ones so only a rep in cur
# 3 and 4 are Roda parent. 6 and 7 are Belm parent.

# get list of Line.ID with 2 replicates
two_reps <- df_seed_line[df_seed_line$AvgSeedNum_n == 2, "Line.ID.T"]$Line.ID.T
# pull those out of the main dataframe
two_reps_seeds <- Dat_2021_TwoTrt[Dat_2021_TwoTrt$Line.ID.T %in% two_reps, c("Treatment", "Population", "Line", "Replicate", "Line.ID", "Line.ID.T", "AvgSeedNum", "l10_AvgSeedNum")]

# find the difference in values
two_reps_diff <-  two_reps_seeds %>% 
  group_by(Line.ID.T) %>% 
  summarize(across(.col = c(AvgSeedNum, l10_AvgSeedNum), .fns = ~diff(.x)))


# table of just RIL parent seed numbers
ril_pars_seed <- Dat_2021_TwoTrt[Dat_2021_TwoTrt$Line.ID %in% c("BELM12", "RODA47"), c("Treatment", "Population", "Line", "Replicate", "Line.ID", "Line.ID.T", "AvgSeedNum", "l10_AvgSeedNum")]
ril_pars_seed[order(ril_pars_seed$Line.ID.T), ]
```


3) figure out which genotypes are the outliers on the residuals plot.The low outlier on the residuals plot is B12. Note for interpretation that the residuals are scaled (so not log scale but divided by the mean)
```{r}
tmp_seed <- lmer(l10_AvgSeedNum ~ Treatment * Population + (1|Population:Line) , data = Dat_2021_TwoTrt, contrasts = list(Treatment=contr.sum, Population = contr.sum))

df_seed <- data.frame(fitted = fitted(tmp_seed),
                      resid = residuals(tmp_seed, type = "pearson", scaled = TRUE),
                      pop = Dat_2021_TwoTrt$Population[complete.cases(Dat_2021_TwoTrt$l10_AvgSeedNum)],
                      trt = Dat_2021_TwoTrt$Treatment[complete.cases(Dat_2021_TwoTrt$l10_AvgSeedNum)],
                      line = Dat_2021_TwoTrt$Line[complete.cases(Dat_2021_TwoTrt$l10_AvgSeedNum)],
                      pot.id = Dat_2021_TwoTrt$Pot.ID[complete.cases(Dat_2021_TwoTrt$l10_AvgSeedNum)])

plot_seed <-  ggplot(df_seed) +
  geom_point(aes(x=fitted, y=resid,
       col = pop, shape = trt, text=pot.id))+
  scale_color_manual(values = c("red", "blue"))+
  theme_classic()
ggplotly(plot_seed)
```

make a plot with genotype on the x axis and seed count per fruit on the y axis. color by pop and treatment.

```{r, echo = F}
ps_plot <- ggplot(Dat_2021_TwoTrt)+
  geom_point(aes(x = Line.ID.T, y = AvgSeedNum, col = Population, shape = Treatment, text = Pot.ID))+
  scale_color_manual(values = c("red", "blue"))+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
ggplotly(ps_plot)

```


# test selection gradient
```{r, eval = FALSE}
# create subset for each pop x trt combination (4 subsets) - only keeping traits related to avoidance/escape that are not too highly correlated and keeps the number of predictors decently low

# sweden - cur has 27 plants, fut has 25
sw_cur <- Dat_2021_TwoTrt[Dat_2021_TwoTrt$Population == "RODA" & Dat_2021_TwoTrt$Treatment == "Current", c("Pot.ID", "Population", "Line", "Replicate", "Treatment", "RWC", "SLA", "EmergeToFlwr", "fitness")]

sw_fut <- Dat_2021_TwoTrt[Dat_2021_TwoTrt$Population == "RODA" & Dat_2021_TwoTrt$Treatment == "Future", c("Pot.ID", "Population", "Line", "Replicate", "Treatment", "RWC", "SLA", "EmergeToFlwr", "fitness")]

# italy - cur has 18 plants, fut has 15 plants
it_cur <- Dat_2021_TwoTrt[Dat_2021_TwoTrt$Population == "BELM" & Dat_2021_TwoTrt$Treatment == "Current", c("Pot.ID", "Population", "Line", "Replicate", "Treatment", "RWC", "SLA", "EmergeToFlwr", "fitness")]

it_fut <- Dat_2021_TwoTrt[Dat_2021_TwoTrt$Population == "BELM" & Dat_2021_TwoTrt$Treatment == "Future", c("Pot.ID", "Population", "Line", "Replicate", "Treatment", "RWC", "SLA", "EmergeToFlwr", "fitness")]

# need to calculate relative fitness and scale the variables
sw_cur <- sw_cur %>%
  mutate(rel.fit = fitness / max(fitness, na.rm = T),
         RWC = c(scale(RWC, center = T, scale = T)),
         SLA = c(scale(SLA, center = T, scale = T)),
         EmergeToFlwr = c(scale(EmergeToFlwr, center = T, scale = T))
  )

sw_fut <- sw_fut %>%
  mutate(rel.fit = fitness / max(fitness, na.rm = T),
         RWC = c(scale(RWC, center = T, scale = T)),
         SLA = c(scale(SLA, center = T, scale = T)),
         EmergeToFlwr = c(scale(EmergeToFlwr, center = T, scale = T))
  )

it_cur <- it_cur %>%
  mutate(rel.fit = fitness / max(fitness, na.rm = T),
         RWC = c(scale(RWC, center = T, scale = T)),
         SLA = c(scale(SLA, center = T, scale = T)),
         EmergeToFlwr = c(scale(EmergeToFlwr, center = T, scale = T))
  )

it_fut <- it_fut %>%
  mutate(rel.fit = fitness / max(fitness, na.rm = T),
         RWC = c(scale(RWC, center = T, scale = T)),
         SLA = c(scale(SLA, center = T, scale = T)),
         EmergeToFlwr = c(scale(EmergeToFlwr, center = T, scale = T))
  )

# run the selection gradients
sg_sw_cur <- lmer(rel.fit ~ RWC + SLA + EmergeToFlwr + (1|Line), data = sw_cur)
summary(sg_sw_cur) # nothing significant

sg_sw_fut <- lmer(rel.fit ~ RWC + SLA + EmergeToFlwr + (1|Line), data = sw_fut)
summary(sg_sw_fut) # actually does show selection for earlier flowering

sg_it_cur <- lmer(rel.fit ~ RWC + SLA + EmergeToFlwr + (1|Line), data = it_cur)
summary(sg_it_cur) # nothing significant

sg_it_fut <- lmer(rel.fit ~ RWC + SLA + EmergeToFlwr + (1|Line), data = it_fut)
summary(sg_it_fut) # nothing significant

```

End of code.