---
title: "03_Figures"
author: "Sophie Buysse"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# description

This code creates figures from both experiments. First, there are figures specific to each experiment showing phenology trends throughout the growing period. Then, there are reaction norms for the traits only collected in one experiment. Finally, there are reaction norms for the traits that were measured in both experiments with lines from both on the same graph. Below is the coloring guide to be consistent between all experiments and be color-blind accessible.

```{r, eval = FALSE, echo = FALSE}
#Full Okabe Ito color palette:
c("#E69F00",  "#56B4E9", "#0072B2",  "#D55E00","#CC79A7",  "#999999", "#000000", "#009E73", "#F5C710")
```

Sweden: #009E73 - a green color \
Italy: #CC79A7 - light pink \
Current: x axis, if needed #56B4E9 - a light blue or a solid line \
Future: x axis, if needed #D55E00 - a dark orange or a dashed line \
2021: solid lines \
2022: dashed lines \

# Prep code

load packages

```{r}
library(dplyr)
library(ggplot2)
library(tidyr)
library(ggpubr)

# would like to phase these out but still used for the growth curves
# let's make some functions
std_mean <- function(x){
  sd(x, na.rm = TRUE)/sqrt(length(x))
}

# Confidence interval multiplier for standard error
    # Calculate t-statistic for confidence interval: 
    # e.g., if conf.interval is .95, use .975 (above/below), and use df=N-1

conf_int <- function(x, conf.interval = 0.95){
  ciMult <- qt((1+conf.interval)/2, length(x)-1)
  ci <- std_mean(x) * ciMult
  return(ci)
}

```

Set theme for plots
```{r}
theme_set(theme_classic())
theme_update(legend.title = element_text(family = "serif", color = "black", size = 12),
    legend.text = element_text(family = "serif", color = "black", size = 12),
    axis.title = element_text(family = "serif", color = "black", size = 12),
    axis.text = element_text(family = "serif", color = "black", size = 12),
    axis.line = element_line(colour = 'black', linewidth = 1),
    axis.ticks = element_line(color = 'black', linewidth = 1),
    axis.ticks.length=unit(.15, "cm"),
    legend.spacing.y = unit(0.03, "cm"),
    panel.border = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank()
    )

# If I want to add back in that the y axis should always start at 0:
#  expand_limits(y=0)
    
```



# read in data

Read in the data needed for all plotting.
```{r}
# 2021
Dat_2021 <- read.csv("data/CleanData_2021.csv")
# set factors
Dat_2021 <- Dat_2021 %>% dplyr::mutate(
  Population = as.factor(Population),
  Line = as.factor(Line),
  Treatment = as.factor(Treatment),
  Soil.Mix.Batch.Number.x = as.factor(Soil.Mix.Batch.Number.x),
  BranchStructure = as.factor(BranchStructure),
  DoneFlwr = as.factor(DoneFlwr)
)
# subset to two treatments
Dat_2021 <- Dat_2021[which(Dat_2021$Treatment == "Current"| Dat_2021$Treatment == "Future") , ]
# for plotting - this file was made with only two treatments
load("data/ModelMeans_2021.robj")
load("data/SampleSizes_2021.robj")


# 2022
Dat_2022 <- read.csv("data/CleanData_2022.csv")
# make some things factors
Dat_2022$Treatment <- as.factor(Dat_2022$Treatment)
Dat_2022$Population <- as.factor(Dat_2022$Population)
Dat_2022$PotID <- as.factor(Dat_2022$PotID)
Dat_2022$Flat <- as.factor(Dat_2022$Flat)
Dat_2022$Chamber <- as.factor(Dat_2022$Chamber)
Dat_2022$Transplanted <- as.factor(Dat_2022$Transplanted)

load("data/ModelMeans_2022.robj")
load("data/SampleSizes_2022.robj")
```

# phenology
Graphs showing general phenology trends. The goal of these is a quick comparison, likely will not be in the manuscript.

2021
```{r, echo = FALSE}
ggplot(data = Dat_2021)+
  geom_density(aes(Emergence, col = Population, linetype = Treatment), size = 1.25)+
  geom_density(aes((Emergence + DayToBolt), col = Population, linetype = Treatment), size = 1.25)+
  geom_density(aes((Emergence + DayToBolt + DayToFlwr), col = Population, linetype = Treatment), size = 1.25)+
  labs(x = "Days Since Planting")+
  scale_color_manual(name = "Population",
                     labels = c("Italy", "Sweden"),
                     values = c("#CC79A7", "#009E73"))+
  theme_classic() +
  theme(
    legend.title = element_text(color = "black", size = 16),
    legend.text = element_text(color = "black", size = 16),
    axis.title = element_text(color = "black", size = 16),
    axis.text = element_text(color = "black", size = 16),
    legend.spacing.y = unit(0.03, "cm"),
    panel.border = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank()
  )

```

2022 - note: plants that died or were transplanted are not shown here because they may have an emergence date and no bolting date (if died) or a bolting date and no emergence date (if transplanted)
```{r, echo = FALSE}
ggplot(data = Dat_2022)+
  geom_density(aes(DaysToEmergence, col = Population, linetype = Treatment), size = 1.25)+
  geom_density(aes((DaysToEmergence + EmergenceToBolting), col = Population, linetype = Treatment), size = 1.25)+
  labs(x = "Days Since Planting")+
  scale_color_manual(name = "Population",
                     labels = c("Italy", "Sweden"),
                     values = c("#CC79A7", "#009E73"))+
  theme_classic() +
  theme(
    legend.title = element_text(color = "black", size = 16),
    legend.text = element_text(color = "black", size = 16),
    axis.title = element_text(color = "black", size = 16),
    axis.text = element_text(color = "black", size = 16),
    legend.spacing.y = unit(0.03, "cm"),
    panel.border = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank()
  )

```


Takeaways: phenology is quicker for both populations in the 2022 experiment. In 2022, Italy bolts around 70 days post planting but it is closer to 85 days in 2021. Similarly, in 2021 Sweden takes over 100 days to bolt but it bolted at about 90 days in 2022. There were some experimental differences (like light and the chambers used and the conditions through vernalization) that could contribute to this difference. 

# 2021 only

Reaction norms to make here: \
- DaysToBolting(b/c exp differences) \
- DaysBoltingToFlwr \
- LeafNum_duringvern (if included) \
- LeafNum_Oct4 (if included) \
- RosetteLeafNum_harvest \
- RosetteDry_harvest \
- DryReproG \
- Repro_to_Ros \
- LatBranches \
- PrimaryStalks \
- Height_cm \
- IJ_FruitCount (alsokept other fruit count in the forplot table just cause) \
- AG_biomass (Ros+Repro) \
- fitness \
- AvgSeedWt \
- AvgSeedNum \
- NumFlwrLeft (no model as of 7/31/2024, not included here, not measured for everything) \


## Phenology

Days Emergence to bolting

```{r, echo = FALSE}
ggplot(data = Means_2021[["bolting_means_21"]], aes(x = Treatment, y = Mean)) +
  geom_point(aes(col=Population), size = 4, show.legend = TRUE) +
  geom_errorbar(aes(ymin=LL_95, ymax = UL_95, col = Population), width = 0.1, size = 1, show.legend = TRUE)+
  geom_line(aes(group = Population, col = Population), size = 1.5)+ 
  labs(y="Days Emergence to Bolting", 
       x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_x_discrete(expand = c(0.6,0))
```


DaysBoltingToFlwr


```{r, echo = FALSE}
ggplot(data = Means_2021[["flowering_means_21"]], aes(x = Treatment, y = Mean)) +
  geom_point(aes(col=Population), size = 4, show.legend = TRUE) +
  geom_errorbar(aes(ymin=LL_95, ymax = UL_95, col = Population), width = 0.1, size = 1, show.legend = TRUE)+
  geom_line(aes(group = Population, col = Population), size = 1.5)+ 
  labs(y="Days Bolting to Flowering", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_x_discrete(expand = c(0.6,0))
```


Days Emergence to Flower
```{r, echo = FALSE}
detf <- ggplot(data = Means_2021[["eTof_means_21"]], aes(x = Treatment, y = Mean)) +
  geom_point(aes(col=Population), size = 4, show.legend = TRUE) +
  geom_errorbar(aes(ymin=LL_95, ymax = UL_95, col = Population), width = 0.1, size = 1, show.legend = TRUE)+
  geom_line(aes(group = Population, col = Population), size = 1.5)+ 
  labs(y="Days to Flowering", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_x_discrete(expand = c(0.6,0))
detf
```

## Leaf Number
LeafNum_duringvern (if included)

```{r, echo = FALSE}
ggplot(data = Means_2021[["LN_Vern_means_21"]], aes(x = Treatment, y = Mean)) +
  geom_point(aes(col=Population), size = 4, show.legend = TRUE) +
  geom_errorbar(aes(ymin=LL_95, ymax = UL_95, col = Population), width = 0.1, size = 1, show.legend = TRUE)+
  geom_line(aes(group = Population, col = Population), size = 1.5)+ 
  labs(y="Leaf Number: during vern", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_x_discrete(expand = c(0.6,0))
```

LeafNum_Oct4 (if included)

```{r, echo = FALSE}
ggplot(data = Means_2021[["LN_PostVern_means_21"]], aes(x = Treatment, y = Mean)) +
  geom_point(aes(col=Population), size = 4, show.legend = TRUE) +
  geom_errorbar(aes(ymin=LL_95, ymax = UL_95, col = Population), width = 0.1, size = 1, show.legend = TRUE)+
  geom_line(aes(group = Population, col = Population), size = 1.5)+ 
  labs(y="Leaf Number: 6 weeks post vern", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_x_discrete(expand = c(0.6,0))
```

RosetteLeafNum_harvest

```{r, echo = FALSE}
ln_harv <- ggplot(data = Means_2021[["LN_harv_means_21"]], aes(x = Treatment, y = Mean)) +
  geom_point(aes(col=Population), size = 4, show.legend = TRUE) +
  geom_errorbar(aes(ymin=LL_95, ymax = UL_95, col = Population), width = 0.1, size = 1, show.legend = TRUE)+
  geom_line(aes(group = Population, col = Population), size = 1.5)+ 
  labs(y="Leaf Number \n at harvest", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_x_discrete(expand = c(0.6,0))
ln_harv
```

Create an extra figure to look at growth curves during this time. I want one line per population and 4 x axis values with the y axis being leaf number.

```{r}
# format a dataframe
tmp_mean_21 <- Dat_2021 %>%
  dplyr::group_by(Population) %>%
  dplyr::group_by(Treatment, .add = TRUE) %>%
  summarize(across(.col = c(LeafNum_.07222021, LeafNum_08262021, RosetteLeafNum_10042021, RosetteLeafNum), .fns = list(mean = ~mean(.x, na.rm = TRUE, .group = "keep"))))
# pivot
tmp_mean_21 <- pivot_longer(tmp_mean_21, cols = !(c(Population, Treatment)), names_to = "Time", values_to = "mean")
tmp_mean_21$Time <- factor(rep(c("Week5", "Week10", "Week14", "Harvest"), times = 4), levels = c("Week5", "Week10", "Week14", "Harvest"))

tmp_ci_21 <- Dat_2021 %>%
  dplyr::group_by(Population) %>%
  dplyr::group_by(Treatment, .add = TRUE) %>%
  summarize(across(.col = c(LeafNum_.07222021, LeafNum_08262021, RosetteLeafNum_10042021, RosetteLeafNum), .fns = list(ci_95 = ~conf_int(.x))))

# pivot
tmp_ci_21 <- pivot_longer(tmp_ci_21, cols = !(c(Population, Treatment)), names_to = "Time", values_to = "ci")
tmp_ci_21$Time <- factor(rep(c("Week5", "Week10", "Week14", "Harvest"), times = 4), levels = c("Week5", "Week10", "Week14", "Harvest"))

forplot3_2021 <- merge(tmp_mean_21, tmp_ci_21)

# plot
ggplot(data = forplot3_2021, aes(x = Time, y = mean)) +
  geom_point(aes(col=Population), size = 4, show.legend = TRUE) +
  #geom_errorbar(aes(ymin=mean-ci, ymax = mean + ci, col = Population), width = 0.1, size = 1, show.legend = TRUE)+
  geom_line(aes(group = Treatment:Population, col = Population, linetype = Treatment), size = 1.5)+ 
  labs(y="Leaf Number", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_x_discrete(expand = c(0.2,0))
```

## Plant Structure
RosetteDry_harvest

```{r, echo = FALSE}
ggplot(data = Means_2021[["Ros_means_21"]], aes(x = Treatment, y = Mean)) +
  geom_point(aes(col=Population), size = 4, show.legend = TRUE) +
  geom_errorbar(aes(ymin=LL_95, ymax = UL_95, col = Population), width = 0.1, size = 1, show.legend = TRUE)+
  geom_line(aes(group = Population, col = Population), size = 1.5)+ 
  labs(y="Log10 Rosette Dry Biomass (g) at harvest", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_x_discrete(expand = c(0.6,0))

# backtransformed
ros_2021 <- ggplot(data = Means_2021[["Ros_means_21"]], aes(x = Treatment, y = Bk_Mean)) +
  geom_point(aes(col=Population), size = 4, show.legend = TRUE) +
  geom_errorbar(aes(ymin=Bk_LL_95, ymax = Bk_UL_95, col = Population), width = 0.1, size = 1, show.legend = TRUE)+
  geom_line(aes(group = Population, col = Population), size = 1.5)+ 
  labs(y="Rosette Biomass (g)\nat harvest", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_x_discrete(expand = c(0.6,0))
ros_2021
```

DryReproG

```{r, echo = FALSE}
ggplot(data = Means_2021[["Repro_means_21"]], aes(x = Treatment, y = Mean)) +
  geom_point(aes(col=Population), size = 4, show.legend = TRUE) +
  geom_errorbar(aes(ymin=LL_95, ymax = UL_95, col = Population), width = 0.1, size = 1, show.legend = TRUE)+
  geom_line(aes(group = Population, col = Population), size = 1.5)+ 
  labs(y="l10 Reproductive Dry Biomass (g)", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_x_discrete(expand = c(0.6,0))

repro <- ggplot(data = Means_2021[["Repro_means_21"]], aes(x = Treatment, y = Bk_Mean)) +
  geom_point(aes(col=Population), size = 4, show.legend = TRUE) +
  geom_errorbar(aes(ymin=Bk_LL_95, ymax = Bk_UL_95, col = Population), width = 0.1, size = 1, show.legend = TRUE)+
  geom_line(aes(group = Population, col = Population), size = 1.5)+ 
  labs(y="Reproductive Biomass (g)", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_x_discrete(expand = c(0.6,0))
repro
```

AG_biomass (Ros+Repro)

```{r, echo = FALSE}
ggplot(data = Means_2021[["AG_means_21"]], aes(x = Treatment, y = Mean)) +
  geom_point(aes(col=Population), size = 4, show.legend = TRUE) +
  geom_errorbar(aes(ymin=LL_95, ymax = UL_95, col = Population), width = 0.1, size = 1, show.legend = TRUE)+
  geom_line(aes(group = Population, col = Population), size = 1.5)+ 
  labs(y="l10 Total Above Ground Biomass (g)", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_x_discrete(expand = c(0.6,0))

ggplot(data = Means_2021[["AG_means_21"]], aes(x = Treatment, y = Bk_Mean)) +
  geom_point(aes(col=Population), size = 4, show.legend = TRUE) +
  geom_errorbar(aes(ymin=Bk_LL_95, ymax = Bk_UL_95, col = Population), width = 0.1, size = 1, show.legend = TRUE)+
  geom_line(aes(group = Population, col = Population), size = 1.5)+ 
  labs(y="Total Above Ground Biomass (g)", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_x_discrete(expand = c(0.6,0))
```

Repro_to_Ros

```{r, echo = FALSE}
ggplot(data = Means_2021[["RR_means_21"]], aes(x = Treatment, y = Mean)) +
  geom_point(aes(col=Population), size = 4, show.legend = TRUE) +
  geom_errorbar(aes(ymin=LL_95, ymax = UL_95, col = Population), width = 0.1, size = 1, show.legend = TRUE)+
  geom_line(aes(group = Population, col = Population), size = 1.5)+ 
  labs(y="l10(Reproductive / Rosette Dry Biomass)", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_x_discrete(expand = c(0.6,0))


rr <- ggplot(data = Means_2021[["RR_means_21"]], aes(x = Treatment, y = Bk_Mean)) +
  geom_point(aes(col=Population), size = 4, show.legend = TRUE) +
  geom_errorbar(aes(ymin=Bk_LL_95, ymax = Bk_UL_95, col = Population), width = 0.1, size = 1, show.legend = TRUE)+
  geom_line(aes(group = Population, col = Population), size = 1.5)+ 
  labs(y=expression(paste(frac("Reproductive", "Rosette"), "   Biomass")), 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_x_discrete(expand = c(0.6,0))
rr
```

LatBranches

```{r, echo = FALSE}
ggplot(data = Means_2021[["LatBranch_means_21"]], aes(x = Treatment, y = Mean)) +
  geom_point(aes(col=Population), size = 4, show.legend = TRUE) +
  geom_errorbar(aes(ymin=LL_95, ymax = UL_95, col = Population), width = 0.1, size = 1, show.legend = TRUE)+
  geom_line(aes(group = Population, col = Population), size = 1.5)+ 
  labs(y="Number of Lateral Branches", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_x_discrete(expand = c(0.6,0))
```

PrimaryStalks - how many stalks are coming out of the top of the rosette?

```{r, echo = FALSE}
ggplot(data = Means_2021[["PrimStalks_means_21"]], aes(x = Treatment, y = Mean)) +
  geom_point(aes(col=Population), size = 4, show.legend = TRUE) +
  geom_errorbar(aes(ymin=LL_95, ymax = UL_95, col = Population), width = 0.1, size = 1, show.legend = TRUE)+
  geom_line(aes(group = Population, col = Population), size = 1.5)+ 
  labs(y="Number of Primary Stalks", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_x_discrete(expand = c(0.6,0))
```

Height_cm (of main stalk)

```{r, echo = FALSE}
ggplot(data = Means_2021[["height_means_21"]], aes(x = Treatment, y = Mean)) +
  geom_point(aes(col=Population), size = 4, show.legend = TRUE) +
  geom_errorbar(aes(ymin=LL_95, ymax = UL_95, col = Population), width = 0.1, size = 1, show.legend = TRUE)+
  geom_line(aes(group = Population, col = Population), size = 1.5)+ 
  labs(y="Height (cm)", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_x_discrete(expand = c(0.6,0))
```

## fitness + fitness related

FruitCount - Basia

```{r, echo = FALSE}
# Basia Fruit Count
ggplot(data = Means_2021[["fruit_means_21"]], aes(x = Treatment, y = Mean)) +
  geom_point(aes(col=Population), size = 4, show.legend = TRUE) +
  geom_errorbar(aes(ymin=LL_95, ymax = UL_95, col = Population), width = 0.1, size = 1, show.legend = TRUE)+
  geom_line(aes(group = Population, col = Population), size = 1.5)+ 
  labs(y="Fruit Count (ImageJ)", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_x_discrete(expand = c(0.6,0))
```

AvgSeedWt

```{r, echo = FALSE}
seed_wt <- ggplot(data = Means_2021[["AvSeedWt_means_21"]], aes(x = Treatment, y = Mean)) +
  geom_point(aes(col=Population), size = 4, show.legend = TRUE) +
  geom_errorbar(aes(ymin=LL_95, ymax = UL_95, col = Population), width = 0.1, size = 1, show.legend = TRUE)+
  geom_line(aes(group = Population, col = Population), size = 1.5)+ 
  labs(y="Average Seed\nWeight (mg)", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_x_discrete(expand = c(0.6,0))
seed_wt
```

AvgSeedNum

```{r, echo = FALSE}
ggplot(data = Means_2021[["seeds_means_21"]], aes(x = Treatment, y = Mean)) +
  geom_point(aes(col=Population), size = 4, show.legend = TRUE) +
  geom_errorbar(aes(ymin=LL_95, ymax = UL_95, col = Population), width = 0.1, size = 1, show.legend = TRUE)+
  geom_line(aes(group = Population, col = Population), size = 1.5)+ 
  labs(y="l10(Average Number of Seeds per Fruit)", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_x_discrete(expand = c(0.6,0))

# backtransformed
ggplot(data = Means_2021[["seeds_means_21"]], aes(x = Treatment, y = Bk_Mean)) +
  geom_point(aes(col=Population), size = 4, show.legend = TRUE) +
  geom_errorbar(aes(ymin=Bk_LL_95, ymax = Bk_UL_95, col = Population), width = 0.1, size = 1, show.legend = TRUE)+
  geom_line(aes(group = Population, col = Population), size = 1.5)+ 
  labs(y="Average Number of Seeds per Fruit", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_x_discrete(expand = c(0.6,0))
```

fitness - avg seed num/fruit * fruit num

```{r, echo = FALSE}
ggplot(data = Means_2021[["fitness_means_21"]], aes(x = Treatment, y = Mean)) +
  geom_point(aes(col=Population), size = 4, show.legend = TRUE) +
  geom_errorbar(aes(ymin=LL_95, ymax = UL_95, col = Population), width = 0.1, size = 1, show.legend = TRUE)+
  geom_line(aes(group = Population, col = Population), size = 1.5)+ 
  labs(y="l10(Fitness)", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_x_discrete(expand = c(0.6,0))

fitness <- ggplot(data = Means_2021[["fitness_means_21"]], aes(x = Treatment, y = Bk_Mean)) +
  geom_point(aes(col=Population), size = 4, show.legend = TRUE) +
  geom_errorbar(aes(ymin=Bk_LL_95, ymax = Bk_UL_95, col = Population), width = 0.1, size = 1, show.legend = TRUE)+
  geom_line(aes(group = Population, col = Population), size = 1.5)+ 
  labs(y="Fitness", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_x_discrete(expand = c(0.6,0))
fitness
```


## graphs by line - include?

Let's look at some things by line instead of Population.

For now, only making a graph by line for seeds per fruit because I want to see if there truly is little variation between lines within a population.The confidence intervals are decently small guess. 

```{r, eval = TRUE}
Dat_2021$Line.ID <- as.factor(paste0(Dat_2021$Population, Dat_2021$Line))

forplot2_2021 <- Dat_2021 %>%
  dplyr::group_by(Treatment) %>%
  dplyr::group_by(Line.ID, .add = TRUE) %>%
  summarize(across(.col = c(Emergence:LeafNum_08262021, RosetteLeafNum_10042021, RosetteLeafNum:DryReproG, Repro_to_Ros, LatBranches:IJ_FruitCount, AG_biomass, AvgSeedWt, AvgSeedNum, fitness), .fns = list(mean = ~mean(.x, na.rm = TRUE, .group = "keep"), ci_95 = ~conf_int(.x))))

# some lines only have 1 pot per treatment
# some lines only have pots in one treatment (3)
forplot2_2021$Population <- substr(forplot2_2021$Line.ID, start = 1, stop = 4)

# Avg Seeds per fruit
ggplot(data = forplot2_2021, aes(x = Treatment, y = AvgSeedNum_mean)) +
  geom_point(aes(col=Population), size = 2, alpha = 0.7, show.legend = TRUE) +
  #geom_errorbar(aes(group = Line.ID, ymin=AvgSeedNum_mean-AvgSeedNum_ci_95, ymax = AvgSeedNum_mean + AvgSeedNum_ci_95, col = Population), width = 0.1, size = 1, alpha = 0.7, show.legend = TRUE)+
  geom_line(aes(group = Line.ID, col = Population), size = 0.7, alpha = 0.7)+
  #geom_line(data = forplot_2021, aes(group = Population, col = Population), size= 1.5)+
  #geom_point(data = forplot_2021, aes(col = Population), size = 4, show.legend = TRUE)+
  labs(y="Average Seeds/Fruit", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_x_discrete(expand = c(0.6,0))


```

This truly shows little variation in line means within each population and treatment. huh. The error bars are very large and maybe skewing the scale so now plotting without error bars and looking at a histogram of means just to see it a different way.

```{r}
ggplot(dat = forplot2_2021)+
  geom_histogram(aes(x = AvgSeedNum_mean, fill = Population), binwidth = 3, alpha = 0.5, position = "identity")+
  scale_fill_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))

ggplot(dat = forplot2_2021)+
  geom_histogram(aes(x = AvgSeedNum_mean, fill = Treatment), binwidth = 3, alpha = 0.5, position = "identity")+
  scale_fill_manual(name = "Treatment",
                     labels = c("Current", "Future"),
                     values = c("#56B4E9", "#D55E00"))
```


# 2022 only

Graphs to make here: \
- LeafNum_Jun6 (if included) \
- LeafNum_Jun13 (if included) \
- LeafNum_bolt \
- AG_Dry_bolt \
- BG_Dry_bolt \
- Root_to_Shoot \
- Stomata_density \

Note: skipping days from emergence to bolting because most of that time is the same treatment. Might want it later but no visualization of that for 2022 at all in this code as of 8/22/2024

## Leaf Number
LeafNum_Jun6 (if included)

```{r, echo = FALSE}
ggplot(data = Means_2022[["LN_Jun6_means_22"]], aes(x = Treatment, y = Mean)) +
  geom_point(aes(col=Population), size = 4, show.legend = TRUE) +
  geom_errorbar(aes(ymin=LL_95, ymax = UL_95, col = Population), width = 0.1, size = 1, show.legend = TRUE)+
  geom_line(aes(group = Population, col = Population), size = 1.5, linetype = "dotted")+ 
  labs(y="Leaf Number 1 week post vern", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_x_discrete(expand = c(0.6,0))
```

LeafNum_Jun13 (if included)

```{r, echo = FALSE}
ggplot(data = Means_2022[["LN_Jun13_means_22"]], aes(x = Treatment, y = Mean)) +
  geom_point(aes(col=Population), size = 4, show.legend = TRUE) +
  geom_errorbar(aes(ymin=LL_95, ymax = UL_95, col = Population), width = 0.1, size = 1, show.legend = TRUE)+
  geom_line(aes(group = Population, col = Population), size = 1.5, linetype = "dotted")+ 
  labs(y="Leaf Number 2 weeks post vern", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_x_discrete(expand = c(0.6,0))
```

LeafNum_bolt

```{r, echo = FALSE}
ggplot(data = Means_2022[["LN_bolt_means_22"]], aes(x = Treatment, y = Mean)) +
  geom_point(aes(col=Population), size = 4, show.legend = TRUE) +
  geom_errorbar(aes(ymin=LL_95, ymax = UL_95, col = Population), width = 0.1, size = 1, show.legend = TRUE)+
  geom_line(aes(group = Population, col = Population), size = 1.5, linetype = "dotted")+ 
  labs(y="l10 Leaf Number at Bolting", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_x_discrete(expand = c(0.6,0))

# back transform
ln_bolt <- ggplot(data = Means_2022[["LN_bolt_means_22"]], aes(x = Treatment, y = Bk_Mean)) +
  geom_point(aes(col=Population), size = 4, show.legend = TRUE) +
  geom_errorbar(aes(ymin=Bk_LL_95, ymax = Bk_UL_95, col = Population), width = 0.1, size = 1, show.legend = TRUE)+
  geom_line(aes(group = Population, col = Population), size = 1.5, linetype = "dotted")+ 
  labs(y="Leaf Number\nat Bolting", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_x_discrete(expand = c(0.6,0))
ln_bolt
```

Create an extra figure to look at growth curves during this time. I want one line per population and 4 x axis values (4 weeks, post vern, 2 weeks post vern) with the y axis being leaf number.

```{r}
# format a dataframe
tmp_mean <- Dat_2022 %>%
  dplyr::group_by(Population) %>%
  dplyr::group_by(Treatment, .add = TRUE) %>%
  summarize(across(.col = c(LeafNumber_PreVern:LeafNumber_Jun13, LeafNumber_Total), .fns = list(mean = ~mean(.x, na.rm = TRUE, .group = "keep"))))
# pivot
tmp_mean <- pivot_longer(tmp_mean, cols = !(c(Population, Treatment)), names_to = "Time", values_to = "mean")
tmp_mean$Time <- factor(rep(c("Week5", "Week9", "Week10", "Bolting"), times = 4), levels = c("Week5", "Week9", "Week10", "Bolting"))

tmp_ci <- Dat_2022 %>%
  dplyr::group_by(Population) %>%
  dplyr::group_by(Treatment, .add = TRUE) %>%
  summarize(across(.col = c(LeafNumber_PreVern:LeafNumber_Jun13, LeafNumber_Total), .fns = list(ci_95 = ~conf_int(.x))))

# pivot
tmp_ci <- pivot_longer(tmp_ci, cols = !(c(Population, Treatment)), names_to = "Time", values_to = "ci")
tmp_ci$Time <- factor(rep(c("Week5", "Week9", "Week10", "Bolting"), times = 4), levels = c("Week5", "Week9", "Week10", "Bolting"))

forplot3_2022 <- merge(tmp_mean, tmp_ci)

# plot
ggplot(data = forplot3_2022, aes(x = Time, y = mean)) +
  geom_point(aes(col=Population), size = 4, show.legend = TRUE) +
  #geom_errorbar(aes(ymin=mean-ci, ymax = mean + ci, col = Population), width = 0.1, size = 1, show.legend = TRUE)+
  geom_line(aes(group = Treatment:Population, col = Population, linetype = Treatment), size = 1.5)+ 
  labs(y="Leaf Number", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_x_discrete(expand = c(0.2,0))
```


## Biomass
AG_Dry_bolt

```{r, echo = FALSE}
ggplot(data = Means_2022[["AG_means_22"]], aes(x = Treatment, y = Mean)) +
  geom_point(aes(col=Population), size = 4, show.legend = TRUE) +
  geom_errorbar(aes(ymin=LL_95, ymax = UL_95, col = Population), width = 0.1, size = 1, show.legend = TRUE)+
  geom_line(aes(group = Population, col = Population), size = 1.5, linetype = "dotted")+ 
  labs(y="l10 Above Ground Biomass (g) at bolting", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_x_discrete(expand = c(0.6,0))

ros_2022 <- ggplot(data = Means_2022[["AG_means_22"]], aes(x = Treatment, y = Bk_Mean)) +
  geom_point(aes(col=Population), size = 4, show.legend = TRUE) +
  geom_errorbar(aes(ymin=Bk_LL_95, ymax = Bk_UL_95, col = Population), width = 0.1, size = 1, show.legend = TRUE)+
  geom_line(aes(group = Population, col = Population), size = 1.5, linetype = "dotted")+ 
  labs(y="Rosette Biomass (g)\nat bolting", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_x_discrete(expand = c(0.6,0))
ros_2022
```

BG_Dry_bolt

```{r, echo = FALSE}
bg <- ggplot(data = Means_2022[["BG_means_22"]], aes(x = Treatment, y = Mean)) +
  geom_point(aes(col=Population), size = 4, show.legend = TRUE) +
  geom_errorbar(aes(ymin=LL_95, ymax = UL_95, col = Population), width = 0.1, size = 1, show.legend = TRUE)+
  geom_line(aes(group = Population, col = Population), size = 1.5, linetype = "dotted")+ 
  labs(y="Root Biomass (g)\nat bolting", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_x_discrete(expand = c(0.6,0))
bg
```

Root_to_Shoot

```{r, echo = FALSE}
ggplot(data = Means_2022[["RS_means_22"]], aes(x = Treatment, y = Mean)) +
  geom_point(aes(col=Population), size = 4, show.legend = TRUE) +
  geom_errorbar(aes(ymin=LL_95, ymax = UL_95, col = Population), width = 0.1, size = 1, show.legend = TRUE)+
  geom_line(aes(group = Population, col = Population), size = 1.5, linetype = "dotted")+ 
  labs(y="l10(Root:Shoot Ratio)", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_x_discrete(expand = c(0.6,0))

# back transformed
rs <- ggplot(data = Means_2022[["RS_means_22"]], aes(x = Treatment, y = Bk_Mean)) +
  geom_point(aes(col=Population), size = 4, show.legend = TRUE) +
  geom_errorbar(aes(ymin=Bk_LL_95, ymax = Bk_UL_95, col = Population), width = 0.1, size = 1, show.legend = TRUE)+
  geom_line(aes(group = Population, col = Population), size = 1.5, linetype = "dotted")+ 
  labs(y=expression(paste(frac("Root", "Shoot"), "   Biomass")), 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_x_discrete(expand = c(0.6,0))
rs
```

## Stomata
Stomata Density

```{r, echo = FALSE}
ggplot(data = Means_2022[["sto_den_means_22"]], aes(x = Treatment, y = Mean)) +
  geom_point(aes(col=Population), size = 4, show.legend = TRUE) +
  geom_errorbar(aes(ymin=LL_95, ymax = UL_95, col = Population), width = 0.1, size = 1, show.legend = TRUE)+
  geom_line(aes(group = Population, col = Population), size = 1.5, linetype = "dotted")+ 
  labs(y="l10 Stomatal Density", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_x_discrete(expand = c(0.6,0))

# back transformed
stoden <- ggplot(data = Means_2022[["sto_den_means_22"]], aes(x = Treatment, y = Bk_Mean)) +
  geom_point(aes(col=Population), size = 4, show.legend = TRUE) +
  geom_errorbar(aes(ymin=Bk_LL_95, ymax = Bk_UL_95, col = Population), width = 0.1, size = 1, show.legend = TRUE)+
  geom_line(aes(group = Population, col = Population), size = 1.5, linetype = "dotted")+ 
  labs(y="Stomatal Density", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_x_discrete(expand = c(0.6,0))
stoden
```

# both
Steps to be completed here.
Subset 2021 dataset. add 2021 column. Subset 2022 dataset. add 2022 column. merge - rbind for each trait. format. plot.

Reaction Norms to create here: \
- Emergence \
- EmergenceToBolting \
- FreshWt \
- HydWt \
- DryWt \
- LeafArea \
- LeafPerimeter \
- SLA \
- LDMC \
- RWC \
- LeafNum_4weeks \

```{r}
# subset 2021 dataset
both_exp <- c("emergence_means_21", "bolting_means_21", "fresh_means_21", "sat_means_21", "dry_means_21", "area_means_21", "per_means_21", "SLA_means_21", "LDMC_means_21", "RWC_means_21", "LN_PreVern_means_21")
both_2021 <- Means_2021[both_exp]
both_2021 <- lapply(both_2021, cbind, Exp = as.factor(c("2021")))

# subset 2022 dataset
both_exp2 <- c("emergence_means_22", "bolting_means_22", "fresh_means_22", "hyd_means_22", "dry_means_22", "area_means_22", "per_means_22", "sla_means_22", "ldmc_means_22", "rwc_means_22", "LN_PreVern_means_22")
both_2022 <- Means_2022[both_exp2]
# change B and R to Belm and Roda
for (i in 1:length(both_2022)){
  both_2022[[i]]$Population <- as.character(both_2022[[i]]$Population)
  both_2022[[i]][both_2022[[i]]$Population == "B", "Population"] <- "BELM"
  both_2022[[i]][both_2022[[i]]$Population == "R", "Population"] <- "RODA"
  both_2022[[i]]$Population <- as.factor(both_2022[[i]]$Population)
}
both_2022 <- lapply(both_2022, cbind, Exp = as.factor(c("2022")))


# merge - do this within a loop
mylist.names <- c("emergence", "bolting", "fresh", "hyd", "dry", "area", "per", "sla", "ldmc", "rwc", "LN")
both_dfs <-  vector("list", length(mylist.names))
names(both_dfs) <- mylist.names

for (i in 1:length(both_dfs)){
  both_dfs[[i]] <- bind_rows(both_2021[[i]], both_2022[[i]])
}
```


## Phenology

Emergence. days to emergence is shown for both but should be interpreted with caution. In the 2022 experiment, all plants were put in the same conditions so there may be genetic differences but treatment differences would be due to chamber placement not due to temperature or water availability.

```{r, echo = FALSE}
#jitter by population
ggplot()+
  geom_point(data = both_dfs[["emergence"]][both_dfs[["emergence"]]$Exp=="2021", ], aes(x = Treatment, y = Mean, group = Population:Exp, col=Population), size = 4, show.legend = FALSE, position = position_dodge2(width = 0.3)) +
  geom_errorbar(data = both_dfs[["emergence"]][both_dfs[["emergence"]]$Exp=="2021", ], aes(group = Population:Exp, x = Treatment, ymin=LL_95, ymax = UL_95, col = Population), width = 0.3, size = 1, position = position_dodge2(width = 0.3),  show.legend = FALSE)+
  geom_line(data = both_dfs[["emergence"]][both_dfs[["emergence"]]$Exp=="2021", ], aes(x = Treatment, y = Mean, group = Population:Exp, col = Population, linetype = Exp), size = 1.5, position = position_dodge(width = 0.3), show.legend = TRUE)+ 
  geom_point(data = both_dfs[["emergence"]][both_dfs[["emergence"]]$Exp=="2022", ], aes(x = Treatment, y = Bk_Mean, group = Population:Exp, col=Population), size = 4, show.legend = FALSE, position = position_dodge2(width = 0.3)) +
  geom_errorbar(data = both_dfs[["emergence"]][both_dfs[["emergence"]]$Exp=="2022", ], aes(group = Population:Exp, x = Treatment, ymin=Bk_LL_95, ymax = Bk_UL_95, col = Population), width = 0.3, size = 1, position = position_dodge2(width = 0.3),  show.legend = FALSE)+
  geom_line(data = both_dfs[["emergence"]][both_dfs[["emergence"]]$Exp=="2022", ], aes(x = Treatment, y = Bk_Mean, group = Population:Exp, col = Population, linetype = Exp), size = 1.5, position = position_dodge(width = 0.3), show.legend = TRUE)+ 
  labs(y="Days to Emergence", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_linetype_manual(name = "Experiment",
                        labels = c("2021", "2022"),
                        values = c("solid", "dotted"))+
  scale_x_discrete(expand = c(0.6,0))

# jitter by experiment
# still trying to figure out how to dodge within experiment and nudge between experiments...
#position = position_dodge(width=0.1) - can add this to the 2021 lines to get them to dodge, but the 2022 lines don't dodge

# I think the actual solution might be to make a temp dataframe for each plot that has the columns that will be plotted intead of calling different geoms for each. then I could also just add dodge to the previous plots to get them to look like I want. Or I would split all the "together" things up so that they plot in the same say.not a focus for now but I will need to decide which figures I want in the end.

ggplot()+
  geom_point(data = both_dfs[["emergence"]][both_dfs[["emergence"]]$Exp=="2021", ], aes(x = Treatment, y = Mean, group = Exp:Population, col=Population), size = 4, show.legend = FALSE ) +
  geom_errorbar(data = both_dfs[["emergence"]][both_dfs[["emergence"]]$Exp=="2021", ], aes(group = Exp:Population, x = Treatment, ymin=LL_95, ymax = UL_95, col = Population), width = 0.1, size = 1,  show.legend = FALSE)+
  geom_line(data = both_dfs[["emergence"]][both_dfs[["emergence"]]$Exp=="2021", ], aes(x = Treatment, y = Mean, group = Exp:Population, col = Population, linetype = Exp), size = 1.5, show.legend = TRUE)+ 
  geom_point(data = both_dfs[["emergence"]][both_dfs[["emergence"]]$Exp=="2022", ], aes(x = Treatment, y = Bk_Mean, group = Exp:Population, col=Population), size = 4, show.legend = FALSE, position = position_nudge(x = 0.15)) +
  geom_errorbar(data = both_dfs[["emergence"]][both_dfs[["emergence"]]$Exp=="2022", ], aes(group = Exp:Population, x = Treatment, ymin=Bk_LL_95, ymax = Bk_UL_95, col = Population), width = 0.1, size = 1, position = position_nudge(x = 0.15),  show.legend = FALSE)+
  geom_line(data = both_dfs[["emergence"]][both_dfs[["emergence"]]$Exp=="2022", ], aes(x = Treatment, y = Bk_Mean, group = Exp:Population, col = Population, linetype = Exp), size = 1.5, position = position_nudge(x = 0.15), show.legend = TRUE)+ 
  labs(y="Days to Emergence", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_linetype_manual(name = "Experiment",
                        labels = c("2021", "2022"),
                        values = c("solid", "dotted"))+
  scale_x_discrete(expand = c(0.6,0))
  
```


EmergenceToBolting. While this was measured in both experiments, it is not comparable becuase in 2022 the conditions were the same for all plants for the first 9 weeks, which is more of the time until the plants start to bolt. Emergence to bolting was moved to an only 2021 trait for the manuscript.

```{r, echo = FALSE, eval = TRUE}
detb <- ggplot(data = both_dfs[["bolting"]], aes(x = Treatment, y = Mean)) +
  geom_point(aes(group = Exp:Population, col=Population), size = 4, position = position_dodge2(width = 0.3), show.legend = FALSE) +
  geom_errorbar(aes(ymin=LL_95, ymax = UL_95,group = Exp:Population, col = Population), width = 0.3, size = 1, position = position_dodge2(width = 0.3), show.legend = FALSE)+
  geom_line(aes(group = Exp:Population,col = Population, linetype = Exp), size = 1.5, position = position_dodge2(width = 0.3), show.legend = TRUE)+
  labs(y="Days to Bolting", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_linetype_manual(name = "Experiment",
                        labels = c("2021", "2022"),
                        values = c("solid", "dotted"))+
  scale_x_discrete(expand = c(0.6,0))
detb
```


## Single Leaf Individual Traits
FreshWt

```{r, echo = FALSE}
ggplot(data = both_dfs[["fresh"]], aes(x = Treatment, y = Bk_Mean)) +
  geom_point(aes(group = Exp:Population, col=Population), size = 4, position = position_dodge(width = 0.3), show.legend = FALSE) +
  geom_errorbar(aes(group = Exp:Population, ymin=Bk_LL_95, ymax = Bk_UL_95, col = Population), width = 0.3, size = 1, position = position_dodge(width = 0.3), show.legend = FALSE)+
  geom_line(aes(group = Exp:Population, col = Population, linetype = Exp), size = 1.5, position = position_dodge(width = 0.3), show.legend = TRUE)+
  labs(y="Fresh Weight (g)", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_linetype_manual(name = "Experiment",
                        labels = c("2021", "2022"),
                        values = c("solid", "dotted"))+
  scale_x_discrete(expand = c(0.6,0))
```


HydWt

```{r, echo = FALSE}
ggplot(data = both_dfs[["hyd"]], aes(x = Treatment, y = Bk_Mean)) +
  geom_point(aes(group = Exp:Population, col=Population), size = 4, position = position_dodge(width = 0.3), show.legend = FALSE) +
  geom_errorbar(aes(group = Exp:Population, ymin=Bk_LL_95, ymax = Bk_UL_95, col = Population), width = 0.3, size = 1, position = position_dodge(width = 0.3), show.legend = FALSE)+
  geom_line(aes(group = Exp:Population, col = Population, linetype = Exp), size = 1.5, position = position_dodge(width = 0.3), show.legend = TRUE)+
  labs(y="Saturated Weight (g)", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_linetype_manual(name = "Experiment",
                        labels = c("2021", "2022"),
                        values = c("solid", "dotted"))+
  scale_x_discrete(expand = c(0.6,0))
```


DryWt

```{r, echo = FALSE}
ggplot() +
  geom_point(data = both_dfs[["dry"]][both_dfs[["dry"]]$Exp=="2022", ], aes(x = Treatment, y = Mean, group = Population:Exp, col=Population), size = 4, show.legend = FALSE, position = position_nudge(x=0.15)) +
  geom_errorbar(data = both_dfs[["dry"]][both_dfs[["dry"]]$Exp=="2022", ], aes(group = Population:Exp, x = Treatment, ymin=LL_95, ymax = UL_95, col = Population), width = 0.1, size = 1, position = position_nudge(x=0.15),  show.legend = FALSE)+
  geom_line(data = both_dfs[["dry"]][both_dfs[["dry"]]$Exp=="2022", ], aes(x = Treatment, y = Mean, group = Population:Exp, col = Population, linetype = Exp), size = 1.5, position = position_nudge(x=0.15), show.legend = TRUE)+ 
  geom_point(data = both_dfs[["dry"]][both_dfs[["dry"]]$Exp=="2021", ], aes(x = Treatment, y = Bk_Mean, group = Population:Exp, col=Population), size = 4, show.legend = FALSE, position = position_identity()) +
  geom_errorbar(data = both_dfs[["dry"]][both_dfs[["dry"]]$Exp=="2021", ], aes(group = Population:Exp, x = Treatment, ymin=Bk_LL_95, ymax = Bk_UL_95, col = Population), width = 0.1, size = 1, position = position_identity(),  show.legend = FALSE)+
  geom_line(data = both_dfs[["dry"]][both_dfs[["dry"]]$Exp=="2021", ], aes(x = Treatment, y = Bk_Mean, group = Population:Exp, col = Population, linetype = Exp), size = 1.5, position = position_identity(), show.legend = TRUE)+ 
  labs(y="Dry Weight (g)", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_linetype_manual(name = "Experiment",
                        labels = c("2021", "2022"),
                        values = c("solid", "dotted"))+
  scale_x_discrete(expand = c(0.6,0))
```


LeafArea

```{r, echo = FALSE}
ggplot() +
  geom_point(data = both_dfs[["area"]][both_dfs[["area"]]$Exp=="2022", ], aes(x = Treatment, y = Mean, group = Population:Exp, col=Population), size = 4, show.legend = FALSE, position = position_nudge(x=0.15)) +
  geom_errorbar(data = both_dfs[["area"]][both_dfs[["area"]]$Exp=="2022", ], aes(group = Population:Exp, x = Treatment, ymin=LL_95, ymax = UL_95, col = Population), width = 0.1, size = 1, position = position_nudge(x=0.15),  show.legend = FALSE)+
  geom_line(data = both_dfs[["area"]][both_dfs[["area"]]$Exp=="2022", ], aes(x = Treatment, y = Mean, group = Population:Exp, col = Population, linetype = Exp), size = 1.5, position = position_nudge(x=0.15), show.legend = TRUE)+ 
  geom_point(data = both_dfs[["area"]][both_dfs[["area"]]$Exp=="2021", ], aes(x = Treatment, y = Bk_Mean, group = Population:Exp, col=Population), size = 4, show.legend = FALSE, position = position_identity()) +
  geom_errorbar(data = both_dfs[["area"]][both_dfs[["area"]]$Exp=="2021", ], aes(group = Population:Exp, x = Treatment, ymin=Bk_LL_95, ymax = Bk_UL_95, col = Population), width = 0.1, size = 1, position = position_identity(),  show.legend = FALSE)+
  geom_line(data = both_dfs[["area"]][both_dfs[["area"]]$Exp=="2021", ], aes(x = Treatment, y = Bk_Mean, group = Population:Exp, col = Population, linetype = Exp), size = 1.5, position = position_identity(), show.legend = TRUE)+ 
  labs(y="Leaf Area (cm^2)", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_linetype_manual(name = "Experiment",
                        labels = c("2021", "2022"),
                        values = c("solid", "dotted"))+
  scale_x_discrete(expand = c(0.6,0))
```


LeafPerimeter

```{r, echo = FALSE}
ggplot() +
  geom_point(data = both_dfs[["per"]][both_dfs[["per"]]$Exp=="2022", ], aes(x = Treatment, y = Mean, group = Population:Exp, col=Population), size = 4, show.legend = FALSE, position = position_nudge(x=0.15)) +
  geom_errorbar(data = both_dfs[["per"]][both_dfs[["per"]]$Exp=="2022", ], aes(group = Population:Exp, x = Treatment, ymin=LL_95, ymax = UL_95, col = Population), width = 0.1, size = 1, position = position_nudge(x=0.15),  show.legend = FALSE)+
  geom_line(data = both_dfs[["per"]][both_dfs[["per"]]$Exp=="2022", ], aes(x = Treatment, y = Mean, group = Population:Exp, col = Population, linetype = Exp), size = 1.5, position = position_nudge(x=0.15), show.legend = TRUE)+ 
  geom_point(data = both_dfs[["per"]][both_dfs[["per"]]$Exp=="2021", ], aes(x = Treatment, y = Bk_Mean, group = Population:Exp, col=Population), size = 4, show.legend = FALSE, position = position_identity()) +
  geom_errorbar(data = both_dfs[["per"]][both_dfs[["per"]]$Exp=="2021", ], aes(group = Population:Exp, x = Treatment, ymin=Bk_LL_95, ymax = Bk_UL_95, col = Population), width = 0.1, size = 1, position = position_identity(),  show.legend = FALSE)+
  geom_line(data = both_dfs[["per"]][both_dfs[["per"]]$Exp=="2021", ], aes(x = Treatment, y = Bk_Mean, group = Population:Exp, col = Population, linetype = Exp), size = 1.5, position = position_identity(), show.legend = TRUE)+ 
  labs(y="Leaf Perimeter (cm)", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_linetype_manual(name = "Experiment",
                        labels = c("2021", "2022"),
                        values = c("solid", "dotted"))+
  scale_x_discrete(expand = c(0.6,0))
```



## Single Leaf Calculated Traits
Specific Leaf Area

```{r, echo = FALSE}
sla <- ggplot(data = both_dfs[["sla"]], aes(x = Treatment, y = Bk_Mean)) +
  geom_point(aes(group = Exp:Population, col=Population), size = 4, position = position_dodge(width = 0.3), show.legend = FALSE) +
  geom_errorbar(aes(group = Exp:Population, ymin=Bk_LL_95, ymax = Bk_UL_95, col = Population), width = 0.3, size = 1, position = position_dodge(width = 0.3), show.legend = FALSE)+
  geom_line(aes(group = Exp:Population, col = Population, linetype = Exp), size = 1.5, position = position_dodge(width = 0.3), show.legend = TRUE)+
  labs(y=expression(paste("Specific Leaf Area  ", (cm^2/g))), 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_linetype_manual(name = "Experiment",
                        labels = c("2021", "2022"),
                        values = c("solid", "dotted"))+
  scale_x_discrete(expand = c(0.6,0))
sla

# both not back transformed
ggplot(data = both_dfs[["sla"]], aes(x = Treatment, y = Mean)) +
  geom_point(aes(group = Exp:Population, col=Population), size = 4, position = position_dodge(width = 0.3), show.legend = FALSE) +
  geom_errorbar(aes(group = Exp:Population, ymin=LL_95, ymax = UL_95, col = Population), width = 0.3, size = 1, position = position_dodge(width = 0.3), show.legend = FALSE)+
  geom_line(aes(group = Exp:Population, col = Population, linetype = Exp), size = 1.5, position = position_dodge(width = 0.3), show.legend = TRUE)+
  labs(y="l10 Specific Leaf Area (cm^2 / g)", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_linetype_manual(name = "Experiment",
                        labels = c("2021", "2022"),
                        values = c("solid", "dotted"))+
  scale_x_discrete(expand = c(0.6,0))


# just 2021 SLA for Emily's tenure talk

ggplot(data = both_2021[["SLA_means_21"]], aes(x = Treatment, y = Bk_Mean)) +
  geom_point(aes(group = Population, col=Population), size = 4, show.legend = FALSE) +
  geom_errorbar(aes(group = Population, ymin=Bk_LL_95, ymax = Bk_UL_95, col = Population), width = 0.3, size = 1, show.legend = FALSE)+
  geom_line(aes(group = Population, col = Population), size = 1.5, show.legend = TRUE)+
  labs(y="Specific Leaf Area (cm^2 / g)", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_x_discrete(expand = c(0.6,0))
ggsave(filename = "figures/SLA_2021.png", dpi=300)

# not back transformed
ggplot(data = both_2021[["SLA_means_21"]], aes(x = Treatment, y = Mean)) +
  geom_point(aes(group = Population, col=Population), size = 4, show.legend = FALSE) +
  geom_errorbar(aes(group = Population, ymin=LL_95, ymax = UL_95, col = Population), width = 0.3, size = 1, show.legend = FALSE)+
  geom_line(aes(group = Population, col = Population), size = 1.5, show.legend = TRUE)+
  labs(y="log10 Specific Leaf Area (cm^2 / g)", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_x_discrete(expand = c(0.6,0))
ggsave(filename = "figures/SLA_l10_2021.png", dpi=300)
```


Leaf Dry Matter Content

```{r, echo = FALSE}
ldmc <- ggplot(data = both_dfs[["ldmc"]], aes(x = Treatment, y = Bk_Mean)) +
  geom_point(aes(group = Exp:Population, col=Population), size = 4, position = position_dodge(width = 0.3), show.legend = FALSE) +
  geom_errorbar(aes(group = Exp:Population, ymin=Bk_LL_95, ymax = Bk_UL_95, col = Population), width = 0.3, size = 1, position = position_dodge(width = 0.3), show.legend = FALSE)+
  geom_line(aes(group = Exp:Population, col = Population, linetype = Exp), size = 1.5, position = position_dodge(width = 0.3), show.legend = TRUE)+
  labs(y="Leaf Dry Matter Content (%)", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_linetype_manual(name = "Experiment",
                        labels = c("2021", "2022"),
                        values = c("solid", "dotted"))+
  scale_x_discrete(expand = c(0.6,0))
ldmc
```


Relative Water Content

```{r, echo = FALSE}
rwc <- ggplot(data = both_dfs[["rwc"]], aes(x = Treatment, y = Mean)) +
  geom_point(aes(group = Exp:Population, col=Population), size = 4, position = position_dodge(width = 0.3), show.legend = FALSE) +
  geom_errorbar(aes(group = Exp:Population, ymin=LL_95, ymax = UL_95, col = Population), width = 0.3, size = 1, position = position_dodge(width = 0.3), show.legend = FALSE)+
  geom_line(aes(group = Exp:Population, col = Population, linetype = Exp), size = 1.5, position = position_dodge(width = 0.3), show.legend = TRUE)+
  labs(y="Relative Water Content (%)", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_linetype_manual(name = "Experiment",
                        labels = c("2021", "2022"),
                        values = c("solid", "dotted"))+
  scale_x_discrete(expand = c(0.6,0))
rwc
```

## Leaf Number
LeafNum_4weeks

```{r, echo = FALSE}
ggplot(data = both_dfs[["LN"]], aes(x = Treatment, y = Mean)) +
  geom_point(aes(group = Exp:Population, col=Population), size = 4, position = position_dodge(width = 0.3), show.legend = FALSE) +
  geom_errorbar(aes(group = Exp:Population, ymin=LL_95, ymax = UL_95, col = Population), width = 0.3, size = 1, position = position_dodge(width = 0.3), show.legend = FALSE)+
  geom_line(aes(group = Exp:Population, col = Population, linetype = Exp), size = 1.5, position = position_dodge(width = 0.3), show.legend = TRUE)+
  labs(y="Leaf Number (4 weeks post planting)", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_linetype_manual(name = "Experiment",
                        labels = c("2021", "2022"),
                        values = c("solid", "dotted"))+
  scale_x_discrete(expand = c(0.6,0))
```

# Manuscript Figures
creating multipanel figures for the manuscript. Updated 4/17/2025 with ideas for cutting down/restructuring results to focus on avoidance/escape

Figure 1 \
```{r}
fig1 <- ggarrange(detf, rwc,
                  labels = c("A", "B"),
                  ncol = 2, nrow = 1,
                  common.legend = TRUE,
                  legend = "right",
                  align = "h")
fig1
```

Figure 2 \
```{r}
fig2 <- ggarrange(sla +rremove("xlab"), ldmc+rremove("xlab"), rs, stoden,
                  labels = c("A", "B", "C", "D"),
                  ncol = 2, nrow = 2,
                  common.legend = TRUE,
                  legend = "right",
                  align = "hv"
                  )
fig2
```

Figure 3 \
```{r}
fig3 <- fitness
fig3
```

# supplemental figures not updated

Figure s1 \
```{r}
figs1 <- ggarrange(detb, repro,
                  labels = c("A", "B"),
                  ncol = 2, nrow = 1,
                  common.legend = TRUE,
                  legend = "right",
                  align = "h"
                  )
figs1
```

##Save figures
```{r}
#fig1
ggsave(fig1, filename = "figures/fig1.jpg", width = 4, height = 2.5, units = "in", dpi = 300)

#fig2
ggsave(fig2, filename = "figures/fig2.jpg", width = 6, height = 2.5, units = "in", dpi = 300)

#fig3
ggsave(fig3, filename = "figures/fig3.jpg", width = 12, height = 5, units = "in", dpi = 300)

#fig4
ggsave(fig4, filename = "figures/fig4.jpg", width = 8, height = 2.5, units = "in", dpi = 300)

#figs1
ggsave(figs1, filename = "figures/figs1.jpg", width = 6, height = 2.5, units = "in", dpi = 300)

```

