---
title: "03_Figures"
author: "Sophie Buysse"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# description

This code creates figures from both experiments. First, there are figures specific to each experiment showing phenology trends throughout the growing period. Then, there are reaction norms for the traits only collected in one experiment. Finally, there are reaction norms for the traits that were measured in both experiments with lines from both on the same graph. Below is the coloring guide to be consistent between all experiments and be color-blind accessible. If a trait was transformed, it is back transformed for the figures.

```{r, eval = FALSE, echo = FALSE}
#Full Okabe Ito color palette:
c("#E69F00",  "#56B4E9", "#0072B2",  "#D55E00","#CC79A7",  "#999999", "#000000", "#009E73", "#F5C710")
```

Sweden: #009E73 - a green color \
Italy: #CC79A7 - light pink \
Current: x axis, if needed #56B4E9 - a light blue or a solid line \
Future: x axis, if needed #D55E00 - a dark orange or a dashed line \
2021: solid lines \
2022: dashed lines \

# Prep code

load packages

```{r}
library(dplyr)
library(ggplot2)
library(tidyr)
library(ggpubr)

# functions used for growth curves
std_mean <- function(x){
  sd(x, na.rm = TRUE)/sqrt(length(x))
}

# Confidence interval multiplier for standard error
conf_int <- function(x, conf.interval = 0.95){
  ciMult <- qt((1+conf.interval)/2, length(x)-1)
  ci <- std_mean(x) * ciMult
  return(ci)
}

```

Set theme for plots
```{r}
theme_set(theme_classic())
theme_update(legend.title = element_text(family = "serif", color = "black", size = 14),
    legend.text = element_text(family = "serif", color = "black", size = 14),
    axis.title = element_text(family = "serif", color = "black", size = 14),
    axis.text = element_text(family = "serif", color = "black", size = 14),
    axis.line = element_line(colour = 'black', linewidth = 1),
    axis.ticks = element_line(color = 'black', linewidth = 1),
    axis.ticks.length=unit(.15, "cm"),
    legend.spacing.y = unit(0.03, "cm"),
    panel.border = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    plot.margin = margin(t=1, r=0.075, b=0.075, l=0.05, unit = "in")
    )

# If I want to add back in that the y axis should always start at 0:
#  expand_limits(y=0)

# what font is serif?
windowsFonts()$serif
    
```



# read in data

Read in the data needed for all plotting.
```{r}
# 2021
Dat_2021 <- read.csv("data/CleanData_2021.csv")
# set factors
Dat_2021 <- Dat_2021 %>% dplyr::mutate(
  Population = as.factor(Population),
  Line = as.factor(Line),
  Treatment = as.factor(Treatment),
  Soil.Mix.Batch.Number.x = as.factor(Soil.Mix.Batch.Number.x),
  BranchStructure = as.factor(BranchStructure),
  DoneFlwr = as.factor(DoneFlwr)
)

# for plotting
load("data/ModelMeans_2021.robj")
load("data/SampleSizes_2021.robj")


# 2022
Dat_2022 <- read.csv("data/CleanData_2022.csv")
# make some things factors
Dat_2022$Treatment <- as.factor(Dat_2022$Treatment)
Dat_2022$Population <- as.factor(Dat_2022$Population)
Dat_2022$PotID <- as.factor(Dat_2022$PotID)
Dat_2022$Flat <- as.factor(Dat_2022$Flat)
Dat_2022$Chamber <- as.factor(Dat_2022$Chamber)
Dat_2022$Transplanted <- as.factor(Dat_2022$Transplanted)

load("data/ModelMeans_2022.robj")
load("data/SampleSizes_2022.robj")
```

# phenology
Graphs showing general phenology trends. The goal of these is a quick comparison.

2021
```{r, echo = FALSE}
ggplot(data = Dat_2021)+
  geom_density(aes(Emergence, col = Population, linetype = Treatment), size = 1.25)+
  geom_density(aes((Emergence + DayToBolt), col = Population, linetype = Treatment), size = 1.25)+
  geom_density(aes((Emergence + DayToBolt + DayToFlwr), col = Population, linetype = Treatment), size = 1.25)+
  labs(x = "Days Since Planting")+
  scale_color_manual(name = "Population",
                     labels = c("Italy", "Sweden"),
                     values = c("#CC79A7", "#009E73"))+
  theme_classic() +
  theme(
    legend.title = element_text(color = "black", size = 16),
    legend.text = element_text(color = "black", size = 16),
    axis.title = element_text(color = "black", size = 16),
    axis.text = element_text(color = "black", size = 16),
    legend.spacing.y = unit(0.03, "cm"),
    panel.border = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank()
  )

```

2022 - note: plants that died or were transplanted are not shown here because they may have an emergence date and no bolting date (if died) or a bolting date and no emergence date (if transplanted)
```{r, echo = FALSE}
ggplot(data = Dat_2022)+
  geom_density(aes(DaysToEmergence, col = Population, linetype = Treatment), size = 1.25)+
  geom_density(aes((DaysToEmergence + EmergenceToBolting), col = Population, linetype = Treatment), size = 1.25)+
  labs(x = "Days Since Planting")+
  scale_color_manual(name = "Population",
                     labels = c("Italy", "Sweden"),
                     values = c("#CC79A7", "#009E73"))+
  theme_classic() +
  theme(
    legend.title = element_text(color = "black", size = 16),
    legend.text = element_text(color = "black", size = 16),
    axis.title = element_text(color = "black", size = 16),
    axis.text = element_text(color = "black", size = 16),
    legend.spacing.y = unit(0.03, "cm"),
    panel.border = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank()
  )

```


Takeaways: phenology is quicker for both populations in the 2022 experiment. In 2022, Italy bolts around 70 days post planting but it is closer to 85 days in 2021. Similarly, in 2021 Sweden takes over 100 days to bolt but it bolted at about 90 days in 2022. There were some experimental differences (like light and the chambers used and the conditions/timing through vernalization) that could contribute to this difference. 

# 2021 only

Reaction norms to make here: \
- DaysToBolting(b/c exp differences) \
- DaysBoltingToFlwr \
- LeafNum_duringvern (if included) \
- LeafNum_Oct4 (if included) \
- RosetteLeafNum_harvest \
- RosetteDry_harvest \
- DryReproG \
- Repro_to_Ros \
- LatBranches \
- PrimaryStalks \
- Height_cm \
- IJ_BasiaCounts \
- AG_biomass (Ros+Repro) \
- fitness \
- AvgSeedWt \
- AvgSeedNum \
\

## Phenology

Days Emergence to bolting

```{r, echo = FALSE}
ggplot(data = Means_2021[["bolting_means_21"]], aes(x = Treatment, y = Mean)) +
  geom_point(aes(col=Population, shape = Population), size = 3, position = position_dodge(width = 0.3), show.legend = TRUE) +
  geom_errorbar(aes(ymin=LL_95, ymax = UL_95, col = Population), width = 0.3, size = 0.8, position = position_dodge(width = 0.3), show.legend = TRUE)+
  geom_line(aes(group = Population, col = Population), size = 1, position = position_dodge(width = 0.3))+ 
  labs(y="Days Emergence to Bolting", 
       x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_x_discrete(expand = c(0.6,0))
```


DaysBoltingToFlwr


```{r, echo = FALSE}
ggplot(data = Means_2021[["flowering_means_21"]], aes(x = Treatment, y = Mean)) +
  geom_point(aes(col=Population, shape = Population), size = 3, position = position_dodge(width = 0.3), show.legend = TRUE) +
  geom_errorbar(aes(ymin=LL_95, ymax = UL_95, col = Population), width = 0.3, size = 0.8, position = position_dodge(width = 0.3), show.legend = TRUE)+
  geom_line(aes(group = Population, col = Population), size = 1, position = position_dodge(width = 0.3))+ 
  labs(y="Days Bolting to Flowering", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_x_discrete(expand = c(0.6,0))
```


Days Emergence to Flower
```{r, echo = FALSE}
detf <- ggplot(data = Means_2021[["eTof_means_21"]], aes(x = Treatment, y = Mean)) +
  geom_point(aes(col=Population, shape = Population), size = 3, position = position_dodge(width = 0.3), show.legend = TRUE) +
  geom_errorbar(aes(ymin=LL_95, ymax = UL_95, col = Population), width = 0.3, size = 0.8, position = position_dodge(width = 0.3), show.legend = TRUE)+
  geom_line(aes(group = Population, col = Population), size = 1, position = position_dodge(width = 0.3))+ 
  labs(y="Days to Flowering", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_x_discrete(expand = c(0.6,0))
detf
```

## Leaf Number
LeafNum_duringvern 

```{r, echo = FALSE}
ggplot(data = Means_2021[["LN_Vern_means_21"]], aes(x = Treatment, y = Mean)) +
  geom_point(aes(col=Population, shape = Population), size = 3, position = position_dodge(width = 0.3), show.legend = TRUE) +
  geom_errorbar(aes(ymin=LL_95, ymax = UL_95, col = Population), width = 0.3, size = 0.8, position = position_dodge(width = 0.3), show.legend = TRUE)+
  geom_line(aes(group = Population, col = Population), size = 1, position = position_dodge(width = 0.3))+ 
  labs(y="Leaf Number: during vern", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_x_discrete(expand = c(0.6,0))
```

LeafNum_Oct4 

```{r, echo = FALSE}
ggplot(data = Means_2021[["LN_PostVern_means_21"]], aes(x = Treatment, y = Mean)) +
  geom_point(aes(col=Population, shape = Population), size = 3, position = position_dodge(width = 0.3), show.legend = TRUE) +
  geom_errorbar(aes(ymin=LL_95, ymax = UL_95, col = Population), width = 0.3, size = 0.8, position = position_dodge(width = 0.3), show.legend = TRUE)+
  geom_line(aes(group = Population, col = Population), size = 1, position = position_dodge(width = 0.3))+ 
  labs(y="Leaf Number: 6 weeks post vern", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_x_discrete(expand = c(0.6,0))
```

RosetteLeafNum_harvest

```{r, echo = FALSE}
ln_harv <- ggplot(data = Means_2021[["LN_harv_means_21"]], aes(x = Treatment, y = Mean)) +
  geom_point(aes(col=Population, shape = Population), size = 3, position = position_dodge(width = 0.3), show.legend = TRUE) +
  geom_errorbar(aes(ymin=LL_95, ymax = UL_95, col = Population), width = 0.3, size = 0.8, position = position_dodge(width = 0.3), show.legend = TRUE)+
  geom_line(aes(group = Population, col = Population), size = 1, position = position_dodge(width = 0.3))+ 
  labs(y="Leaf Number \n at harvest", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_x_discrete(expand = c(0.6,0))
ln_harv
```

Create an extra figure to look at growth curves during this time. I want one line per population and 4 x axis values with the y axis being leaf number.

```{r}
# format a dataframe
tmp_mean_21 <- Dat_2021 %>%
  dplyr::group_by(Population) %>%
  dplyr::group_by(Treatment, .add = TRUE) %>%
  summarize(across(.col = c(LeafNum_.07222021, LeafNum_08262021, RosetteLeafNum_10042021, RosetteLeafNum), .fns = list(mean = ~mean(.x, na.rm = TRUE, .group = "keep"))))
# pivot
tmp_mean_21 <- pivot_longer(tmp_mean_21, cols = !(c(Population, Treatment)), names_to = "Time", values_to = "mean")
tmp_mean_21$Time <- factor(rep(c("Week5", "Week10", "Week14", "Harvest"), times = 4), levels = c("Week5", "Week10", "Week14", "Harvest"))

tmp_ci_21 <- Dat_2021 %>%
  dplyr::group_by(Population) %>%
  dplyr::group_by(Treatment, .add = TRUE) %>%
  summarize(across(.col = c(LeafNum_.07222021, LeafNum_08262021, RosetteLeafNum_10042021, RosetteLeafNum), .fns = list(ci_95 = ~conf_int(.x))))

# pivot
tmp_ci_21 <- pivot_longer(tmp_ci_21, cols = !(c(Population, Treatment)), names_to = "Time", values_to = "ci")
tmp_ci_21$Time <- factor(rep(c("Week5", "Week10", "Week14", "Harvest"), times = 4), levels = c("Week5", "Week10", "Week14", "Harvest"))

forplot3_2021 <- merge(tmp_mean_21, tmp_ci_21)

# plot
ggplot(data = forplot3_2021, aes(x = Time, y = mean)) +
  geom_point(aes(col=Population, shape = Population), size = 3, show.legend = TRUE) +
  #geom_errorbar(aes(ymin=mean-ci, ymax = mean + ci, col = Population), width = 0.1, size = 0.8, show.legend = TRUE)+
  geom_line(aes(group = Treatment:Population, col = Population, linetype = Treatment), size = 1)+ 
  labs(y="Leaf Number", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_x_discrete(expand = c(0.2,0))
```

## Plant Structure
RosetteDry_harvest

```{r, echo = FALSE}
ros_2021 <- ggplot(data = Means_2021[["Ros_means_21"]], aes(x = Treatment, y = Mean)) +
  geom_point(aes(col=Population, shape = Population), size = 3, position = position_dodge(width = 0.3), show.legend = TRUE) +
  geom_errorbar(aes(ymin=LL_95, ymax = UL_95, col = Population), width = 0.3, size = 0.8, position = position_dodge(width = 0.3), show.legend = TRUE)+
  geom_line(aes(group = Population, col = Population), size = 1, position = position_dodge(width = 0.3))+ 
  labs(y="Rosette Dry Biomass (g) at harvest", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_x_discrete(expand = c(0.6,0))


ros_2021
```

DryReproG

```{r, echo = FALSE}
repro <- ggplot(data = Means_2021[["Repro_means_21"]], aes(x = Treatment, y = Mean)) +
  geom_point(aes(col=Population, shape = Population), size = 3, position = position_dodge(width = 0.3), show.legend = TRUE) +
  geom_errorbar(aes(ymin=LL_95, ymax = UL_95, col = Population), width = 0.3, size = 0.8, position = position_dodge(width = 0.3), show.legend = TRUE)+
  geom_line(aes(group = Population, col = Population), size = 1, position = position_dodge(width = 0.3))+ 
  labs(y="Reproductive Dry Biomass (g)", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_x_discrete(expand = c(0.6,0))

repro
```

AG_biomass (Ros+Repro)

```{r, echo = FALSE}
ggplot(data = Means_2021[["AG_means_21"]], aes(x = Treatment, y = Mean)) +
  geom_point(aes(col=Population, shape = Population), size = 3, position = position_dodge(width = 0.3), show.legend = TRUE) +
  geom_errorbar(aes(ymin=LL_95, ymax = UL_95, col = Population), width = 0.3, size = 0.8, position = position_dodge(width = 0.3), show.legend = TRUE)+
  geom_line(aes(group = Population, col = Population), size = 1, position = position_dodge(width = 0.3))+ 
  labs(y="Total Above Ground Biomass (g)", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_x_discrete(expand = c(0.6,0))

```

Repro_to_Ros

```{r, echo = FALSE}
rr <- ggplot(data = Means_2021[["RR_means_21"]], aes(x = Treatment, y = Mean)) +
  geom_point(aes(col=Population, shape = Population), size = 3, position = position_dodge(width = 0.3), show.legend = TRUE) +
  geom_errorbar(aes(ymin=LL_95, ymax = UL_95, col = Population), width = 0.3, size = 0.8, position = position_dodge(width = 0.3), show.legend = TRUE)+
  geom_line(aes(group = Population, col = Population), size = 1, position = position_dodge(width = 0.3))+ 
  labs(y="Reproductive / Rosette Dry Biomass", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_x_discrete(expand = c(0.6,0))


rr
```

LatBranches

```{r, echo = FALSE}
ggplot(data = Means_2021[["LatBranch_means_21"]], aes(x = Treatment, y = Mean)) +
  geom_point(aes(col=Population, shape = Population), size = 3, position = position_dodge(width = 0.3), show.legend = TRUE) +
  geom_errorbar(aes(ymin=LL_95, ymax = UL_95, col = Population), width = 0.3, size = 0.8, position = position_dodge(width = 0.3), show.legend = TRUE)+
  geom_line(aes(group = Population, col = Population), size = 1, position = position_dodge(width = 0.3))+ 
  labs(y="Number of Lateral Branches", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_x_discrete(expand = c(0.6,0))
```

PrimaryStalks - how many stalks are coming out of the top of the rosette?

```{r, echo = FALSE}
ggplot(data = Means_2021[["PrimStalks_means_21"]], aes(x = Treatment, y = Mean)) +
  geom_point(aes(col=Population, shape = Population), size = 3, position = position_dodge(width = 0.3), show.legend = TRUE) +
  geom_errorbar(aes(ymin=LL_95, ymax = UL_95, col = Population), width = 0.3, size = 0.8, position = position_dodge(width = 0.3), show.legend = TRUE)+
  geom_line(aes(group = Population, col = Population), size = 1, position = position_dodge(width = 0.3))+ 
  labs(y="Number of Primary Stalks", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_x_discrete(expand = c(0.6,0))
```

Height_cm (of main stalk)

```{r, echo = FALSE}
ggplot(data = Means_2021[["height_means_21"]], aes(x = Treatment, y = Mean)) +
  geom_point(aes(col=Population, shape = Population), size = 3, position = position_dodge(width = 0.3), show.legend = TRUE) +
  geom_errorbar(aes(ymin=LL_95, ymax = UL_95, col = Population), width = 0.3, size = 0.8, position = position_dodge(width = 0.3), show.legend = TRUE)+
  geom_line(aes(group = Population, col = Population), size = 1, position = position_dodge(width = 0.3))+ 
  labs(y="Height (cm)", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_x_discrete(expand = c(0.6,0))
```

## fitness + fitness related

FruitCount - BL

```{r, echo = FALSE}
# Basia Fruit Count
fruits <- ggplot(data = Means_2021[["fruit_means_21"]], aes(x = Treatment, y = Mean)) +
  geom_point(aes(col=Population, shape = Population), size = 3, position = position_dodge(width = 0.3), show.legend = TRUE) +
  geom_errorbar(aes(ymin=LL_95, ymax = UL_95, col = Population), width = 0.3, size = 0.8, position = position_dodge(width = 0.3), show.legend = TRUE)+
  geom_line(aes(group = Population, col = Population), size = 1, position = position_dodge(width = 0.3))+ 
  labs(y="Fruit Count", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_x_discrete(expand = c(0.6,0))

fruits
```

AvgSeedWt

```{r, echo = FALSE}
seed_wt <- ggplot(data = Means_2021[["AvSeedWt_means_21"]], aes(x = Treatment, y = Mean)) +
  geom_point(aes(col=Population, shape = Population), size = 3, position = position_dodge(width = 0.3), show.legend = TRUE) +
  geom_errorbar(aes(ymin=LL_95, ymax = UL_95, col = Population), width = 0.3, size = 0.8, position = position_dodge(width = 0.3), show.legend = TRUE)+
  geom_line(aes(group = Population, col = Population), size = 1, position = position_dodge(width = 0.3))+ 
  labs(y="Seed Weight (mg)", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_x_discrete(expand = c(0.6,0))
seed_wt
```

AvgSeedNum

```{r, echo = FALSE}
seed_per_fruit <- ggplot(data = Means_2021[["seeds_means_21"]], aes(x = Treatment, y = Mean)) +
  geom_point(aes(col=Population, shape = Population), size = 3, position = position_dodge(width = 0.3), show.legend = TRUE) +
  geom_errorbar(aes(ymin=LL_95, ymax = UL_95, col = Population), width = 0.3, size = 0.8, position = position_dodge(width = 0.3), show.legend = TRUE)+
  geom_line(aes(group = Population, col = Population), size = 1, position = position_dodge(width = 0.3))+ 
  labs(y="Seeds per Fruit", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_x_discrete(expand = c(0.6,0))

seed_per_fruit
```

fitness - avg seed num/fruit * fruit num

```{r, echo = FALSE}
fitness <- ggplot(data = Means_2021[["fitness_means_21"]], aes(x = Treatment, y = Mean)) +
  geom_point(aes(col=Population, shape = Population), size = 3, position = position_dodge(width = 0.3), show.legend = TRUE) +
  geom_errorbar(aes(ymin=LL_95, ymax = UL_95, col = Population), width = 0.3, size = 0.8, position = position_dodge(width = 0.3), show.legend = TRUE)+
  geom_line(aes(group = Population, col = Population), size = 1, position = position_dodge(width = 0.3))+ 
  labs(y="Total Fitness", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_x_discrete(expand = c(0.6,0))
fitness

```


# 2022 only

Graphs to make here: \
- LeafNum_Jun6 (if included) \
- LeafNum_Jun13 (if included) \
- LeafNum_bolt \
- AG_Dry_bolt \
- BG_Dry_bolt \
- Root_to_Shoot \
- Stomata_density \

Note: skipping days from emergence to bolting because most of that time is the same treatment.

## Leaf Number
LeafNum_Jun6 

```{r, echo = FALSE}
ggplot(data = Means_2022[["LN_Jun6_means_22"]], aes(x = Treatment, y = Mean)) +
  geom_point(aes(col=Population, shape = Population), size = 3, position = position_dodge(width = 0.3), show.legend = TRUE) +
  geom_errorbar(aes(ymin=LL_95, ymax = UL_95, col = Population), width = 0.3, size = 0.8, position = position_dodge(width = 0.3), show.legend = TRUE)+
  geom_line(aes(group = Population, col = Population), size = 1, linetype = "dotted", position = position_dodge(width = 0.3))+ 
  labs(y="Leaf Number 1 week post vern", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_x_discrete(expand = c(0.6,0))
```

LeafNum_Jun13 
```{r, echo = FALSE}
ggplot(data = Means_2022[["LN_Jun13_means_22"]], aes(x = Treatment, y = Mean)) +
  geom_point(aes(col=Population, shape = Population), size = 3, position = position_dodge(width = 0.3), show.legend = TRUE) +
  geom_errorbar(aes(ymin=LL_95, ymax = UL_95, col = Population), width = 0.3, size = 0.8, position = position_dodge(width = 0.3), show.legend = TRUE)+
  geom_line(aes(group = Population, col = Population), size = 1, linetype = "dotted", position = position_dodge(width = 0.3))+ 
  labs(y="Leaf Number 2 weeks post vern", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_x_discrete(expand = c(0.6,0))
```

LeafNum_bolt

```{r, echo = FALSE}
ln_bolt <- ggplot(data = Means_2022[["LN_bolt_means_22"]], aes(x = Treatment, y = Mean)) +
  geom_point(aes(col=Population, shape = Population), size = 3, position = position_dodge(width = 0.3), show.legend = TRUE) +
  geom_errorbar(aes(ymin=LL_95, ymax = UL_95, col = Population), width = 0.3, size = 0.8, position = position_dodge(width = 0.3), show.legend = TRUE)+
  geom_line(aes(group = Population, col = Population), size = 1, linetype = "dotted", position = position_dodge(width = 0.3))+ 
  labs(y="Leaf Number at Bolting", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_x_discrete(expand = c(0.6,0))

ln_bolt
```

Create an extra figure to look at growth curves during this time. I want one line per population and 4 x axis values (4 weeks, post vern, 2 weeks post vern) with the y axis being leaf number.

```{r}
# format a dataframe
tmp_mean <- Dat_2022 %>%
  dplyr::group_by(Population) %>%
  dplyr::group_by(Treatment, .add = TRUE) %>%
  summarize(across(.col = c(LeafNumber_PreVern:LeafNumber_Jun13, LeafNumber_Total), .fns = list(mean = ~mean(.x, na.rm = TRUE, .group = "keep"))))
# pivot
tmp_mean <- pivot_longer(tmp_mean, cols = !(c(Population, Treatment)), names_to = "Time", values_to = "mean")
tmp_mean$Time <- factor(rep(c("Week5", "Week9", "Week10", "Bolting"), times = 4), levels = c("Week5", "Week9", "Week10", "Bolting"))

tmp_ci <- Dat_2022 %>%
  dplyr::group_by(Population) %>%
  dplyr::group_by(Treatment, .add = TRUE) %>%
  summarize(across(.col = c(LeafNumber_PreVern:LeafNumber_Jun13, LeafNumber_Total), .fns = list(ci_95 = ~conf_int(.x))))

# pivot
tmp_ci <- pivot_longer(tmp_ci, cols = !(c(Population, Treatment)), names_to = "Time", values_to = "ci")
tmp_ci$Time <- factor(rep(c("Week5", "Week9", "Week10", "Bolting"), times = 4), levels = c("Week5", "Week9", "Week10", "Bolting"))

forplot3_2022 <- merge(tmp_mean, tmp_ci)

# plot
ggplot(data = forplot3_2022, aes(x = Time, y = mean)) +
  geom_point(aes(col=Population, shape = Population), size = 2, show.legend = TRUE) +
  #geom_errorbar(aes(ymin=mean-ci, ymax = mean + ci, col = Population), width = 0.1, size = 0.8, show.legend = TRUE)+
  geom_line(aes(group = Treatment:Population, col = Population, linetype = Treatment), size = 1)+ 
  labs(y="Leaf Number", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_x_discrete(expand = c(0.2,0))
```


## Biomass
AG_Dry_bolt

```{r, echo = FALSE}
ros_2022 <- ggplot(data = Means_2022[["AG_means_22"]], aes(x = Treatment, y = Mean)) +
  geom_point(aes(col=Population, shape = Population), size = 3, position = position_dodge(width = 0.3), show.legend = TRUE) +
  geom_errorbar(aes(ymin=LL_95, ymax = UL_95, col = Population), width = 0.3, size = 0.8, position = position_dodge(width = 0.3), show.legend = TRUE)+
  geom_line(aes(group = Population, col = Population), size = 1, linetype = "dotted", position = position_dodge(width = 0.3))+ 
  labs(y="Shoot Biomass (g)", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_x_discrete(expand = c(0.6,0))

ros_2022

# what are the percent decreases for each pop?
# Italy
(0.1654970 - 0.2128705)/0.2128705
# Sweden
(0.4577489 - 0.4678267) / 0.4678267

```

BG_Dry_bolt

```{r, echo = FALSE}
bg <- ggplot(data = Means_2022[["BG_means_22"]], aes(x = Treatment, y = Mean)) +
  geom_point(aes(col=Population, shape = Population), size = 3, position = position_dodge(width = 0.3), show.legend = TRUE) +
  geom_errorbar(aes(ymin=LL_95, ymax = UL_95, col = Population), width = 0.3, size = 0.8, position = position_dodge(width = 0.3), show.legend = TRUE)+
  geom_line(aes(group = Population, col = Population), size = 1, linetype = "dotted", position = position_dodge(width = 0.3))+ 
  labs(y="Root Biomass (g)", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_x_discrete(expand = c(0.6,0))
bg

# what are the percent decreases?
#Italy
(0.02796990 - 0.03572211 )/ 0.03572211 
# Sweden
(0.05366088 - 0.05903125 ) / 0.05903125 
```

Root_to_Shoot

```{r, echo = FALSE}
rs <- ggplot(data = Means_2022[["RS_means_22"]], aes(x = Treatment, y = Mean)) +
  geom_point(aes(col=Population, shape = Population), size = 3, position = position_dodge(width = 0.3), show.legend = TRUE) +
  geom_errorbar(aes(ymin=LL_95, ymax = UL_95, col = Population), width = 0.3, size = 0.8, position = position_dodge(width = 0.3), show.legend = TRUE)+
  geom_line(aes(group = Population, col = Population), size = 1, linetype = "dotted", position = position_dodge(width = 0.3))+ 
  labs(y="Root:Shoot Ratio", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_x_discrete(expand = c(0.6,0))

rs

# old fraction label: expression(paste(frac("Root", "Shoot"), "   Biomass"))
```

## Stomata
Stomata Density

```{r, echo = FALSE}
stoden <- ggplot(data = Means_2022[["sto_den_means_22"]], aes(x = Treatment, y = Mean)) +
  geom_point(aes(col=Population, shape = Population), size = 3, position = position_dodge(width = 0.3), show.legend = TRUE) +
  geom_errorbar(aes(ymin=LL_95, ymax = UL_95, col = Population), width = 0.3, size = 0.8, position = position_dodge(width = 0.3), show.legend = TRUE)+
  geom_line(aes(group = Population, col = Population), size = 1, linetype = "dotted", position = position_dodge(width = 0.3))+ 
  labs(y=expression(paste("Stomatal Density  ", (mm^-2))), 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_x_discrete(expand = c(0.6,0))

stoden
```

# both
Steps to be completed here.
Subset 2021 dataset. add 2021 column. Subset 2022 dataset. add 2022 column. merge - rbind for each trait. format. plot.

Reaction Norms to create here: \
- Emergence \
- EmergenceToBolting \
- FreshWt \
- HydWt \
- DryWt \
- LeafArea \
- LeafPerimeter \
- SLA \
- LDMC \
- RWC \
- LeafNum_4weeks \

```{r}
# subset 2021 dataset
both_exp <- c("emergence_means_21", "bolting_means_21", "fresh_means_21", "sat_means_21", "dry_means_21", "area_means_21", "per_means_21", "SLA_means_21", "LDMC_means_21", "RWC_means_21", "LN_PreVern_means_21")
both_2021 <- Means_2021[both_exp]
both_2021 <- lapply(both_2021, cbind, Exp = as.factor(c("2021")))

# subset 2022 dataset
both_exp2 <- c("emergence_means_22", "bolting_means_22", "fresh_means_22", "hyd_means_22", "dry_means_22", "area_means_22", "per_means_22", "sla_means_22", "ldmc_means_22", "rwc_means_22", "LN_PreVern_means_22")
both_2022 <- Means_2022[both_exp2]
# change B and R to Belm and Roda
for (i in 1:length(both_2022)){
  both_2022[[i]]$Population <- as.character(both_2022[[i]]$Population)
  both_2022[[i]][both_2022[[i]]$Population == "B", "Population"] <- "BELM"
  both_2022[[i]][both_2022[[i]]$Population == "R", "Population"] <- "RODA"
  both_2022[[i]]$Population <- as.factor(both_2022[[i]]$Population)
}
both_2022 <- lapply(both_2022, cbind, Exp = as.factor(c("2022")))


# merge - do this within a loop
mylist.names <- c("emergence", "bolting", "fresh", "hyd", "dry", "area", "per", "sla", "ldmc", "rwc", "LN")
both_dfs <-  vector("list", length(mylist.names))
names(both_dfs) <- mylist.names

for (i in 1:length(both_dfs)){
  both_dfs[[i]] <- bind_rows(both_2021[[i]], both_2022[[i]])
}
```


## Phenology

Emergence. days to emergence is shown for both but should be interpreted with caution. In the 2022 experiment, all plants were put in the same conditions so there may be genetic differences but treatment differences would be due to chamber placement not due to temperature or water availability.

```{r, echo = FALSE}
ggplot(data = both_dfs[["emergence"]], aes(x = Treatment, y = Mean)) +
  geom_point(aes(group = Exp:Population, col=Population, shape = Population), size = 3, position = position_dodge(width = 0.3), show.legend = FALSE) +
  geom_errorbar(aes(group = Exp:Population, ymin=LL_95, ymax = UL_95, col = Population), width = 0.3, size = 0.8, position = position_dodge(width = 0.3), show.legend = FALSE)+
  geom_line(aes(group = Exp:Population, col = Population, linetype = Exp), size = 1, position = position_dodge(width = 0.3), show.legend = TRUE)+
  labs(y="Days To Emergence", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_linetype_manual(name = "Experiment",
                        labels = c("2021", "2022"),
                        values = c("solid", "dotted"))+
  scale_x_discrete(expand = c(0.6,0))

```


EmergenceToBolting. While this was measured in both experiments, it is not comparable because in 2022 the conditions were the same for all plants for the first 9 weeks, which is more of the time until the plants start to bolt. Emergence to bolting was moved to an only 2021 trait and is not included in the manuscript.

```{r, echo = FALSE, eval = TRUE}
detb <- ggplot(data = both_dfs[["bolting"]], aes(x = Treatment, y = Mean)) +
  geom_point(aes(group = Exp:Population, col=Population, shape = Population), size = 3, position = position_dodge2(width = 0.3), show.legend = FALSE) +
  geom_errorbar(aes(ymin=LL_95, ymax = UL_95,group = Exp:Population, col = Population), width = 0.3, size = 0.8, position = position_dodge2(width = 0.3), show.legend = FALSE)+
  geom_line(aes(group = Exp:Population,col = Population, linetype = Exp), size = 1, position = position_dodge2(width = 0.3), show.legend = TRUE)+
  labs(y="Days to Bolting", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_linetype_manual(name = "Experiment",
                        labels = c("2021", "2022"),
                        values = c("solid", "dotted"))+
  scale_x_discrete(expand = c(0.6,0))
detb
```


## Single Leaf Individual Traits
FreshWt

```{r, echo = FALSE}
freshwt <- ggplot(data = both_dfs[["fresh"]], aes(x = Treatment, y = Mean)) +
  geom_point(aes(group = Exp:Population, col=Population, shape = Population), size = 3, position = position_dodge(width = 0.3), show.legend = FALSE) +
  geom_errorbar(aes(group = Exp:Population, ymin=LL_95, ymax = UL_95, col = Population), width = 0.3, size = 0.8, position = position_dodge(width = 0.3), show.legend = FALSE)+
  geom_line(aes(group = Exp:Population, col = Population, linetype = Exp), size = 1, position = position_dodge(width = 0.3), show.legend = TRUE)+
  labs(y="Fresh Weight (g)", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_linetype_manual(name = "Experiment",
                        labels = c("2021", "2022"),
                        values = c("solid", "dotted"))+
  scale_x_discrete(expand = c(0.6,0))
```


HydWt

```{r, echo = FALSE}
hydwt <- ggplot(data = both_dfs[["hyd"]], aes(x = Treatment, y = Mean)) +
  geom_point(aes(group = Exp:Population, col=Population, shape = Population), size = 3, position = position_dodge(width = 0.3), show.legend = FALSE) +
  geom_errorbar(aes(group = Exp:Population, ymin=LL_95, ymax = UL_95, col = Population), width = 0.3, size = 0.8, position = position_dodge(width = 0.3), show.legend = FALSE)+
  geom_line(aes(group = Exp:Population, col = Population, linetype = Exp), size = 1, position = position_dodge(width = 0.3), show.legend = TRUE)+
  labs(y="Rehydrated Weight (g)", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_linetype_manual(name = "Experiment",
                        labels = c("2021", "2022"),
                        values = c("solid", "dotted"))+
  scale_x_discrete(expand = c(0.6,0))
```


DryWt

```{r, echo = FALSE}
drywt <- ggplot(data = both_dfs[["dry"]], aes(x = Treatment, y = Mean)) +
  geom_point(aes(group = Exp:Population, col=Population, shape = Population), size = 3, position = position_dodge(width = 0.3), show.legend = FALSE) +
  geom_errorbar(aes(group = Exp:Population, ymin=LL_95, ymax = UL_95, col = Population), width = 0.3, size = 0.8, position = position_dodge(width = 0.3), show.legend = FALSE)+
  geom_line(aes(group = Exp:Population, col = Population, linetype = Exp), size = 1, position = position_dodge(width = 0.3), show.legend = TRUE)+
  labs(y="Dried Weight (g)", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_linetype_manual(name = "Experiment",
                        labels = c("2021", "2022"),
                        values = c("solid", "dotted"))+
  scale_x_discrete(expand = c(0.6,0))
```


LeafArea

```{r, echo = FALSE}
area <- ggplot(data = both_dfs[["area"]], aes(x = Treatment, y = Mean)) +
  geom_point(aes(group = Exp:Population, col=Population, shape = Population), size = 3, position = position_dodge(width = 0.3), show.legend = FALSE) +
  geom_errorbar(aes(group = Exp:Population, ymin=LL_95, ymax = UL_95, col = Population), width = 0.3, size = 0.8, position = position_dodge(width = 0.3), show.legend = FALSE)+
  geom_line(aes(group = Exp:Population, col = Population, linetype = Exp), size = 1, position = position_dodge(width = 0.3), show.legend = TRUE)+
  labs(y=expression(paste("Leaf Area ", (cm^2))), 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_linetype_manual(name = "Experiment",
                        labels = c("2021", "2022"),
                        values = c("solid", "dotted"))+
  scale_x_discrete(expand = c(0.6,0))
```


LeafPerimeter

```{r, echo = FALSE}
perim <- ggplot(data = both_dfs[["per"]], aes(x = Treatment, y = Mean)) +
  geom_point(aes(group = Exp:Population, col=Population, shape = Population), size = 3, position = position_dodge(width = 0.3), show.legend = FALSE) +
  geom_errorbar(aes(group = Exp:Population, ymin=LL_95, ymax = UL_95, col = Population), width = 0.3, size = 0.8, position = position_dodge(width = 0.3), show.legend = FALSE)+
  geom_line(aes(group = Exp:Population, col = Population, linetype = Exp), size = 1, position = position_dodge(width = 0.3), show.legend = TRUE)+
  labs(y="Leaf Perimeter (cm)", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_linetype_manual(name = "Experiment",
                        labels = c("2021", "2022"),
                        values = c("solid", "dotted"))+
  scale_x_discrete(expand = c(0.6,0))
```



## Single Leaf Calculated Traits
Specific Leaf Area

```{r, echo = FALSE}
sla <- ggplot(data = both_dfs[["sla"]], aes(x = Treatment, y = Mean)) +
  geom_point(aes(group = Exp:Population, col=Population, shape = Population), size = 3, position = position_dodge(width = 0.3), show.legend = FALSE) +
  geom_errorbar(aes(group = Exp:Population, ymin=LL_95, ymax = UL_95, col = Population), width = 0.3, size = 0.8, position = position_dodge(width = 0.3), show.legend = FALSE)+
  geom_line(aes(group = Exp:Population, col = Population, linetype = Exp), size = 1, position = position_dodge(width = 0.3), show.legend = TRUE)+
  labs(y=expression(paste("Specific Leaf Area ", (cm^2 / g))),
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_linetype_manual(name = "Experiment",
                        labels = c("2021", "2022"),
                        values = c("solid", "dotted"))+
  scale_x_discrete(expand = c(0.6,0))


# just 2021 SLA for Emily's tenure talk
ggplot(data = both_2021[["SLA_means_21"]], aes(x = Treatment, y = Mean)) +
  geom_point(aes(group = Population, col=Population), size = 2, show.legend = FALSE) +
  geom_errorbar(aes(group = Population, ymin=LL_95, ymax = UL_95, col = Population), width = 0.3, size = 0.8, show.legend = FALSE)+
  geom_line(aes(group = Population, col = Population), size = 1, show.legend = TRUE)+
  labs(y="Specific Leaf Area (cm^2 / g)", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_x_discrete(expand = c(0.6,0))
ggsave(filename = "figures/SLA_2021.png", dpi=300)
```


Leaf Dry Matter Content

```{r, echo = FALSE}
ldmc <- ggplot(data = both_dfs[["ldmc"]], aes(x = Treatment, y = Mean)) +
  geom_point(aes(group = Exp:Population, col=Population, shape = Population), size = 3, position = position_dodge(width = 0.3), show.legend = FALSE) +
  geom_errorbar(aes(group = Exp:Population, ymin=LL_95, ymax = UL_95, col = Population), width = 0.3, size = 0.8, position = position_dodge(width = 0.3), show.legend = FALSE)+
  geom_line(aes(group = Exp:Population, col = Population, linetype = Exp), size = 1, position = position_dodge(width = 0.3), show.legend = TRUE)+
  labs(y="Leaf Dry Matter Content (%)", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_linetype_manual(name = "Experiment",
                        labels = c("2021", "2022"),
                        values = c("solid", "dotted"))+
  scale_x_discrete(expand = c(0.6,0))
ldmc
```


Relative Water Content

```{r, echo = FALSE}
rwc <- ggplot(data = both_dfs[["rwc"]], aes(x = Treatment, y = Mean)) +
  geom_point(aes(group = Exp:Population, col=Population, shape = Population), size = 3, position = position_dodge(width = 0.3), show.legend = FALSE) +
  geom_errorbar(aes(group = Exp:Population, ymin=LL_95, ymax = UL_95, col = Population), width = 0.3, size = 0.8, position = position_dodge(width = 0.3), show.legend = FALSE)+
  geom_line(aes(group = Exp:Population, col = Population, linetype = Exp), size = 0.8, position = position_dodge(width = 0.3), show.legend = TRUE)+
  labs(y="Relative Water Content (%)", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_linetype_manual(name = "Experiment",
                        labels = c("2021", "2022"),
                        values = c("solid", "dotted"))+
  scale_x_discrete(expand = c(0.6,0))
rwc
```

## Leaf Number
LeafNum_4weeks

```{r, echo = FALSE}
ggplot(data = both_dfs[["LN"]], aes(x = Treatment, y = Mean)) +
  geom_point(aes(group = Exp:Population, col=Population, shape = Population), size = 3, position = position_dodge(width = 0.3), show.legend = FALSE) +
  geom_errorbar(aes(group = Exp:Population, ymin=LL_95, ymax = UL_95, col = Population), width = 0.3, size = 0.8, position = position_dodge(width = 0.3), show.legend = FALSE)+
  geom_line(aes(group = Exp:Population, col = Population, linetype = Exp), size = 1, position = position_dodge(width = 0.3), show.legend = TRUE)+
  labs(y="Leaf Number (4 weeks post planting)", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_linetype_manual(name = "Experiment",
                        labels = c("2021", "2022"),
                        values = c("solid", "dotted"))+
  scale_x_discrete(expand = c(0.6,0))
```

# Manuscript Figures
creating multipanel figures for the manuscript. 
Misc updates to restructure and change figure numbers made 4/17/2025, 5/1/2025, and 7/1/2025. \
Draft 2 figures updated 7/3/2025 \
Draft 3 figures updated 7/30/2025 \

```{r}
fig1 <- ggarrange(rwc+rremove("xlab"), sla+rremove("xlab"), detf, stoden,
                  labels = c("A", "B", "C", "D"),
                  ncol = 2, nrow = 2,
                  common.legend = TRUE, # created based on the first plot
                  legend = "none",
                  align = "hv"
)
#label.x = c(0, -0.1, -0.1, -0.1, -0.1, 0)
#vjust = 00
fig1

fig2 <- ggarrange(rs, ros_2022, bg,
                  labels = c("A", "B", "C"),
                  ncol = 3, nrow = 1,
                  common.legend = TRUE,
                  legend = "none",
                  align = "hv")
fig2

fitness_fig <- ggarrange(fruits, seed_per_fruit, fitness,
                  labels = c("A", "B", "C"),
                  ncol = 3, nrow = 1,
                  common.legend = TRUE,
                  legend = "none",
                  align = "h"
)
fitness_fig
```


# supplemental figures

Figure s1 - single leaf traits \
```{r}
figs1 <- ggarrange(ldmc+rremove("xlab"),area+rremove("xlab"), perim+rremove("xlab"), freshwt, hydwt, drywt, 
                  labels = c("A", "B", "C", "D", "E", "F"),
                  ncol = 3, nrow = 2,
                  common.legend = TRUE,
                  legend = "none",
                  align = "h"
                  )
figs1
```

Figure s2 - biomass + fitness \
```{r}
figs2 <- ggarrange(ros_2021+rremove("xlab"), repro+rremove("xlab"), ln_harv+rremove("xlab"),  ln_bolt, rr, seed_wt,
                  labels = c("A", "B", "C", "D", "E", "F"),
                  ncol = 3, nrow = 2,
                  common.legend = TRUE,
                  legend = "none",
                  align = "h"
                  )
figs2
```


Create Legend for main figures

For the traits figure, I need the colors with the shaped points and I need the line types with the years of the experiment
```{r}
legend_points <- data.frame(xvals = c(1.1, 1.1), yvals = c(4, 3), text = c("Italy", "Sweden"), Pop = as.factor(c("IT", "SW")))
#legend_lines <- c(xvals = c(1, 1), yvals = c(2, 1), text = c("2021", "2022"))

# legend to take up space of a plot
ggplot()+
  geom_point(data = legend_points, aes(x = xvals, y = yvals, col = Pop, shape = Pop), size = 4, show.legend = FALSE)+
  geom_segment(aes(x = 1.09, y = 2, xend = 1.11, yend = 2), linetype = "solid", linewidth = 1.25)+
  geom_segment(aes(x = 1.09, y = 1, xend = 1.11, yend = 1), linetype = "dotted", linewidth = 1.25)+
  geom_text(aes(x = c(1.11, 1.11, 1.115, 1.115), y = c(4, 3, 2, 1), label = c("Italy", "Sweden", "2021", "2022")))+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  xlim(c(1.087, 1.12))+
  theme_void()
  #theme(plot.margin = margin(t = 0, r = 0.5, b = 0.075, l=0, unit = "in"))

# top legend
legend_points2 <- data.frame(xvals = c(1, 1.5), yvals = c(1, 1), text = c("Italy", "Sweden"), Pop = as.factor(c("IT", "SW")))
legend_fig <- ggplot()+
  geom_point(data = legend_points2, aes(x = xvals, y = yvals, col = Pop, shape = Pop), size = 4, show.legend = FALSE)+
  geom_segment(aes(x = 2.2, y = 1, xend = 2.7, yend = 1), linetype = "solid", linewidth = 1.25)+
  geom_segment(aes(x = 3.2, y = 1, xend = 3.7, yend = 1), linetype = "dotted", linewidth = 1.25)+
  geom_text(aes(x = c(1.2, 1.8, 2.9, 3.9), y = c(1, 1, 1, 1), label = c("Italy", "Sweden", "2021", "2022")))+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  xlim(c(0.087, 4.12))+
  theme_void()
legend_fig
```



##Save figures
```{r}

# trait figs
ggsave(fig1, filename = "figures/fig1.jpg", width = 6, height = 8, units = "in", dpi = 600)
ggsave(fig2, filename = "figures/fig2.jpg", width = 9, height = 4, units = "in", dpi = 600)

# fitness fig
ggsave(fitness_fig, filename = "figures/fig3.jpg", width = 9, height = 4, units = "in", dpi = 600)

# legend_fig
ggsave(legend_fig, filename = "figures/legend_fig.jpg", width = 6, height = 0.5, units = "in", dpi = 600)

#figs1
ggsave(figs1, filename = "figures/figs1.jpg", width = 9, height = 8, units = "in", dpi = 600)

#figs2
ggsave(figs2, filename = "figures/figs2.jpg", width = 9, height = 8, units = "in", dpi = 600)


```

