---
title: "03_Figures"
author: "Sophie Buysse"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# description

This code creates figures from both experiments. First, there are figures specific to each experiment showing phenology trends throughout the growing period. Then, there are reaction norms for the traits only collected in one experiment. Finally, there are reaction norms for the traits that were measured in both experiments with lines from both on the same graph. Below is the coloring guide to be consistent between all experiments and be color-blind accessible.

```{r, eval = FALSE, echo = FALSE}
#Full Okabe Ito color palette:
c("#E69F00",  "#56B4E9", "#0072B2",  "#D55E00","#CC79A7",  "#999999", "#000000", "#009E73", "#F5C710")
```

Sweden: #009E73 - a green color \
Italy: #CC79A7 - light pink \
Current: x axis, if needed #56B4E9 - a light blue or a solid line \
Future: x axis, if needed #D55E00 - a dark orange or a dashed line \
2021: solid lines \
2022: dashed lines \

# Prep code

load packages

```{r}
library(dplyr)
library(ggplot2)
library(tidyr)

# make function
std_mean <- function(x){
  sd(x, na.rm = TRUE)/sqrt(length(x))
}

# Confidence interval multiplier for standard error
    # Calculate t-statistic for confidence interval: 
    # e.g., if conf.interval is .95, use .975 (above/below), and use df=N-1

conf_int <- function(x, conf.interval = 0.95){
  ciMult <- qt((1+conf.interval)/2, length(x)-1)
  ci <- std_mean(x) * ciMult
  return(ci)
}
```

Set theme for plots
```{r}
theme_set(theme_classic())
theme_update(legend.title = element_text(family = "serif", color = "black", size = 20),
    legend.text = element_text(family = "serif", color = "black", size = 20),
    axis.title = element_text(family = "serif", color = "black", size = 20),
    axis.text = element_text(family = "serif", color = "black", size = 20),
    axis.line = element_line(colour = 'black', linewidth = 1),
    axis.ticks = element_line(color = 'black', linewidth = 1),
    axis.ticks.length=unit(.15, "cm"),
    legend.spacing.y = unit(0.03, "cm"),
    panel.border = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank()
    )
    
```



# read in data

Read in the data needed for all plotting
```{r}
# 2021
Dat_2021 <- read.csv("data/CleanData_2021.csv")
# set factors
Dat_2021 <- Dat_2021 %>% dplyr::mutate(
  Population = as.factor(Population),
  Line = as.factor(Line),
  Treatment = as.factor(Treatment),
  Soil.Mix.Batch.Number.x = as.factor(Soil.Mix.Batch.Number.x),
  BranchStructure = as.factor(BranchStructure),
  DoneFlwr = as.factor(DoneFlwr)
)
# subset to two treatments
Dat_2021 <- Dat_2021[which(Dat_2021$Treatment == "Current"| Dat_2021$Treatment == "Future") , ]

Sum_2021 <- read.csv("data/PopMeansByTreatment_2021.csv")
# subset to only two treatments
Sum_2021 <- Sum_2021[ , c(1, 2, 4, 5, 7)]

# 2022
Dat_2022 <- read.csv("data/CleanData_2022.csv")
# make some things factors
Dat_2022$Treatment <- as.factor(Dat_2022$Treatment)
Dat_2022$Population <- as.factor(Dat_2022$Population)
Dat_2022$PotID <- as.factor(Dat_2022$PotID)
Dat_2022$Flat <- as.factor(Dat_2022$Flat)
Dat_2022$Chamber <- as.factor(Dat_2022$Chamber)
Dat_2022$Transplanted <- as.factor(Dat_2022$Transplanted)

Sum_2022 <- read.csv("data/PopMeansByTreatment_2022.csv")
```

# phenology
Graphs showing general phenology trends. The goal of these is a quick comparison, likely will not be in the manuscript.

2021
```{r, echo = FALSE}
ggplot(data = Dat_2021)+
  geom_density(aes(Emergence, col = Population, linetype = Treatment), size = 1.25)+
  geom_density(aes((Emergence + DayToBolt), col = Population, linetype = Treatment), size = 1.25)+
  geom_density(aes((Emergence + DayToBolt + DayToFlwr), col = Population, linetype = Treatment), size = 1.25)+
  labs(x = "Days Since Planting")+
  scale_color_manual(name = "Population",
                     labels = c("Italy", "Sweden"),
                     values = c("#CC79A7", "#009E73"))+
  theme_classic() +
  theme(
    legend.title = element_text(color = "black", size = 16),
    legend.text = element_text(color = "black", size = 16),
    axis.title = element_text(color = "black", size = 16),
    axis.text = element_text(color = "black", size = 16),
    legend.spacing.y = unit(0.03, "cm"),
    panel.border = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank()
  )

```

2022 - note: plants that died or were transplanted are not shown here because they may have an emergence date and no bolting date (if died) or a bolting date and no emergence date (if transplanted)
```{r, echo = FALSE}
ggplot(data = Dat_2022)+
  geom_density(aes(DaysToEmergence, col = Population, linetype = Treatment), size = 1.25)+
  geom_density(aes((DaysToEmergence + EmergenceToBolting), col = Population, linetype = Treatment), size = 1.25)+
  labs(x = "Days Since Planting")+
  scale_color_manual(name = "Population",
                     labels = c("Italy", "Sweden"),
                     values = c("#CC79A7", "#009E73"))+
  theme_classic() +
  theme(
    legend.title = element_text(color = "black", size = 16),
    legend.text = element_text(color = "black", size = 16),
    axis.title = element_text(color = "black", size = 16),
    axis.text = element_text(color = "black", size = 16),
    legend.spacing.y = unit(0.03, "cm"),
    panel.border = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank()
  )

```


Takeaways: phenology is quicker for both populations in the 2022 experiment. In 2022, Italy bolts around 70 days post planting but it is closer to 85 days in 2021. Similarly, in 2021 Sweden takes over 100 days to bolt but it bolted at about 90 days in 2022. There were some experimental differences (like light and the chambers used and the conditions through vernalization) that could contribute to this difference. 

# 2021 only

Reaction norms to make here: \
- DaysBoltingToFlwr \
- LeafNum_duringvern (if included) \
- LeafNum_Oct4 (if included) \
- RosetteLeafNum_harvest \
- RosetteDry_harvest \
- DryReproG \
- Repro_to_Ros \
- LatBranches \
- PrimaryStalks \
- Height_cm \
- IJ_FruitCount (alsokept other fruit count in the forplot table just cause) \
- AG_biomass (Ros+Repro) \
- fitness \
- AvgSeedWt \
- AvgSeedNum \
- NumFlwrLeft (no model as of 7/31/2024, not included here, not measured for everything) \

Subset to only that data from the summary table + format
```{r}
forplot_2021 <- data.frame(t(Sum_2021))
colnames(forplot_2021) <- forplot_2021[1, ]
forplot_2021 <- forplot_2021[c(2:5), ]
forplot_2021 <- forplot_2021 %>%
  select(Population, Treatment, DayToFlwr_mean, DayToFlwr_ci_95, LeafNum_08262021_mean:RosetteLeafNum_10042021_ci_95, RosetteLeafNum_mean:DryReproG_ci_95, Repro_to_Ros_mean:IJ_FruitCount_ci_95, AG_biomass_mean:fitness_ci_95) %>%
  mutate(across(Population:Treatment, as.factor),
         across(DayToFlwr_mean:fitness_ci_95, as.numeric))
```

## Phenology
DaysBoltingToFlwr


```{r, echo = FALSE}
ggplot(data = forplot_2021, aes(x = Treatment, y = DayToFlwr_mean)) +
  geom_point(aes(col=Population), size = 4, show.legend = TRUE) +
  geom_errorbar(aes(ymin=DayToFlwr_mean-DayToFlwr_ci_95, ymax = DayToFlwr_mean + DayToFlwr_ci_95, col = Population), width = 0.1, size = 1, show.legend = TRUE)+
  geom_line(aes(group = Population, col = Population), size = 1.5)+ 
  labs(y="Days Bolting to Flowering", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_x_discrete(expand = c(0.6,0))+
  expand_limits(y=0)
```


## Leaf Number
LeafNum_duringvern (if included)

```{r, echo = FALSE}
ggplot(data = forplot_2021, aes(x = Treatment, y = LeafNum_08262021_mean)) +
  geom_point(aes(col=Population), size = 4, show.legend = TRUE) +
  geom_errorbar(aes(ymin=LeafNum_08262021_mean-LeafNum_08262021_ci_95, ymax = LeafNum_08262021_mean + LeafNum_08262021_ci_95, col = Population), width = 0.1, size = 1, show.legend = TRUE)+
  geom_line(aes(group = Population, col = Population), size = 1.5)+ 
  labs(y="Leaf Number: during vern", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_x_discrete(expand = c(0.6,0))+
  expand_limits(y=0)
```

LeafNum_Oct4 (if included)

```{r, echo = FALSE}
ggplot(data = forplot_2021, aes(x = Treatment, y = RosetteLeafNum_10042021_mean)) +
  geom_point(aes(col=Population), size = 4, show.legend = TRUE) +
  geom_errorbar(aes(ymin=RosetteLeafNum_10042021_mean - RosetteLeafNum_10042021_ci_95, ymax = RosetteLeafNum_10042021_mean + RosetteLeafNum_10042021_ci_95, col = Population), width = 0.1, size = 1, show.legend = TRUE)+
  geom_line(aes(group = Population, col = Population), size = 1.5)+ 
  labs(y="Leaf Number: 6 weeks post vern", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_x_discrete(expand = c(0.6,0))+
  expand_limits(y=0)
```

RosetteLeafNum_harvest

```{r, echo = FALSE}
ggplot(data = forplot_2021, aes(x = Treatment, y = RosetteLeafNum_mean)) +
  geom_point(aes(col=Population), size = 4, show.legend = TRUE) +
  geom_errorbar(aes(ymin=RosetteLeafNum_mean-RosetteLeafNum_ci_95, ymax = RosetteLeafNum_mean + RosetteLeafNum_ci_95, col = Population), width = 0.1, size = 1, show.legend = TRUE)+
  geom_line(aes(group = Population, col = Population), size = 1.5)+ 
  labs(y="Leaf Number: at harvest", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_x_discrete(expand = c(0.6,0))+
  expand_limits(y=0)
```

Create an extra figure to look at growth curves during this time. I want one line per population and 4 x axis values with the y axis being leaf number.

```{r}
# format a dataframe
tmp_mean_21 <- Dat_2021 %>%
  dplyr::group_by(Population) %>%
  dplyr::group_by(Treatment, .add = TRUE) %>%
  summarize(across(.col = c(LeafNum_.07222021, LeafNum_08262021, RosetteLeafNum_10042021, RosetteLeafNum), .fns = list(mean = ~mean(.x, na.rm = TRUE, .group = "keep"))))
# pivot
tmp_mean_21 <- pivot_longer(tmp_mean_21, cols = !(c(Population, Treatment)), names_to = "Time", values_to = "mean")
tmp_mean_21$Time <- factor(rep(c("Week5", "Week10", "Week14", "Harvest"), times = 4), levels = c("Week5", "Week10", "Week14", "Harvest"))

tmp_ci_21 <- Dat_2021 %>%
  dplyr::group_by(Population) %>%
  dplyr::group_by(Treatment, .add = TRUE) %>%
  summarize(across(.col = c(LeafNum_.07222021, LeafNum_08262021, RosetteLeafNum_10042021, RosetteLeafNum), .fns = list(ci_95 = ~conf_int(.x))))

# pivot
tmp_ci_21 <- pivot_longer(tmp_ci_21, cols = !(c(Population, Treatment)), names_to = "Time", values_to = "ci")
tmp_ci_21$Time <- factor(rep(c("Week5", "Week10", "Week14", "Harvest"), times = 4), levels = c("Week5", "Week10", "Week14", "Harvest"))

forplot3_2021 <- merge(tmp_mean_21, tmp_ci_21)

# plot
ggplot(data = forplot3_2021, aes(x = Time, y = mean)) +
  geom_point(aes(col=Population), size = 4, show.legend = TRUE) +
  #geom_errorbar(aes(ymin=mean-ci, ymax = mean + ci, col = Population), width = 0.1, size = 1, show.legend = TRUE)+
  geom_line(aes(group = Treatment:Population, col = Population, linetype = Treatment), size = 1.5)+ 
  labs(y="Leaf Number", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_x_discrete(expand = c(0.2,0))+
  expand_limits(y=0)
```

## Plant Structure
RosetteDry_harvest

```{r, echo = FALSE}
ggplot(data = forplot_2021, aes(x = Treatment, y = DryRosetteG_mean)) +
  geom_point(aes(col=Population), size = 4, show.legend = TRUE) +
  geom_errorbar(aes(ymin=DryRosetteG_mean-DryRosetteG_ci_95, ymax = DryRosetteG_mean + DryRosetteG_ci_95, col = Population), width = 0.1, size = 1, show.legend = TRUE)+
  geom_line(aes(group = Population, col = Population), size = 1.5)+ 
  labs(y="Rosette Dry Biomass (g) at harvest", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_x_discrete(expand = c(0.6,0))+
  expand_limits(y=0)
```

DryReproG

```{r, echo = FALSE}
ggplot(data = forplot_2021, aes(x = Treatment, y = DryReproG_mean)) +
  geom_point(aes(col=Population), size = 4, show.legend = TRUE) +
  geom_errorbar(aes(ymin=DryReproG_mean-DryReproG_ci_95, ymax = DryReproG_mean + DryReproG_ci_95, col = Population), width = 0.1, size = 1, show.legend = TRUE)+
  geom_line(aes(group = Population, col = Population), size = 1.5)+ 
  labs(y="Reproductive Dry Biomass (g)", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_x_discrete(expand = c(0.6,0))+
  expand_limits(y=0)
```

AG_biomass (Ros+Repro)

```{r, echo = FALSE}
ggplot(data = forplot_2021, aes(x = Treatment, y = AG_biomass_mean)) +
  geom_point(aes(col=Population), size = 4, show.legend = TRUE) +
  geom_errorbar(aes(ymin=AG_biomass_mean-AG_biomass_ci_95, ymax = AG_biomass_mean + AG_biomass_ci_95, col = Population), width = 0.1, size = 1, show.legend = TRUE)+
  geom_line(aes(group = Population, col = Population), size = 1.5)+ 
  labs(y="Total Above Ground Biomass (g)", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_x_discrete(expand = c(0.6,0))+
  expand_limits(y=0)
```

Repro_to_Ros

```{r, echo = FALSE}
ggplot(data = forplot_2021, aes(x = Treatment, y = Repro_to_Ros_mean)) +
  geom_point(aes(col=Population), size = 4, show.legend = TRUE) +
  geom_errorbar(aes(ymin=Repro_to_Ros_mean-Repro_to_Ros_ci_95, ymax = Repro_to_Ros_mean + Repro_to_Ros_ci_95, col = Population), width = 0.1, size = 1, show.legend = TRUE)+
  geom_line(aes(group = Population, col = Population), size = 1.5)+ 
  labs(y="Reproductive / Rosette Dry Biomass", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_x_discrete(expand = c(0.6,0))+
  expand_limits(y=0)
```

LatBranches

```{r, echo = FALSE}
ggplot(data = forplot_2021, aes(x = Treatment, y = LatBranches_mean)) +
  geom_point(aes(col=Population), size = 4, show.legend = TRUE) +
  geom_errorbar(aes(ymin=LatBranches_mean-LatBranches_ci_95, ymax = LatBranches_mean + LatBranches_ci_95, col = Population), width = 0.1, size = 1, show.legend = TRUE)+
  geom_line(aes(group = Population, col = Population), size = 1.5)+ 
  labs(y="Number of Lateral Branches", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_x_discrete(expand = c(0.6,0))+
  expand_limits(y=0)
```

PrimaryStalks - how many stalks are coming out of the top of the rosette?

```{r, echo = FALSE}
ggplot(data = forplot_2021, aes(x = Treatment, y = PrimaryStalks_mean)) +
  geom_point(aes(col=Population), size = 4, show.legend = TRUE) +
  geom_errorbar(aes(ymin=PrimaryStalks_mean-PrimaryStalks_ci_95, ymax = PrimaryStalks_mean + PrimaryStalks_ci_95, col = Population), width = 0.1, size = 1, show.legend = TRUE)+
  geom_line(aes(group = Population, col = Population), size = 1.5)+ 
  labs(y="Number of Primary Stalks", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_x_discrete(expand = c(0.6,0))+
  expand_limits(y=0)
```

Height_cm (of main stalk)

```{r, echo = FALSE}
ggplot(data = forplot_2021, aes(x = Treatment, y = Height_cm_mean)) +
  geom_point(aes(col=Population), size = 4, show.legend = TRUE) +
  geom_errorbar(aes(ymin=Height_cm_mean-Height_cm_ci_95, ymax = Height_cm_mean + Height_cm_ci_95, col = Population), width = 0.1, size = 1, show.legend = TRUE)+
  geom_line(aes(group = Population, col = Population), size = 1.5)+ 
  labs(y="Height (cm)", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_x_discrete(expand = c(0.6,0))+
  expand_limits(y=0)
```

## fitness + fitness related

IJ_FruitCount (alsokept other fruit count in the forplot table just cause, is the first graph here)

```{r, echo = FALSE}
# hand count - think is wrong
ggplot(data = forplot_2021, aes(x = Treatment, y = TotFruit_mean)) +
  geom_point(aes(col=Population), size = 4, show.legend = TRUE) +
  geom_errorbar(aes(ymin=TotFruit_mean-TotFruit_ci_95, ymax = TotFruit_mean + TotFruit_ci_95, col = Population), width = 0.1, size = 1, show.legend = TRUE)+
  geom_line(aes(group = Population, col = Population), size = 1.5)+ 
  labs(y="Fruit Count (Machine Learning)", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_x_discrete(expand = c(0.6,0))+
  expand_limits(y=0)

# IJ count
ggplot(data = forplot_2021, aes(x = Treatment, y = IJ_FruitCount_mean)) +
  geom_point(aes(col=Population), size = 4, show.legend = TRUE) +
  geom_errorbar(aes(ymin=IJ_FruitCount_mean-IJ_FruitCount_ci_95, ymax = IJ_FruitCount_mean + IJ_FruitCount_ci_95, col = Population), width = 0.1, size = 1, show.legend = TRUE)+
  geom_line(aes(group = Population, col = Population), size = 1.5)+ 
  labs(y="Fruit Count (ImageJ)", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_x_discrete(expand = c(0.6,0))+
  expand_limits(y=0)
```

AvgSeedWt

```{r, echo = FALSE}
ggplot(data = forplot_2021, aes(x = Treatment, y = AvgSeedWt_mean)) +
  geom_point(aes(col=Population), size = 4, show.legend = TRUE) +
  geom_errorbar(aes(ymin=AvgSeedWt_mean-AvgSeedWt_ci_95, ymax = AvgSeedWt_mean + AvgSeedWt_ci_95, col = Population), width = 0.1, size = 1, show.legend = TRUE)+
  geom_line(aes(group = Population, col = Population), size = 1.5)+ 
  labs(y="Average Seed Weight (mg)", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_x_discrete(expand = c(0.6,0))+
  expand_limits(y=0)
```

AvgSeedNum

```{r, echo = FALSE}
ggplot(data = forplot_2021, aes(x = Treatment, y = AvgSeedNum_mean)) +
  geom_point(aes(col=Population), size = 4, show.legend = TRUE) +
  geom_errorbar(aes(ymin=AvgSeedNum_mean-AvgSeedNum_ci_95, ymax = AvgSeedNum_mean + AvgSeedNum_ci_95, col = Population), width = 0.1, size = 1, show.legend = TRUE)+
  geom_line(aes(group = Population, col = Population), size = 1.5)+ 
  labs(y="Average Number of Seeds per Fruit", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_x_discrete(expand = c(0.6,0))+
  expand_limits(y=0)
```

fitness - avg seed num/fruit * fruit num

```{r, echo = FALSE}
ggplot(data = forplot_2021, aes(x = Treatment, y = fitness_mean)) +
  geom_point(aes(col=Population), size = 4, show.legend = TRUE) +
  geom_errorbar(aes(ymin=fitness_mean-fitness_ci_95, ymax = fitness_mean + fitness_ci_95, col = Population), width = 0.1, size = 1, show.legend = TRUE)+
  geom_line(aes(group = Population, col = Population), size = 1.5)+ 
  labs(y="Fitness", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_x_discrete(expand = c(0.6,0))+
  expand_limits(y=0)

```




## graphs by line - include?

Let's look at some things by line instead of Population.

For now, only making a graph by line for seeds per fruit because I want to see if there truly is little variation between lines within a population.The confidence intervals are decently small guess. 

```{r, eval = TRUE}
Dat_2021$Line.ID <- as.factor(paste0(Dat_2021$Population, Dat_2021$Line))

forplot2_2021 <- Dat_2021 %>%
  dplyr::group_by(Treatment) %>%
  dplyr::group_by(Line.ID, .add = TRUE) %>%
  summarize(across(.col = c(Emergence:LeafNum_08262021, RosetteLeafNum_10042021, RosetteLeafNum:DryReproG, Repro_to_Ros, LatBranches:IJ_FruitCount, AG_biomass, AvgSeedWt, AvgSeedNum, fitness), .fns = list(mean = ~mean(.x, na.rm = TRUE, .group = "keep"), ci_95 = ~conf_int(.x))))

# some lines only have 1 pot per treatment
# some lines only have pots in one treatment (3)
forplot2_2021$Population <- substr(forplot2_2021$Line.ID, start = 1, stop = 4)

# Avg Seeds per fruit
ggplot(data = forplot2_2021, aes(x = Treatment, y = AvgSeedNum_mean)) +
  geom_point(aes(col=Population), size = 2, alpha = 0.7, show.legend = TRUE) +
  #geom_errorbar(aes(group = Line.ID, ymin=AvgSeedNum_mean-AvgSeedNum_ci_95, ymax = AvgSeedNum_mean + AvgSeedNum_ci_95, col = Population), width = 0.1, size = 1, alpha = 0.7, show.legend = TRUE)+
  geom_line(aes(group = Line.ID, col = Population), size = 0.7, alpha = 0.7)+
  #geom_line(data = forplot_2021, aes(group = Population, col = Population), size= 1.5)+
  #geom_point(data = forplot_2021, aes(col = Population), size = 4, show.legend = TRUE)+
  labs(y="Average Seeds/Fruit", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_x_discrete(expand = c(0.6,0))


```

This truly shows little variation in line means within each population and treatment. huh. The error bars are very large and maybe skewing the scale so now plotting without error bars and looking at a histogram of means just to see it a different way.

```{r}
ggplot(dat = forplot2_2021)+
  geom_histogram(aes(x = AvgSeedNum_mean, fill = Population), binwidth = 3, alpha = 0.5, position = "identity")+
  scale_fill_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))

ggplot(dat = forplot2_2021)+
  geom_histogram(aes(x = AvgSeedNum_mean, fill = Treatment), binwidth = 3, alpha = 0.5, position = "identity")+
  scale_fill_manual(name = "Treatment",
                     labels = c("Current", "Future"),
                     values = c("#56B4E9", "#D55E00"))
```


# 2022 only

Graphs to make here: \
- NumEmerged \
- LeafNum_Jun6 (if included) \
- LeafNum_Jun13 (if included) \
- LeafNum_bolt \
- AG_Dry_bolt \
- BG_Dry_bolt \
- Root_to_Shoot \
- Stomata_density \

```{r}
forplot_2022 <- data.frame(t(Sum_2022))
colnames(forplot_2022) <- forplot_2022[1, ]
forplot_2022 <- forplot_2022[c(2:5), ]
forplot_2022 <- forplot_2022 %>%
  select(Population, Treatment, NumEmerged_mean, NumEmerged_ci_95, LeafNumber_Jun6_mean:LeafNumber_Jun13_ci_95, LeafNumber_Total_mean, LeafNumber_Total_ci_95, AG_DryBiomass_mean:BG_DryBiomass_ci_95, Root_to_Shoot_mean:Stomata_density_ci_95) %>%
  mutate(across(Population:Treatment, as.factor),
         across(NumEmerged_mean:Stomata_density_ci_95, as.numeric))
```

## Num Emerged
Number Emerged

```{r, echo = FALSE}
ggplot(data = forplot_2022, aes(x = Treatment, y = NumEmerged_mean)) +
  geom_point(aes(col=Population), size = 4, show.legend = TRUE) +
  geom_errorbar(aes(ymin=NumEmerged_mean-NumEmerged_ci_95, ymax = NumEmerged_mean + NumEmerged_ci_95, col = Population), width = 0.1, size = 1, show.legend = TRUE)+
  geom_line(aes(group = Population, col = Population), size = 1.5, linetype = "dotted")+ 
  labs(y="Number Emerged Seeds", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_x_discrete(expand = c(0.6,0))+
  expand_limits(y=0)
```

## Leaf Number
LeafNum_Jun6 (if included)

```{r, echo = FALSE}
ggplot(data = forplot_2022, aes(x = Treatment, y = LeafNumber_Jun6_mean)) +
  geom_point(aes(col=Population), size = 4, show.legend = TRUE) +
  geom_errorbar(aes(ymin=LeafNumber_Jun6_mean-LeafNumber_Jun6_ci_95, ymax = LeafNumber_Jun6_mean + LeafNumber_Jun6_ci_95, col = Population), width = 0.1, size = 1, show.legend = TRUE)+
  geom_line(aes(group = Population, col = Population), size = 1.5, linetype = "dotted")+ 
  labs(y="Leaf Number 1 week post vern", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_x_discrete(expand = c(0.6,0))+
  expand_limits(y=0)
```

LeafNum_Jun13 (if included)

```{r, echo = FALSE}
ggplot(data = forplot_2022, aes(x = Treatment, y = LeafNumber_Jun13_mean)) +
  geom_point(aes(col=Population), size = 4, show.legend = TRUE) +
  geom_errorbar(aes(ymin=LeafNumber_Jun13_mean-LeafNumber_Jun13_ci_95, ymax = LeafNumber_Jun13_mean + LeafNumber_Jun13_ci_95, col = Population), width = 0.1, size = 1, show.legend = TRUE)+
  geom_line(aes(group = Population, col = Population), size = 1.5, linetype = "dotted")+ 
  labs(y="Leaf Number 2 weeks post vern", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_x_discrete(expand = c(0.6,0))+
  expand_limits(y=0)
```

LeafNum_bolt

```{r, echo = FALSE}
ggplot(data = forplot_2022, aes(x = Treatment, y = LeafNumber_Total_mean)) +
  geom_point(aes(col=Population), size = 4, show.legend = TRUE) +
  geom_errorbar(aes(ymin=LeafNumber_Total_mean-LeafNumber_Total_ci_95, ymax = LeafNumber_Total_mean + LeafNumber_Total_ci_95, col = Population), width = 0.1, size = 1, show.legend = TRUE)+
  geom_line(aes(group = Population, col = Population), size = 1.5, linetype = "dotted")+ 
  labs(y="Leaf Number at Bolting", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_x_discrete(expand = c(0.6,0))+
  expand_limits(y=0)
```

Create an extra figure to look at growth curves during this time. I want one line per population and 4 x axis values (4 weeks, post vern, 2 weeks post vern) with the y axis being leaf number.

```{r}
# format a dataframe
tmp_mean <- Dat_2022 %>%
  dplyr::group_by(Population) %>%
  dplyr::group_by(Treatment, .add = TRUE) %>%
  summarize(across(.col = c(LeafNumber_PreVern:LeafNumber_Jun13, LeafNumber_Total), .fns = list(mean = ~mean(.x, na.rm = TRUE, .group = "keep"))))
# pivot
tmp_mean <- pivot_longer(tmp_mean, cols = !(c(Population, Treatment)), names_to = "Time", values_to = "mean")
tmp_mean$Time <- factor(rep(c("Week5", "Week9", "Week10", "Bolting"), times = 4), levels = c("Week5", "Week9", "Week10", "Bolting"))

tmp_ci <- Dat_2022 %>%
  dplyr::group_by(Population) %>%
  dplyr::group_by(Treatment, .add = TRUE) %>%
  summarize(across(.col = c(LeafNumber_PreVern:LeafNumber_Jun13, LeafNumber_Total), .fns = list(ci_95 = ~conf_int(.x))))

# pivot
tmp_ci <- pivot_longer(tmp_ci, cols = !(c(Population, Treatment)), names_to = "Time", values_to = "ci")
tmp_ci$Time <- factor(rep(c("Week5", "Week9", "Week10", "Bolting"), times = 4), levels = c("Week5", "Week9", "Week10", "Bolting"))

forplot3_2022 <- merge(tmp_mean, tmp_ci)

# plot
ggplot(data = forplot3_2022, aes(x = Time, y = mean)) +
  geom_point(aes(col=Population), size = 4, show.legend = TRUE) +
  #geom_errorbar(aes(ymin=mean-ci, ymax = mean + ci, col = Population), width = 0.1, size = 1, show.legend = TRUE)+
  geom_line(aes(group = Treatment:Population, col = Population, linetype = Treatment), size = 1.5)+ 
  labs(y="Leaf Number", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_x_discrete(expand = c(0.2,0))+
  expand_limits(y=0)
```


## Biomass
AG_Dry_bolt

```{r, echo = FALSE}
ggplot(data = forplot_2022, aes(x = Treatment, y = AG_DryBiomass_mean)) +
  geom_point(aes(col=Population), size = 4, show.legend = TRUE) +
  geom_errorbar(aes(ymin=AG_DryBiomass_mean-AG_DryBiomass_ci_95, ymax = AG_DryBiomass_mean + AG_DryBiomass_ci_95, col = Population), width = 0.1, size = 1, show.legend = TRUE)+
  geom_line(aes(group = Population, col = Population), size = 1.5, linetype = "dotted")+ 
  labs(y="Above Ground Biomass (g) at bolting", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_x_discrete(expand = c(0.6,0))+
  expand_limits(y=0)
```

BG_Dry_bolt

```{r, echo = FALSE}
ggplot(data = forplot_2022, aes(x = Treatment, y = BG_DryBiomass_mean)) +
  geom_point(aes(col=Population), size = 4, show.legend = TRUE) +
  geom_errorbar(aes(ymin=BG_DryBiomass_mean-BG_DryBiomass_ci_95, ymax = BG_DryBiomass_mean + BG_DryBiomass_ci_95, col = Population), width = 0.1, size = 1, show.legend = TRUE)+
  geom_line(aes(group = Population, col = Population), size = 1.5, linetype = "dotted")+ 
  labs(y="Below Ground Biomass (g) at bolting", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_x_discrete(expand = c(0.6,0))+
  expand_limits(y=0)
```

Root_to_Shoot

```{r, echo = FALSE}
ggplot(data = forplot_2022, aes(x = Treatment, y = Root_to_Shoot_mean)) +
  geom_point(aes(col=Population), size = 4, show.legend = TRUE) +
  geom_errorbar(aes(ymin=Root_to_Shoot_mean-Root_to_Shoot_ci_95, ymax = Root_to_Shoot_mean + Root_to_Shoot_ci_95, col = Population), width = 0.1, size = 1, show.legend = TRUE)+
  geom_line(aes(group = Population, col = Population), size = 1.5, linetype = "dotted")+ 
  labs(y="Root:Shoot Ratio", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_x_discrete(expand = c(0.6,0))+
  expand_limits(y=0)
```

## Stomata
Stomata Density

```{r, echo = FALSE}
ggplot(data = forplot_2022, aes(x = Treatment, y = Stomata_density_mean)) +
  geom_point(aes(col=Population), size = 4, show.legend = TRUE) +
  geom_errorbar(aes(ymin=Stomata_density_mean-Stomata_density_ci_95, ymax = Stomata_density_mean + Stomata_density_ci_95, col = Population), width = 0.1, size = 1, show.legend = TRUE)+
  geom_line(aes(group = Population, col = Population), size = 1.5, linetype = "dotted")+ 
  labs(y="Stomatal Density", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_x_discrete(expand = c(0.6,0))+
  expand_limits(y=0)
```

# both
Steps to be completed here.
Subset 2021 dataset. add 2021 column. Subset 2022 dataset. add 2022 column. make column names the same. merge. format. plot.

Reaction Norms to create here: \
- Emergence \
- EmergenceToBolting \
- FreshWt \
- HydWt \
- DryWt \
- LeafArea \
- LeafPerimeter \
- SLA \
- LDMC \
- RWC \
- LeafNum_4weeks \

```{r}
# subset 2021 dataset
tmp_1 <- data.frame(t(Sum_2021))
colnames(tmp_1) <- tmp_1[1, ]
tmp_1 <- tmp_1[c(2:5), ]
tmp_1 <- tmp_1 %>%
  select(Population, Treatment, Emergence_mean:DayToBolt_ci_95, FreshWt_g_mean:LeafNum_.07222021_ci_95) %>%
  mutate(across(Population:Treatment, as.factor),
         across(Emergence_mean:LeafNum_.07222021_ci_95, as.numeric))
tmp_1$Exp <- as.factor("2021")

# subset 2022 dataset
tmp_2 <- data.frame(t(Sum_2022))
colnames(tmp_2) <- tmp_2[1, ]
tmp_2 <- tmp_2[c(2:5), ]
tmp_2[tmp_2$Population == "B", "Population"] <- "BELM"
tmp_2[tmp_2$Population == "R", "Population"] <- "RODA"
tmp_2 <- tmp_2 %>%
  select(Population, Treatment, DaysToEmergence_mean:EmergenceToBolting_ci_95, SL_FreshWt_mean:SL_Perim_ci_95, SLA_mean:RWC_ci_95, LeafNumber_PreVern_mean, LeafNumber_PreVern_ci_95) %>%
  mutate(across(Population:Treatment, as.factor),
         across(DaysToEmergence_mean:LeafNumber_PreVern_ci_95, as.numeric))
tmp_2$Exp <- as.factor("2022")


# test columns are in the same order in both 
colnames(tmp_1)
colnames(tmp_2)

# set colnames to be the same for both
names <- c("Population", "Treatment", "Emergence_mean", "Emergence_ci", "EmergeToBolt_mean", "EmergeToBolt_ci", "Fresh_mean", "Fresh_ci", "Sat_mean", "Sat_ci", "Dry_mean", "Dry_ci", "area_mean", "area_ci", "perim_mean", "perim_ci", "SLA_mean", "SLA_ci", "LDMC_mean", "LDMC_ci", "RWC_mean", "RWC_ci","LeafNum_4wk_mean", "LeafNum_4wk_ci","Exp")

colnames(tmp_1) <- names
colnames(tmp_2) <- names

# merge
forplot_both <- rbind(tmp_1, tmp_2)
```



## Phenology

Emergence

```{r, echo = FALSE}
ggplot(data = forplot_both, aes(x = Treatment, y = Emergence_mean)) +
  geom_point(aes(group = Population:Exp, col=Population), size = 4, show.legend = FALSE, position = position_dodge2(width = 0.3)) +
  geom_errorbar(aes(group = Population:Exp, ymin=Emergence_mean-Emergence_ci, ymax = Emergence_mean + Emergence_ci, col = Population), width = 0.3, size = 1, position = position_dodge2(width = 0.3),  show.legend = FALSE)+
  geom_line(aes(group = Population:Exp, col = Population, linetype = Exp), size = 1.5, position = position_dodge(width = 0.3), show.legend = TRUE)+ 
  labs(y="Days to Emergence", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_linetype_manual(name = "Experiment",
                        labels = c("2021", "2022"),
                        values = c("solid", "dotted"))+
  scale_x_discrete(expand = c(0.6,0))+
  expand_limits(y=0)
```


EmergenceToBolting

```{r, echo = FALSE}
ggplot(data = forplot_both, aes(x = Treatment, y = EmergeToBolt_mean)) +
  geom_point(aes(group = Population:Exp, col=Population), size = 4, position = position_dodge(width = 0.3), show.legend = FALSE) +
  geom_errorbar(aes(group = Population:Exp, ymin=EmergeToBolt_mean-EmergeToBolt_ci, ymax = EmergeToBolt_mean + EmergeToBolt_ci, col = Population), width = 0.3, size = 1, position = position_dodge(width = 0.3), show.legend = FALSE)+
  geom_line(aes(group = Population:Exp, col = Population, linetype = Exp), size = 1.5, position = position_dodge(width = 0.3), show.legend = TRUE)+
  labs(y="Days Emergence to Bolting", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_linetype_manual(name = "Experiment",
                        labels = c("2021", "2022"),
                        values = c("solid", "dotted"))+
  scale_x_discrete(expand = c(0.6,0))+
  expand_limits(y=0)
```


## Single Leaf Individual Traits
FreshWt

```{r, echo = FALSE}
ggplot(data = forplot_both, aes(x = Treatment, y = Fresh_mean)) +
  geom_point(aes(group = Population:Exp, col=Population), size = 4, position = position_dodge(width = 0.3), show.legend = FALSE) +
  geom_errorbar(aes(group = Population:Exp, ymin=Fresh_mean - Fresh_ci, ymax = Fresh_mean + Fresh_ci, col = Population), width = 0.3, size = 1, position = position_dodge(width = 0.3), show.legend = FALSE)+
  geom_line(aes(group = Population:Exp, col = Population, linetype = Exp), size = 1.5, position = position_dodge(width = 0.3), show.legend = TRUE)+
  labs(y="Fresh Weight (g)", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_linetype_manual(name = "Experiment",
                        labels = c("2021", "2022"),
                        values = c("solid", "dotted"))+
  scale_x_discrete(expand = c(0.6,0))+
  expand_limits(y=0)
```


HydWt

```{r, echo = FALSE}
ggplot(data = forplot_both, aes(x = Treatment, y = Sat_mean)) +
  geom_point(aes(group = Population:Exp, col=Population), size = 4, position = position_dodge(width = 0.3), show.legend = FALSE) +
  geom_errorbar(aes(group = Population:Exp, ymin=Sat_mean - Sat_ci, ymax = Sat_mean + Sat_ci, col = Population), width = 0.3, size = 1, position = position_dodge(width = 0.3), show.legend = FALSE)+
  geom_line(aes(group = Population:Exp, col = Population, linetype = Exp), size = 1.5, position = position_dodge(width = 0.3), show.legend = TRUE)+
  labs(y="Saturated Weight (g)", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_linetype_manual(name = "Experiment",
                        labels = c("2021", "2022"),
                        values = c("solid", "dotted"))+
  scale_x_discrete(expand = c(0.6,0))+
  expand_limits(y=0)
```


DryWt

```{r, echo = FALSE}
ggplot(data = forplot_both, aes(x = Treatment, y = Dry_mean)) +
  geom_point(aes(group = Population:Exp, col=Population), size = 4, position = position_dodge(width = 0.3), show.legend = FALSE) +
  geom_errorbar(aes(group = Population:Exp, ymin=Dry_mean - Dry_ci, ymax = Dry_mean + Dry_ci, col = Population), width = 0.3, size = 1, position = position_dodge(width = 0.3), show.legend = FALSE)+
  geom_line(aes(group = Population:Exp, col = Population, linetype = Exp), size = 1.5, position = position_dodge(width = 0.3), show.legend = TRUE)+
  labs(y="Dry Weight (g)", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_linetype_manual(name = "Experiment",
                        labels = c("2021", "2022"),
                        values = c("solid", "dotted"))+
  scale_x_discrete(expand = c(0.6,0))+
  expand_limits(y=0)
```


LeafArea

```{r, echo = FALSE}
ggplot(data = forplot_both, aes(x = Treatment, y = area_mean)) +
  geom_point(aes(group = Population:Exp, col=Population), size = 4, position = position_dodge(width = 0.3), show.legend = FALSE) +
  geom_errorbar(aes(group = Population:Exp, ymin=area_mean - area_ci, ymax = area_mean + area_ci, col = Population), width = 0.3, size = 1, position = position_dodge(width = 0.3), show.legend = FALSE)+
  geom_line(aes(group = Population:Exp, col = Population, linetype = Exp), size = 1.5, position = position_dodge(width = 0.3), show.legend = TRUE)+ 
  labs(y="Leaf Area (cm^2)", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_linetype_manual(name = "Experiment",
                        labels = c("2021", "2022"),
                        values = c("solid", "dotted"))+
  scale_x_discrete(expand = c(0.6,0))+
  expand_limits(y=0)
```


LeafPerimeter

```{r, echo = FALSE}
ggplot(data = forplot_both, aes(x = Treatment, y = perim_mean)) +
  geom_point(aes(group = Population:Exp, col=Population), size = 4, position = position_dodge(width = 0.3), show.legend = FALSE) +
  geom_errorbar(aes(group = Population:Exp, ymin=perim_mean - perim_ci, ymax = perim_mean + perim_ci, col = Population), width = 0.3, size = 1, position = position_dodge(width = 0.3), show.legend = FALSE)+
  geom_line(aes(group = Population:Exp, col = Population, linetype = Exp), size = 1.5, position = position_dodge(width = 0.3), show.legend = TRUE)+
  labs(y="Leaf Perimeter (cm)", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_linetype_manual(name = "Experiment",
                        labels = c("2021", "2022"),
                        values = c("solid", "dotted"))+
  scale_x_discrete(expand = c(0.6,0))+
  expand_limits(y=0)
```



## Single Leaf Calculated Traits
Specific Leaf Area

```{r, echo = FALSE}
ggplot(data = forplot_both, aes(x = Treatment, y = SLA_mean)) +
  geom_point(aes(group = Population:Exp, col=Population), size = 4, position = position_dodge(width = 0.3),show.legend = FALSE) +
  geom_errorbar(aes(group = Population:Exp, ymin=SLA_mean - SLA_ci, ymax = SLA_mean + SLA_ci, col = Population), width = 0.3, size = 1, position = position_dodge(width = 0.3), show.legend = FALSE)+
  geom_line(aes(group = Population:Exp, col = Population, linetype = Exp), size = 1.5, position = position_dodge(width = 0.3), show.legend = TRUE)+
  labs(y="Specific Leaf Area (cm^2 / g)", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_linetype_manual(name = "Experiment",
                        labels = c("2021", "2022"),
                        values = c("solid", "dotted"))+
  scale_x_discrete(expand = c(0.6,0))+
  expand_limits(y=0)
```


Leaf Dry Matter Content

```{r, echo = FALSE}
ggplot(data = forplot_both, aes(x = Treatment, y = LDMC_mean)) +
  geom_point(aes(group = Population:Exp, col=Population), size = 4, position = position_dodge(width = 0.3),show.legend = FALSE) +
  geom_errorbar(aes(group = Population:Exp, ymin=LDMC_mean - LDMC_ci, ymax = LDMC_mean + LDMC_ci, col = Population), width = 0.3, size = 1, position = position_dodge(width = 0.3),show.legend = FALSE)+
  geom_line(aes(group = Population:Exp, col = Population, linetype = Exp), size = 1.5, position = position_dodge(width = 0.3), show.legend = TRUE)+
  labs(y="Leaf Dry Matter Content (%)", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_linetype_manual(name = "Experiment",
                        labels = c("2021", "2022"),
                        values = c("solid", "dotted"))+
  scale_x_discrete(expand = c(0.6,0))+
  expand_limits(y=0)
```


Relative Water Content

```{r, echo = FALSE}
ggplot(data = forplot_both, aes(x = Treatment, y = RWC_mean)) +
  geom_point(aes(group = Population:Exp, col=Population), size = 4, position = position_dodge(width = 0.3), show.legend = FALSE) +
  geom_errorbar(aes(group = Population:Exp, ymin=RWC_mean - RWC_ci, ymax = RWC_mean + RWC_ci, col = Population), width = 0.3, size = 1, position = position_dodge(width = 0.3), show.legend = FALSE)+
  geom_line(aes(group = Population:Exp, col = Population, linetype = Exp), size = 1.5, position = position_dodge(width = 0.3), show.legend = TRUE)+
  labs(y="Relative Water Content (%)", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_linetype_manual(name = "Experiment",
                        labels = c("2021", "2022"),
                        values = c("solid", "dotted"))+
  scale_x_discrete(expand = c(0.6,0))
  #expand_limits(y=0)
```

## Leaf Number
LeafNum_4weeks

```{r, echo = FALSE}
ggplot(data = forplot_both, aes(x = Treatment, y = LeafNum_4wk_mean)) +
  geom_point(aes(group = Population:Exp, col=Population), size = 4, position = position_dodge(width = 0.3), show.legend = FALSE) +
  geom_errorbar(aes(group = Population:Exp, ymin=LeafNum_4wk_mean - LeafNum_4wk_ci, ymax = LeafNum_4wk_mean + LeafNum_4wk_ci, col = Population), width = 0.3, size = 1, position = position_dodge(width = 0.3), show.legend = FALSE)+
  geom_line(aes(group = Population:Exp, col = Population, linetype = Exp), size = 1.5, position = position_dodge(width = 0.3), show.legend = TRUE)+
  labs(y="Leaf Number (4 weeks post planting)", 
      x="Treatment")+
  scale_color_manual(name = "Population",
                     labels = c("IT", "SW"),
                     values = c("#CC79A7", "#009E73"))+
  scale_linetype_manual(name = "Experiment",
                        labels = c("2021", "2022"),
                        values = c("solid", "dotted"))+
  scale_x_discrete(expand = c(0.6,0))+
  expand_limits(y=0)
```